---
title: "PipseqPTSubcluster_V0822"
author: "Xin Wang"
date: "2023-08-22"
description: Proximal tubule subclusters analyses:
  - Subclusters of PT
  - Annotation of subclusters using the candidate markers
  - Psudotime of PT
  - Basic statistics of PT
output: html_document
---
# Loading the packages
```{r}
# check the session information
sessionInfo()
library(dplyr)
library(Seurat)
library(patchwork)

#install.packages('devtools') #assuming it is not already installed

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
library(devtools)
#install_github('andreacirilloac/updateR')
# install.packages("sctransform")
# install.packages("ggthemes")
# loading the libarary
library(updateR)

library("sctransform")
library("ggthemes")
library(Seurat)

# BiocManager::install("limma")
# BiocManager::install("ComplexHeatmap")
# BiocManager::install("biomaRt")
# BiocManager::install("methrix")BiocManager::install("org.Hs.eg.db")
# devtools::install_github("saeyslab/nichenetr")
library(limma)
library(nichenetr)
library(tidyverse)
library(dplyr)

library(biomaRt)
library("data.table")


library(umap)
library(patchwork)
library(cowplot)
library(installr)

#install.packages("Matrix", repos = "http://cran.r-project.org")
### we used discre color palettes from ggsci

library("ggsci")
library("ggplot2")
library("gridExtra")

#options(buildtools.check = function(action) TRUE )
#devtools::install_url('https://cran.r-project.org/src/contrib/Archive/Matrix.utils/Matrix.utils_0.9.7.tar.gz')
#devtools::install_github('cole-trapnell-lab/monocle3')
#library(monocle3)

library(harmony)
library(Seurat)
#library(SeuratData)
library(tidyverse)

#library(BPCells)
library(ggplot2)
library("devtools")
library("AnnotationDbi")
library("org.Hs.eg.db")
#library(tibble)
#library(future)
library(here)
library(patchwork)
library(future)
library(monocle3)
#devtools::install_dev("remotes")
#remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
#remotes::install_github(repo ='chris-mcginnis-ucsf/DoubletFinder' )
#install.packages("DoubletFinder")
library(DoubletFinder)
library(future.callr)
library(ComplexHeatmap)

library(RColorBrewer)
library(circlize)
library(monocle3)
#install.packages("magick")
library(magick)
#K means with 6 groups

```


## Step 2: Set up the workspace environment:
```{r}
rm(list=ls())
setwd("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/PT_Subclusters/")
Indir <-c("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/")
Outdir <-c( "/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/PT_Subclusters/")
# setting up the seed 
set.seed(10000)
```


# generate the health cluster of PT cells

```{r}

# Read the preclustered object

scrna <- readRDS(file = paste0(Indir,"Pyelonephritis_singlecell_harmony_withAnnotation_v0803.RDS"))

# We only collect the PT cells
#CellType[CellType$cluster %in% c("1_1","0", "9","12","18", "19","3"),2] <- "PT" # "Proximal tubule"

scrna @meta.data

PTcells <- subset(scrna, subset = Raw_cell_type == "PT")

PTcells@meta.data
# We then recluster the PT cells
# recluster the cells
PTcells <- SCTransform(PTcells, verbose = TRUE)
PTcells <- RunPCA(PTcells,npcs = 5, verbose = TRUE)

# Run harmony for the PT cells
PTcells <- RunHarmony(PTcells, 
				group.by.vars = c("DataSet"), 
				reduction = "pca", assay.use = "SCT", reduction.save = "harmony")
PTcells <- RunUMAP(PTcells, dims = 1:5, verbose = TRUE, reduction = "harmony")
PTcells <- FindNeighbors(PTcells, dims = 1:5,  verbose = TRUE, reduction = "harmony")

PTcells<- FindClusters(PTcells,  verbose = TRUE, resolution = 0.3)

PTcellsDim<- DimPlot(PTcells,label = T)
PTcellsDim
ggsave(filename = "PTcellsDim.pdf", plot = PTcellsDim, width = 8, height = 6)

PTcellsDimSplit<-DimPlot(PTcells, split.by = "DataSet",label = T)
PTcellsDimSplit
ggsave(filename = "PTcellsDimSplit.pdf", plot = PTcellsDimSplit, width = 24, height = 6)

```

# check the markers
```{r}
# the healthy cells
HealthMarkers<- c("Slc22a30", "Atp11a", "Trim7", "Slc22a6", "Slc7a13", "Slc5a2", "Slc5a10", "Spp2",  "Gatm", "Dcxr")

DotPlot(PTcells, features =  HealthMarkers, col.min = -0.5) +RotatedAxis()

CyclingMarker <- "Mki67"
DotPlot(PTcells, features =  CyclingMarker, col.min = -0.5) +RotatedAxis()

EarlyInjuryMarkers <- c("Cdh13", "Cdh6", "Havcr1")
DotPlot(PTcells, features =  EarlyInjuryMarkers, col.min = -0.5) +RotatedAxis()
LateInjuryMarkers <-c( "Ccl2", "Gm17268", "Cald1")
DotPlot(PTcells, features =  LateInjuryMarkers, col.min = -0.5) +RotatedAxis()

ReparingMarkers <- "Top2a"

DotPlot(PTcells, features =  ReparingMarkers, col.min = -0.5) +RotatedAxis()

FailedFepairMarkers <- c("Ypel2","Kcnip4")

DotPlot(PTcells, features =  FailedFepairMarkers, col.min = -0.5) +RotatedAxis()
# 
MaladptiveMarkers <- c("Vcam1", "Havcr1",  "Icam1", "Prom1","Cxcl2","Ccl3","Myc")

DotPlot(PTcells, features =  MaladptiveMarkers, col.min = -0.5) +RotatedAxis()


# Combined 
#"Lrp2","Slc34a1","Slc13a1"
KidneyStatuesMarkers<- c( "Slc5a2","Slc34a1","Inmt","Apom", "Slc7a13","Slc22a6", "Prom1","Top2a","Cyp7b1","Mki67","Cdh13", "Cdh6", "Havcr1","Ccl2", "Gm17268", "Cald1","Vcam1", "Ypel2","Kcnip4","Hsp90aa1","Col27a1","Plin2")

 DotPlot(PTcells, features =  KidneyStatuesMarkers,col.min = -1) +RotatedAxis()

KidneyStatuesTypeMarkers<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "Spp1", "Cst3","Gdf15", "Clu", "Igfbp7", "Apoe", "Aldob", "B2m", "S100a6","S100a8","S100a9","Pdgfb","Bak1","Bad","Casp8","Ptgs2","Casp1","Hmox1")
 DotPlot(PTcells, features =  KidneyStatuesTypeMarkers) +RotatedAxis()
 
 KidneyStatuesTypeMarkers2<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "S100a6","S100a8","S100a9","Pdgfb","Bak1","Bad","Casp8","Ptgs2","Casp1","Hmox1","Nfkbia")
  DotPlot(PTcells, features =  KidneyStatuesTypeMarkers2) +RotatedAxis()

 #, cluster 5: Fn1, Havcr1, Cst3, Igfbp7, S100a8, Il1b, Slc7a13 : Profibrotic PT
 # cluster 1: Spp1, Prom1, Kcnip4, Col27a1,Top2a,Nfkbia: Cycling PT
 # cluster 8: Cdh13, Hsp90aa1,Inmt, Mki67 : Proliferating PT
 
 # cluster 1: ASpp1, Prom1, Kcnipp4: Progenitor PT
 
 # cluster 0: Aldob
```

# the biomarkers for the 
```{r}
CandidateMarkers <- c("Aldob","Slc34a1","Spp1","Spp2","Lcn2", "Slc22a8","Slc5a2","Inmt","Prom1","Defb1","Top2a","Nfkbia", "Kcnip4", "Slc22a6", "Apoe","Havcr1", "Fn1","C3", "Il1b","S100a8", "Igfbp7", "Cst3", "Col27a1", "Hsp90aa1","Cdh13","Slc4a11","Slc3a1","Ren1","Slc12a1","Umod","Klk1")

CandidateMarkersDotPlot<-  DotPlot(PTcells, features =  CandidateMarkers, col.min = -2) +RotatedAxis()

ggsave("PT_CandidateMarkersDotPlot.pdf", plot = CandidateMarkersDotPlot, height = 6, width = 12)

```

# biomarker from other cell types
```{r}
# Cfh: Parietal epithelial cells
Finalcelltypemarkers<- c("Lrp2","Slc34a1","Slc13a3", # Proximal tubule cells (PT)
                      "Slc12a3","Pvalb","Wnk1",# Distal connecting tublue (DCT)

                     "Umod","Slc12a1","Cldn10", # Thick ascending limb cells (TAL)
                      "Klk1","Slc8a1","Calb1", # Connecting tubule (CNT)
                     "Slc14a2","Fst","Bst1", # descending thin  limb cells (DTL)
                     #"Akr1b1","Sh3gl3","Prox1", # ascending thin limb cells (ATL)
                     #"Enox1","Thsd4", #Macula densa
                      "Aqp2","Hsd11b2","Scnn1g", # principal cells (PC)
                    "Atp6v1g3","Aqp6","Slc26a7", # Intercalated cells
                    "Krt5","Upk1b","Krt14", # Urothelium
                    "Cdh5","Igfbp3","Pecam1", # Endothelium
                    "Col3a1","Fbn1","Lum","Col1a1", # Fibroblast
                    "Ireb2","Alas2","C78334", # Plasmacytoid dentritic cells
                    "S100a8","S100a9","Il1b", # Neutrophil
                    "C1qa","C1qb","Aif1", # Macrophage
                    "Cxcr6","Cd247","Cd3e", # T cells
                    "Igkc","Cd79a","Cd79b", # B cells
                    "Ccl5","Nkg7","Cd7" # NK/Cd8 cells
)

OtherMarkerGenesDotPlots<- DotPlot(PTcells, features =  Finalcelltypemarkers, col.min = -0.5) +RotatedAxis()
ggsave(filename = "OtherMarkerGenesExpression_ClusterAnnotate.pdf", plot = OtherMarkerGenesDotPlots, width = 15, height = 8)

```

# #####################################
## generate the Trajectories of PT cells
# #####################################
# 2.2 Prepared cds the object
```{r}
# build cds object

PTcells_expr_matrix = GetAssayData(PTcells, slot = "counts")

# Cell meta annotation
p_data <- PTcells@meta.data 
head(p_data)
# Gene meta annotation
f_data<- data.frame(gene_short_name = row.names(PTcells), row.names = row.names(PTcells))


pre_cds <- new_cell_data_set(PTcells_expr_matrix,
                         cell_metadata = p_data,
                         gene_metadata = f_data)
pData(pre_cds) %>% colnames()

cds <- preprocess_cds(pre_cds) #标准化+PCA降维

```

# 2.2 Umap by Monocle3
```{r}
cds <- reduce_dimension(cds, preprocess_method = "PCA")
head(cds@int_colData$reducedDims$UMAP)

plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="SCT_snn_res.0.2")
dev.off()
# UMAP by Seurat
seurat_umap <- Embeddings(PTcells, reduction = "umap")[colnames(cds),]
cds@int_colData$reducedDims$UMAP <- seurat_umap
head(cds@int_colData$reducedDims$UMAP)
plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="SCT_snn_res.0.2")

```

# Draw the monocle state
```{r}
#分群(类似monocle的State)

cds <- cluster_cells(cds)

#预测轨迹
cds <- learn_graph(cds)
#交互式确定root节点，可以选择多个。我这里选择了一个
cds <- order_cells(cds) 

PT_pseudotime <-plot_cells(cds, color_cells_by = "pseudotime", 
           label_cell_groups = FALSE, 
           label_leaves = TRUE,  
           label_branch_points = FALSE)

ggsave(filename = "PT_pseudotime_MonocaCluster.pdf", plot = PT_pseudotime, width = 4, height = 3)


```

```{r}
# Determine the psedudotime related genes
Track_genes <- graph_test(cds, neighbor_graph="principal_graph")

Track_genes %>% head()

```

# plot the expression of interesting genes during the PT pseudotime
```{r}

# re-order the tracked genes by the q value and morans I
Track_genes <- Track_genes[,c(5,2,3,4,1,6)] %>% 
  dplyr::arrange(desc(morans_I),q_value)

# Plot some interesting genes for the PT pseudotime : Top 2 genes, Napsa, Kap, Ttr, Igfbp7, Havcr1 and Prom1
KapTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[1],] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_KapTrack.pdf", plot = KapTrack, width = 5, height = 3)
NapsaTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[2],] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_NapsaTrack.pdf", plot = NapsaTrack, width = 5, height = 3)

TtrTrack <- plot_genes_in_pseudotime(cds["Ttr",] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_TtrTrack.pdf", plot = TtrTrack, width = 5, height = 3)

Igfbp7Track <- plot_genes_in_pseudotime(cds["Igfbp7",] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_Igfbp7Track.pdf", plot = Igfbp7Track, width = 5, height = 3)

Track_genes$gene_short_name[1:5]


Track_genes$gene_short_name[2]
Track_genes$gene_short_name[1]
Havcr1Track <- plot_genes_in_pseudotime(cds["Havcr1",] , 
                         min_expr=0.5)

ggsave(filename = "PT_pseudotime_Havcr1Track.pdf", plot = Havcr1Track, width = 5, height = 3)

# For the 
Prom1Track <- plot_genes_in_pseudotime(cds["Prom1",],min_expr=0.5 )
ggsave(filename = "PT_pseudotime_Prom1Track.pdf", plot = Prom1Track, width = 5, height = 3)

FeaturePlot(PTcells, features = "Havcr1")

```

# choose the top signifcant for visualization
```{r}
diff_pseudo_gene <- Track_genes %>% 
  dplyr::arrange(q_value) %>%
  # remove the genes with mitochondrion
  filter(!grepl('^mt-|Hbb-', gene_short_name)) %>%
  rownames() %>% head(60)

# we ignore the mitochondria genes



diff_pseudo_gene

# generate the heatmap for those genes

diff_pseudo_gene.matrix <- exprs(cds)[match(diff_pseudo_gene,rownames(rowData(cds))),order(pseudotime(cds))]

diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){smooth.spline(x,df=3)$y}))
diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){(x-mean(x))/sd(x)}))
rownames(diff_pseudo_gene.matrix) <- diff_pseudo_gene;

#pheatmap(diff_pseudo_gene.matrix)

# because the plot_pseudotime_heatmap no longer have, we used the complexheatmap to draw the figure
# Reference: https://github.com/cole-trapnell-lab/monocle-release/issues/295
pdf("PT_pseudotime_psudotimeGenes.pdf", height = 6, width = 8)
htkm <- Heatmap(
  diff_pseudo_gene.matrix,
  name    = "z-score",
  col     = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names  = TRUE,
  show_column_names = FALSE,
  row_names_gp  = gpar(fontsize = 6),
  km = 8,
  row_title_rot = 0,
  use_raster=TRUE,
  cluster_rows  = TRUE,
  cluster_row_slices = FALSE,
  cluster_columns   = FALSE)

print (htkm)
dev.off()
#ggsave(filename = "PT_pseudotime_psudotimeGenes.pdf", plot = htkm, width = 8, height = 6)


#Ward.D2 Hierarchical Clustering
pdf("PT_pseudotime_psudotimeGenes_Hierarchicalcluster.pdf", height = 7, width = 6)
hthc <- Heatmap(
  diff_pseudo_gene.matrix,
  name                         = "z-score",
  col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names               = TRUE,
  show_column_names            = FALSE,
  row_names_gp                 = gpar(fontsize = 6),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  use_raster= TRUE,
  cluster_rows                 = TRUE,
  cluster_row_slices           = FALSE,
  cluster_columns              = FALSE)

print(hthc)
dev.off()

cds["Havcr1",] %>% head()
# Generate the Dotplot for these genes
plot_genes_in_pseudotime(cds[c("Havcr1"),] , 
                         min_expr=0.5)

DotPlot(PTcells, features =diff_pseudo_gene ) +RotatedAxis()

DotPlot(PTcells, features =c("Spp1","Havcr1","Itgb1","Top2a",
                             "Prom1","Pdgfb","Defb1","Fn1","Col6a1"), col.min = -0.5 ) +RotatedAxis()

FeaturePlot(PTcells, features = c("Lcn2","Defb1","Egf","Havcr1"))
# the following no longer have 

Track_genes
# p = plot_pseudotime_heatmap(cds[diff_pseudo_gene,], 
#                         num_clusters = 3,  # default 6
#                         return_heatmap=T)
# 

```




# generate the subclusters of PT, DCT, IC, PC
```{r}
# dividing into cycling, transitioning, adap- tive (successful or maladaptive repair) and degenerative (damaged or stressed). 
# cycling: Mki67
# Early Injury: Cdh13, Cdh6, Havcr1
# Late Injury: Ccl2, Gm17268, Cald1
# conserved: Slc13a1, Slc5a2,  Slc7a13, Cyp7b1



# Panoptosis
# Il1b, MLKL, Gsdmd,Ripk3

# Fibrosis 
# Fn1, Col1a1, Col3a1, Tgfb1
# adaptive
# VCAM1, DCDC2 and HAVCR1, C3, ICAM1
# VCAM1 PROM1 ITGB8
# Vcam1, Havcr1, Dcdc2, Icam1, Prom1, ITgb8

#degenerative
# SPP1, CST3, CLU and IGFBP7, APOE ALDOB
# B2M S100A6
# Spp1, Cst3, GDF15,Clu, Igfbp7, Apoe, Aldob, B2m, S100a6
# agionegious PEC,
# EGFR, VEGF, Cdh5

# Egfr, Vegf, Cdh5
KidneyStatuesMarkers<- c("Slc13a1", "Slc5a2", "Slc7a13", "Cyp7b1","Mki67","Cdh13", "Cdh6", "Havcr1","Ccl2", "Gm17268", "Cald1")

KidneyStatuesTypeMarkers<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "Spp1", "Cst3","Gdf15", "Clu", "Igfbp7", "Apoe", "Aldob", "B2m", "S100a6")

DotPlot(scrna, features =  KidneyStatuesMarkers, col.min = -0.5) +RotatedAxis()
DotPlot(scrna, features =  KidneyStatuesTypeMarkers, col.min = -0.5) +RotatedAxis()


for (i in KidneyStatuesMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}

for (i in KidneyStatuesTypeMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesTypeMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}


# with the inflammation markers

# with the fibrosis markers

# with the apoptosis markers

# with the necrosis markers

# with the endothelial markers


```


