---
title: "PipseqPTSubclusterDEG_GOKEGG_V0822"
author: "Xin Wang"
date: "2023-09-03"
description: Proximal tubule subclusters analyses:
  - DEGs from various clusters and timepoints
  - Their GO enrihments and KEGG Pathways
output: html_document
---
# Loading the packages
```{r}
# check the session information
sessionInfo()
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

library(dplyr)
library(Seurat)
library(patchwork)
library(devtools)
library(updateR)

library("sctransform")
library("ggthemes")
library(Seurat)
library(limma)
library(nichenetr)
library(tidyverse)
library(dplyr)
library(biomaRt)
library("data.table")


library(umap)
library(patchwork)
library(cowplot)
library(installr)
library("ggsci")
library("ggplot2")
library("gridExtra")
library(harmony)
library(Seurat)
#library(SeuratData)
library(tidyverse)
library(tibble)
#library(BPCells)
library(ggplot2)
library("devtools")
library("AnnotationDbi")
library("org.Hs.eg.db")
#library(tibble)
#library(future)
library(here)
library(patchwork)
library(future)
library(monocle3)
library(readr)
library(DoubletFinder)
library(future.callr)
library(ComplexHeatmap)

library(RColorBrewer)
library(circlize)
library(monocle3)
library(magick)
library(progeny)
library(pheatmap)
library(decoupleR)
library(clusterProfiler)
```


## Set up the workspace environment:
```{r}
rm(list =ls())
setwd("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/PT_Subclusters/")
Indir <-c("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/PT_Subclusters/")
Outdir <-c( "/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/PT_Subclusters/")
```

# Create 
```{r}
# create folders for DEGs, Progeny, GO, KEGG and TF analyses
# directory for DEGs
DEGDir <- paste0(Outdir,"DEGs/")
# directory for Progeny
ProgenyDir <- paste0(Outdir,"Progeny/")
# directory for GO enrichments
GOenrichDir <- paste0(Outdir,"GOenrich/")
# directory for Transcript Factors
TFDir <-  paste0(Outdir,"TranscriptFactors/")
# directory for KEGG
KEGGDir <-  paste0(Outdir,"KEGG/")
# directory for CellCommunication
CellCommDir <-  paste0(Outdir,"CellCommunication/")
# directory for pyseudotime
PyseudotimeDir <-  paste0(Outdir,"Pyseudotime/")

# directory for biomarkers

BioMarkerDir <- paste0(Outdir,"Biomarkers/")

TimeBioMarkerDir <- paste0(Outdir,"TimePointBiomarkers/")

dir.create(DEGDir, showWarnings = FALSE)
dir.create(ProgenyDir, showWarnings = FALSE)
dir.create(GOenrichDir, showWarnings = FALSE)
dir.create(TFDir, showWarnings = FALSE)
dir.create(KEGGDir, showWarnings = FALSE)
dir.create(CellCommDir, showWarnings = FALSE)
dir.create(BioMarkerDir, showWarnings = FALSE)
dir.create(PyseudotimeDir, showWarnings = FALSE)
dir.create(TimeBioMarkerDir, showWarnings = FALSE)

set.seed(10000)
```

# Read the cluster of PT cells
```{r}
# Read the preclustered object
scrna <- readRDS(file = paste0(Indir,"PIPSeq_Annotated_v0823.rds"))

# We only collect the PT cells
PTcells <- readRDS(file = paste0(Indir,"PIPSeq_Annotated_OnlyPT_v0823.rds"))

PTcells @meta.data %>% head()
# set up the PT cells as the number of cluster
Idents(PTcells) <- PTcells @meta.data$DataSet
Idents(PTcells)
```

# Biomarker detection from PT that caused by infections
```{r}
DefaultAssay(PTcells) <- "RNA"
#SubPTBiomarkerDir <- paste0(BioMarkerDir, "SubPTBiomarker")
#dir.create(TimeBioMarkerDir)
setwd(TimeBioMarkerDir)
PTcells <- ScaleData(object = PTcells, features = rownames(PTcells))
# MAST was used to calculate the cluster
# Identify the markers that are differentially expressed from controls
Dates<-levels(Idents(PTcells))
DateMarkers<- list()
Top5UpMarkers <-list()
Top5DownMarkers <-list()
AllInfectUpMarkers <-list()
AllInfectDownMarkers<-list()
for (m in c(2:4)) {
  #m=2
  SpecifiDate <- Dates[[m]]
  # detect the differentially expressed genes compared to controls
  DateMarkers[[SpecifiDate]] <-
    FindMarkers(
      PTcells,
      ident.1 = SpecifiDate,
      ident.2 = "0DPI",
      logfc.threshold = 0.25,
      test.use = "MAST",
      min.pct = 0.25
    )
  # for upregulated genes
  # we ignore the mitochondriondplyr::arrange(q_value) %>%
  # remove the genes with mitochondrion
  
  Top5UpMarkers[[SpecifiDate]] <-
    DateMarkers[[SpecifiDate]] %>% filter (avg_log2FC > 0.25) %>%  slice_max(order_by = avg_log2FC, n = 5) %>% row.names()
  Top5DownMarkers[[SpecifiDate]] <-
    DateMarkers[[SpecifiDate]] %>% filter (avg_log2FC < -0.25) %>%  slice_max(order_by = -avg_log2FC, n = 5) %>% row.names()
  
  AllInfectUpMarkers[[SpecifiDate]] <-
    DateMarkers[[SpecifiDate]] %>% rownames_to_column('GeneID') %>% filter (avg_log2FC >= 0.25 & p_val_adj <= 0.05 & !grepl("mt-", GeneID)) %>% pull(GeneID)
  
  # AllInfectUpMarkers[[SpecifiDate]]
  
  AllInfectDownMarkers[[SpecifiDate]] <-
    DateMarkers[[SpecifiDate]] %>% rownames_to_column('GeneID') %>% filter (avg_log2FC <= -0.25 & p_val_adj <= 0.05 & !grepl("mt-", GeneID)) %>% pull(GeneID)
  # Top5DownMarkers[[SpecifiDate]]
  # save all the specific date into files
  write.csv(DateMarkers[[SpecifiDate]],
            file = paste0(SpecifiDate, "InfectTimePTBiomarker.csv"),)
  
}

#DateMarkers[["3DPI"]] %>% head()
```


# using the top 5 biomarkers and see their distribution, Dotplot, heatmap and featurePlot 
```{r}
setwd(TimeBioMarkerDir)
# select the top biomarkers that are in response to infection in PT cells
FinalTopMarkers<- c(Top5UpMarkers %>% unlist() %>% unique(),Top5DownMarkers %>% unlist() %>% unique())
PTInfectionMarkersDotPlot<-DotPlot(PTcells, features =FinalTopMarkers ) +RotatedAxis() +coord_flip()

ggsave(
  "PTInfectionMarkersDotPlot.pdf",
  plot = PTInfectionMarkersDotPlot,
  height = 9,
  width = 6
)

PTInfectionMarkersDotheatmap<-DoHeatmap(PTcells, features = FinalTopMarkers , slot = "scale.data") + scale_fill_gradientn(colors = c("blue", "white", "red")) + theme(axis.text.y = element_text(size = 10))
ggsave(
  "PTInfectionMarkersHeatmap.pdf",
  plot = PTInfectionMarkersDotheatmap,
  height = 9,
  width = 7
)

# create a fold to store all these biomarker featureplots
FeaturePTBiomarkerDir <- paste0(TimeBioMarkerDir, "PTInfectBiomarkers")
dir.create(FeaturePTBiomarkerDir)
setwd(FeaturePTBiomarkerDir)

for (i in FinalTopMarkers) {
  PTmarkerFeatureSplit <-
    FeaturePlot(PTcells,
                features = i,
                split.by = "DataSet",
                slot = "scale.data")
  PTmarkerFeature <-
    FeaturePlot(PTcells, features = i, slot = "scale.data")
  
  ggsave(
    paste0(i, "_PTInfectedBiomarkerFeatureSplit.pdf"),
    plot = PTmarkerFeatureSplit,
    height = 3,
    width = 12
  )
  
  ggsave(
    paste0(i, "_PTSubBiomarkerFeature.pdf"),
    plot = PTmarkerFeature,
    height = 3,
    width = 4
  )
}

```

# Perform the GO enrichment and see the specific pathways in each categories
```{r}
TimeGOenrichDir<- paste0(TimeBioMarkerDir,"GOenrich")
dir.create(TimeGOenrichDir)
setwd(TimeGOenrichDir)

# Create a GO comparison for the figure
# for the up-regulated genes
# we have stored the up-regulated genes into the AllInfectUpMarkers

# Comparing the Cluster
PTInfectedSpecifcUpgeneGOenrich <-
  compareCluster(
    geneClusters = AllInfectUpMarkers,
    fun = "enrichGO",
    OrgDb = "org.Mm.eg.db",
    ont = "BP",
    keyType = 'SYMBOL',
    pvalueCutoff=1
  )

write.csv(PTInfectedSpecifcUpgeneGOenrich,
          "PTInfectedSpecifcUpgeneGOenrich.csv")

PTInfectedSpecifcUpgeneGOenrichDotPlot <-
  dotplot(
    PTInfectedSpecifcUpgeneGOenrich,
    font.size = 12,
    label_format = 80,
    showCategory = 15,
    #includeAll=FALSE
  )
? dotplot
#low = "LightBlue", midpoint=0, mid="white", high = "FireBrick",
ggsave(
  "PTInfectedSpecifcUpgeneGOenrichDotPlot.pdf",
  plot = PTInfectedSpecifcUpgeneGOenrichDotPlot,
  height = 11,
  width = 8
)

## for the downregualted genes

# we have stored the down-regulated genes into the AllInfectDownMarkers

# Comparing the Cluster
PTInfectedSpecifcDowngeneGOenrich <-
  compareCluster(
    geneClusters = AllInfectDownMarkers,
    fun = "enrichGO",
    OrgDb = "org.Mm.eg.db",
    ont = "BP",
    keyType = 'SYMBOL',
    pvalueCutoff=1
  )

write.csv(PTInfectedSpecifcDowngeneGOenrich,
          "PTInfectedSpecifcDowngeneGOenrich.csv")

PTInfectedSpecifcDowngeneGOenrichDotPlot <-
  dotplot(
    PTInfectedSpecifcDowngeneGOenrich,
    font.size = 12,
    label_format = 80,
    showCategory = 15,
    #includeAll=FALSE
  )
? dotplot
#low = "LightBlue", midpoint=0, mid="white", high = "FireBrick",
ggsave(
  "PTInfectedSpecifcDowngeneGOenrichDotPlot.pdf",
  plot = PTInfectedSpecifcDowngeneGOenrichDotPlot,
  height = 11,
  width = 8
)

```

# KEGG pathway 
```{r}

TimeKEGGDir <- paste0(TimeBioMarkerDir,"KEGG/")
dir.create(TimeKEGGDir)
setwd(TimeKEGGDir)

KEGGpathways <-list()
KEGGpathwaysForView<-list()
PTInfectedDEGTransferUnirotID <-list()
PTInfectedDEGTransferENTREZID <-list()
## put the list of UNIPROT into the list
# we generate genes for the kegg pathview
#as.data.frame(PTsubclusterBiomarkers)
library("org.Mm.eg.db")
PTInfecteds <- levels(Idents(PTcells))
TransferIDs<-list()
AllInfectDEGMarkers<-list()
#SectionDEGlistWithInfection
for (m in c(2:4)) {

  DateName <- PTInfecteds[[m]]
  DateName
  AllInfectDEGMarkers[[DateName]] <-
   DateMarkers[[SpecifiDate]] %>% rownames_to_column('GeneID') %>% filter ((avg_log2FC >= 0.25|avg_log2FC <= -0.25) & p_val_adj <= 0.05 & !grepl("mt-", GeneID)) %>% pull(GeneID)
  length(AllInfectDEGMarkers[[DateName]])
  AllInfectDEGMarkers[[DateName]] %>% unique()
  
  
  # transfer biolgical ID to KEGG ID
  TransferIDs[[DateName]] <-
    bitr(
      AllInfectDownMarkers[[DateName]],
      fromType = "SYMBOL",
      toType = c("UNIPROT", "ENSEMBL", "ENTREZID"),
      OrgDb = "org.Mm.eg.db"
    )
  #nrow(TransferIDs[[DateName]])
  TransferIDs[[DateName]] %>% group_by(SYMBOL)
  # we only select the first uniprot id. because the one ID have multiple uniprot ID
  PTInfectedDEGTransferUnirotID[[DateName]] <-
    TransferIDs[[DateName]] %>% group_by(SYMBOL) %>% slice(1) %>% pull(UNIPROT)
  #   mapIds(org.Mm.eg.db, keys=AllInfectDownMarkers[[DateName]], column='UNIPROT', keytype='SYMBOL', multiVals="first") %>% as.matrix() %>% as.data.frame()
  # #length(PTInfectedDEGTransferUnirotID[[DateName]]$V1)
  # as.data.frame(as.matrix(PTInfectedDEGTransferUnirotID[[DateName]]))
  #  #nrow(PTInfectedDEGTransferUnirotID[[DateName]]) 
  
  #length(PTInfectedDEGTransferUnirotID[[DateName]])
  
  PTInfectedDEGTransferENTREZID[[DateName]] <-
    TransferIDs[[DateName]] %>% group_by(SYMBOL) %>% slice(1) %>% pull(ENTREZID)
  
  #
  # egID = bitr(
  #   PTInfectedSpecifcUpgene[[m]] ,
  #   fromType = "SYMBOL",
  #   toType = "ENTREZID",
  #   OrgDb = "org.Mm.eg.db"
  # )
  PTInfectedDEGTransferUnirotID[[DateName]]
  KEGGpathwaysForView[[DateName]] <- enrichKEGG(gene =  PTInfectedDEGTransferENTREZID[[DateName]],
                                                organism = 'mmu',
                                                pvalueCutoff = 0.05)
  KEGGpathways[[DateName]] <-
    enrichKEGG(
      gene =  PTInfectedDEGTransferUnirotID[[DateName]],
      organism = 'mmu',
      pvalueCutoff = 1,
      keyType = 'uniprot'
    )
  
  write.csv(KEGGpathways[[DateName]],
            file = paste0(TimeKEGGDir, DateName, "_KeggDEGEnrichpath.csv"))
  
  
  
}

```

## compare the KEGG pathways among different time Infection
```{r}

length(PTInfectedDEGTransferUnirotID$`3DPI`)
length(PTInfectedDEGTransferUnirotID$`7DPI`)
length(PTInfectedDEGTransferUnirotID$`28DPI`)
length(AllInfectUpMarkers$`3DPI`)
length(AllInfectUpMarkers$`7DPI`)
length(AllInfectUpMarkers$`28DPI`)

AllInfectUpMarkers
# compare all DEG list with GO enrichment, we did not set up the p value
PTInfectedDEGTransferKEGG <-
  compareCluster(
    geneCluster = PTInfectedDEGTransferUnirotID,
    fun = enrichKEGG,
    organism = "mmu",
    keyType = 'uniprot',
    pvalueCutoff = 1
  )
PTInfectedDEGTransferKEGG %>% head()
# setting into readable ID
PTInfectedDEGTransferKEGGReadable <-
  setReadable(PTInfectedDEGTransferKEGG,
              OrgDb = "org.Mm.eg.db",
              keyType = "UNIPROT")

# remove the string - Mus musculus (house mouse)
PTInfectedDEGTransferKEGGReadable@compareClusterResult$Description <-
  gsub(
    pattern = " - Mus musculus (house mouse)",
    replacement = "",
    PTInfectedDEGTransferKEGG@compareClusterResult$Description,
    fixed = T
  )

# Show all the top20 pathways
Top20Comparion <-
  dotplot(
    PTInfectedDEGTransferKEGGReadable,
    font.size = 12,
    label_format = 80,
    showCategory = 20
  )

ggsave(
  "PTInfectedDEGTransferKEGG_DotPlot.pdf",
  plot = Top20Comparion,
  height = 9,
  width = 7
)

write.csv(PTInfectedDEGTransferKEGGReadable, file = "PTInfectedDEGTransferKEGGReadable.csv")
```


```{r}

```


































```{r}
# save the top ID

# show the candidate pathway in the infected regions
selectedPathway<- c("Carbon metabolism","Glyoxylate and dicarboxylate metabolism","Galactose metabolism","Cell adhesion molecules","ECM-receptor interaction","Focal adhesion","Tight junction","Necroptosis","Complement and coagulation cascades","Apoptosis","Endocytosis","Lysosome","Hippo signaling pathway","Phagosome","Neutrophil extracellular trap formation","Chemokine signaling pathway","Hematopoietic cell lineage","Cytokine-cytokine receptor interaction","NOD-like receptor signaling pathway","Osteoclast differentiation","NF-kappa B signaling pathway","Th17 cell differentiation","Antigen processing and presentation","Inflammatory bowel disease","Toll-like receptor signaling pathway","Human cytomegalovirus infection","TNF signaling pathway","C-type lectin receptor signaling pathway","Th1 and Th2 cell differentiation","IL-17 signaling pathway","Fc gamma R-mediated phagocytosis","Cytosolic DNA-sensing pathway","Natural killer cell mediated cytotoxicity","B cell receptor signaling pathway","JAK-STAT signaling pathway","Intestinal immune network for IgA production","Leukocyte transendothelial migration","T cell receptor signaling pathway","Rap1 signaling pathway","Bacterial invasion of epithelial cells","Fc epsilon RI signaling pathway","p53 signaling pathway")

# select these interesting pathway
SelectedPathwayForfigure<-SectionDEGTransferKEGGReadable %>% as.data.frame() %>% filter(Description %in% selectedPathway) %>% arrange(p.adjust)

# setting up the minium p adjust value for size visulization 
maximum<- max(c(-log10(SelectedPathwayForfigure$p.adjust)))

## set the levels of Description and reorder them
SelectedPathwayForfigure$Description<- ordered(SelectedPathwayForfigure$Description, levels=selectedPathway)
#reorder(variable, value)
p<-ggplot(SelectedPathwayForfigure, aes(x=SelectedPathwayForfigure$Cluster, y=SelectedPathwayForfigure$Description))

InterestingComparionsDotplot<- p+ geom_point(aes(size=-log10(p.adjust)),
                                             colour= ifelse(SelectedPathwayForfigure$p.adjust <=0.05,"red",
                                                            ifelse(SelectedPathwayForfigure$p.adjust==1,"white","gray")), 
                                             fill= ifelse(SelectedPathwayForfigure$p.adjust<=0.05,"red",
                                                          ifelse(SelectedPathwayForfigure$p.adjust==1,"white","gray"))) +
  labs(title = "KEGG Pathway Enrichment", x = "Spatial region", y = "KEGG pathway enrichment") +
  theme_bw()+
  theme(plot.background = element_blank(),
        text = element_text(size=10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1)
  ) + scale_size_area(limits=c(0,maximum),breaks=c(0,4,6,8,12))


InterestingComparions <- dotplot(SectionDEGTransferKEGGReadable,font.size = 12,label_format = 80,showCategory = selectedPathway)  +RotatedAxis()

ggsave(paste0(KEGGoutdir, "InterestingPathwayComparion_Sections_KEGG_comp.pdf"), plot = InterestingComparions, width = 8, height = 10)
ggsave(paste0(KEGGoutdir, "InterestingPathwayComparion_Sections_KEGG_comp_format.pdf"), plot = InterestingComparionsDotplot, width = 8, height = 10)
```




```{r}
# Overall DEGs compared between various date to Day 0
Idents(FinalAbcessClustersObjSubSlides) <- FinalAbcessClustersObjSubSlides@meta.data$orig.ident
Idents(FinalAbcessClustersObjSubSlides) 

# check the specific marker by the MAST 
# take a long time
AllClustermarkersTimepoints <- FindAllMarkers(FinalAbcessClustersObjSubSlides,
                min.pct = 0.25,test.use= "MAST") # only markers that are in 25% of cells
# save all biomarkers as a dataframe
write.csv(AllClustermarkersTimepoints, "AllTimepointsMarkers_Mast.csv")

# Generate a table for figure
DEGDifferentTimepointsnumbersFigure<- table(AllClustermarkersTimepoints$cluster) %>% as.data.frame() %>% 
  ggplot(aes(x=Var1, y=Freq)) +geom_bar(stat="identity") +theme_classic() +xlab("Timepoints") +ylab("Differentially expressed gene number")

ggsave("DEGCTimepointSpecificnumbersFigure.pdf", plot = DEGDifferentTimepointsnumbersFigure, height = 4, width = 6)

```



## read the biomarker and plot the heatmap for top5 biomarker in each cluster
```{r}
# we have already generated the DEGs for each cluster, so we uploaded them now ...
AllClustermarkersTimepoints<- read.csv(file = "AllTimepointsMarkers_Mast.csv", header = T)

# reorder the levels in identity if needed
# reduce data, we only choose the top 10 in each cluster for the visualization
top_AllClustermarkersTimpoints <- as.data.frame(AllClustermarkersTimepoints %>%
                group_by(cluster) %>% # group by cluster (cell type)
                slice_max(avg_log2FC, n = 10)) # sort by log2FC and get top 5
top_AllClustermarkersTimpoints 
# change the order of timepoints match by the Idents orders, 

# reorder 
TimepiointOrder<-c("PyeloD0_1", "PyeloD1_2", "PyeloD3_1", "PyeloD5_2", "PyeloD7_1", "PyeloD28_1")

top_AllClustermarkersTimpointsReorder<- top_AllClustermarkersTimpoints[order(match(top_AllClustermarkersTimpoints$cluster, TimepiointOrder)),]

#top_AllClustermarkersTimpointsReorder$cluster

# explore top markers
mapal <- colorRampPalette(RColorBrewer::brewer.pal(11,"RdBu"))(256)
Top10TimepointsHeatmap<-DoHeatmap(object = FinalAbcessClustersObjSubSlides, features = top_AllClustermarkersTimpointsReorder$gene) + scale_fill_gradientn(colours = rev(mapal))

Top10TimepointsHeatmap<-DoHeatmap(object = FinalAbcessClustersObjSubSlides, features = top_AllClustermarkersTimpointsReorder$gene) +  scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick"))

#low = "LightBlue", midpoint=0, mid="white", high = "FireBrick",
ggsave(paste0("Top10_Cluster_marker_Heatmap_Timpoints.pdf"), plot = Top10TimepointsHeatmap, height = 15, width = 20 )
```

# generate the spatial feature plot for each biomarkers detected in each timepoint
```{r}

for (Cls in TimepiointOrder){
  
  # generate a folder for the clusters
  ClusterFilePath<- paste0(outdir,Cls,"_FeaturePlot/")
  dir.create(file.path(ClusterFilePath))
  setwd(file.path(ClusterFilePath))
  top_AllClustermarkersReorder %>% head()
  # generate the list of top 5 biomarker
  top5Timpointsbiomarker <- top_AllClustermarkersTimpointsReorder %>% filter(cluster == Cls) %>% pull(gene) %>% unique()
  top5Timpointsbiomarker
  # generate the feature plots for each biomarkers
  
  for (t in top5Timpointsbiomarker){
    
     TimepointSpecificSpatilaFeatureVariouCategoriesCut5 <- 
          SpatialFeaturePlot(FinalAbcessClustersObjSubSlides,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
     # save into a file
    ggsave(filename = paste0(Cls,t, "_SpatialFeatureplot_Cut5.pdf"), 
       plot =TimepointSpecificSpatilaFeatureVariouCategoriesCut5, height = 8, width = 15 )
    
    TimepointSpecificSpatilaFeatureVariouCategoriesCut8 <- 
          SpatialFeaturePlot(FinalAbcessClustersObjSubSlides,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,8))
     # save into a file
    ggsave(filename = paste0(Cls,t, "_SpatialFeatureplot_Cut8.pdf"), 
       plot =TimepointSpecificSpatilaFeatureVariouCategoriesCut8, height = 8, width = 15 )
    
     TimepointSpecificSpatilaFeatureVariouCategoriesCut10 <- 
          SpatialFeaturePlot(FinalAbcessClustersObjSubSlides,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,10))
     # save into a file
    ggsave(filename = paste0(Cls,t, "_SpatialFeatureplot_Cut10.pdf"), 
       plot =TimepointSpecificSpatilaFeatureVariouCategoriesCut10, height = 8, width = 15 )
 
  }
  
}

```




# perform the GO enrichment for each clusters
```{r}
SectionDEGlistTimepoints <- list()
TimepointsUniqUpGO<-list()
TimepointsUniqUpGOSimplified <-list()
# save the gene list into Timepoints order
for (n in TimepiointOrder){
  # put he DEG into the DEGlist for comparision
  # we only consider the Cortex, Medulla, Pelvis 
  #if( m == "Cortex" | m == "Medulla" | m == "Pelvis" | m== "Processing Infection"){
     SectionDEGlistTimepoints[[n]] <- AllClustermarkersTimepoints %>% filter(cluster %in% n & avg_log2FC >= 0.25) %>% pull(gene)
      # TimepointsUniqUpGO[[n]] <- enrichGO( SectionDEGlistTimepoints[[n]], OrgDb = "org.Mm.eg.db", ont = 'BP', pAdjustMethod = 'BH', pvalueCutoff = 1, qvalueCutoff = 1, keyType = 'SYMBOL')
      # write.csv(TimepointsUniqUpGO[[n]], file = paste0("Timepoints_uniq_gene.", n, ".UpregulatedGO.csv"))
      # 
      # TimepointsUniqUpGOSimplified[[n]] <- clusterProfiler::simplify(TimepointsUniqUpGO[[n]])
      # write.csv(TimepointsUniqUpGOSimplified[[n]], file = paste0("Timepoints_uniq_gene.", n, ".UpregulatedGO.simplifed.csv"))
  #}
 }
#TimepointsUniqUpGO
```


```{r}
length(SectionDEGlistTimepointss)

# Comparing the Timepoints 
SectionDEGlistGOPlotPerTimepoints <- compareCluster(geneClusters = SectionDEGlistTimepoints, fun = "enrichGO", OrgDb = "org.Mm.eg.db",ont="BP",keyType = 'SYMBOL')

?compareCluster 
# simplify the GO enrichment
SectionDEGlistGOPlotPerTimepointsSimplifed <- clusterProfiler::simplify(SectionDEGlistGOPlotPerTimepoints)

write.csv(SectionDEGlistGOPlotPerTimepoints,"All_Timepointsmarkers_Mast_GO.csv")

SectionDEGlistGOPlotPerTimepointsDotPlot<- dotplot(SectionDEGlistGOPlotPerTimepoints,font.size = 12,label_format = 80,showCategory = 10) 
?dotplot
#low = "LightBlue", midpoint=0, mid="white", high = "FireBrick",
ggsave(paste0("SectionDEGlistGOPlotPerTimepointsSimplifedDotPlot.pdf"), plot = SectionDEGlistGOPlotPerTimepointsDotPlot, height = 15, width = 15 )

```

# generate the 
```{r}

DefaultAssay(allen_reference) <- "RNA"

D0<-subset(allen_reference, subset = DataSet =="D0")
Biomarkers<- FindAllMarkers(object = D0, logfc.threshold = 0.25, min.diff.pct =.2 )
## save into the table 
write.csv(Biomarkers, file = sprintf("%s/Biomarkers.Wilcox.D0.csv", outdir) )
## choose the top 10 DEGs in each cluster and print them into heatmaps

### comparing between cluster 0 and 4
TwoPTBiomarkers<- FindMarkers(object = D0, ident.1 = c(0,1,5,13), ident.2 = 4, logfc.threshold = 0.25, min.diff.pct =.2,test.use ="MAST" )

?FindMarkers
## check the GO enrichment
BimaarkerGO<- enrichGO(rownames(TwoPTBiomarkers), OrgDb = "org.Mm.eg.db", ont='BP',pAdjustMethod = 'BH',pvalueCutoff = 1, qvalueCutoff = 1,keyType = 'SYMBOL')

pdf(file="PromimalTubule_Biomarkers_GO.pdf", width=9, height=4)
GOfigure<-barplot(BimaarkerGO,label_format = 80)
print(GOfigure)
dev.off()


FirstPTBiomarkers<- FindMarkers(object = D0, ident.1 = 0, ident.2 = NULL, logfc.threshold = 0.25, min.diff.pct =.2 )
FirstPTBiomarkers

pdf(file="TwoProximalTubuleCellComp_Biomarker_ForHeatmap_D0.pdf",height = 6, width = 10)
DoHeatmap(subset(allen_reference, subset = DataSet =="D0"), features = rownames(TwoPTBiomarkers)) + scale_fill_gradientn(colors = c("blue", "white", "red")) + theme(axis.text.y = element_text(size = 10))
dev.off()

```


# #####################################
## generate the Progeny signaling pathways
# #####################################

```{r}
# set up the directory
setwd(ProgenyDir)
## we set up with a high memory

options(future.globals.maxSize = 891289600000) 
## We compute the Progeny activity scores and add them to our Seurat object
## as a new assay called Progeny. 

# create cluster data frame with the cells, Celltypes
PTInfected<- data.frame(Cell = names(Idents(PTcells)),   
                              CellType = as.character(Idents(PTcells)),
                              stringsAsFactors = FALSE)

PTcells <- progeny(PTcells, scale=FALSE, organism="Mouse", top=500, perm=1, return_assay = TRUE)
PTcells
# setup the default assay as progeny
DefaultAssay(object = PTcells) <- "progeny"

PTcells <- Seurat::ScaleData(PTcells, assay = "progeny") 
## We can now directly apply Seurat functions in our Progeny scores. 
## For instance, we scale the pathway activity scores. 
progeny_scores_df <- 
  as.data.frame(t(GetAssayData(PTcells, slot = "scale.data", 
                               assay = "progeny"))) %>%
  rownames_to_column("Cell") %>%
  gather(Pathway, Activity, -Cell) 

progeny_scores_df %>% filter(Cell =="0DPI_AAAAAAAACCACCAGA")
## We match Progeny scores with the cell clusters.
progeny_scores_df <- inner_join(progeny_scores_df, PTInfected)

progeny_scores_df %>% head()
## We summarize the Progeny scores by cellpopulation
summarized_progeny_scores <- progeny_scores_df %>% 
  group_by(Pathway, CellType) %>%
  summarise(avg = mean(Activity), std = sd(Activity))
summarized_progeny_scores


## We prepare the data for the plot
summarized_progeny_scores_df <- summarized_progeny_scores %>%
  dplyr::select(-std) %>%   
  spread(Pathway, avg) %>%
  data.frame(row.names = 1, check.names = FALSE, stringsAsFactors = FALSE) 

paletteLength = 80
myColor = colorRampPalette(c("royalblue", "white","lightcoral"))(paletteLength)

progenyBreaks = c(seq(min(summarized_progeny_scores_df), 0, 
                      length.out=ceiling(paletteLength/2) + 1),
                  seq(max(summarized_progeny_scores_df)/paletteLength, 
                      max(summarized_progeny_scores_df), 
                      length.out=floor(paletteLength/2)))
ProximalTubuleCellForHeatmap<-as.data.frame(t(summarized_progeny_scores_df[,-1]))

# generate the heatmap for PT
pdf(file="ProximalTubuleCell_Progeny_ForHeatmap.pdf",height = 5, width = 5)
progeny_hmap = pheatmap(ProximalTubuleCellForHeatmap,fontsize=8, 
                        fontsize_row = 8,  show_rownames=T, cluster_cols=T, cluster_rows=T, scale="row",
                        cex=1, clustering_distance_rows="euclidean", cex=1,
                        # clustering_distance_cols="euclidean", clustering_method="complete", 
                        color=myColor, breaks = progenyBreaks, 
                        main = "PROGENy (500)", angle_col = 45,
                        border_color = NA,
                        treeheight_col = 0,   treeheight_row=0)

print(progeny_hmap)
dev.off()
test<- FeaturePlot(PTcells, features = "NFkB")
ggsave("test_NFKB.pdf", plot = test, width = 5, height = 5)
# generate the featureplot 
ProgenyFeatures <- summarized_progeny_scores$Pathway %>% unique()
ProgenyFeatures
for (i in ProgenyFeatures ){
  p2 <-(FeaturePlot(PTcells, features = i,split.by = "DataSet") & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red'))& 
  ggtitle(paste0(i," activity"))
  
  ggsave(paste0(i, "_PT_cluster_featurePlot.pdf"), plot = p2, width = 20, height = 5)
}


```

# Transcription factor activity inference from Immune scRNA-seq
```{r}
# CollecTRI network
# set up the directory
setwd(TFDir)
netTF <-  readRDS("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Datasets/get_collectri_human_mouse20230716/get_collectri_mouse_process_table.rds")

mat <- as.matrix(PTcells@assays$RNA@data)
# Run ulm
acts <- run_ulm(mat=mat, net=netTF, .source='source', .target='target',
                .mor='mor', minsize = 5)

saveRDS(acts, file="PT_Transcriptfactor.RDS")
# Activity inference with Univariate Linear Model (ULM)
# Extract ulm and store it in tfsulm in pbmc
PTcells[['tfsulm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = PTcells) <- "tfsulm"

# Scale the data
PTcells <- ScaleData(PTcells)
PTcells@assays$tfsulm@data <- PTcells@assays$tfsulm@scale.data

# Explore the top 20 more varaible TFs
n_tfs <- 25

t(as.matrix(PTcells@assays$tfsulm@data))
# Extract activities from object as a long dataframe
df <- t(as.matrix(PTcells@assays$tfsulm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(PTcells)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Get top tfs with more variable means across clusters
tfs <- df %>%
  group_by(source) %>%
  summarise(std = sd(mean)) %>%
  arrange(-abs(std)) %>%
  head(n_tfs) %>%
  pull(source)
# drop "Znf335" due to not exist in the expression
#tfs <- tfs[!tfs == "Znf335"]

#tfs[!tfs == "Znf335"]
# Subset long data frame to top tfs and transform to wide matrix
top_acts_mat <- df %>%
  filter(source %in% tfs) %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-3, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 3, length.out=floor(palette_length/2)))

# Plot
PTTFheatmap<- pheatmap(t(top_acts_mat),border_color = NA, cluster_cols = F, 
                        color=my_color,fontsize=12, fontsize_row = 10, 
                        breaks = my_breaks, 
                        #main = "Predicted TF activities", 
                        angle_col = 45,
                        treeheight_col = 0, treeheight_row=0)


ggsave("PTcellsHeatmap_Transcriptfactor_Progeny_Decouple.pdf", plot = PTTFheatmap, width = 4, height = 5)

# draw the transcriptor factor activity and their expression
# generate the figure of transcript factor activity:
for (m in tfs){
  p2 <- (FeaturePlot(PTcells, features = m) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) +
  ggtitle(paste0(m, ' activity'))
  
  ggsave(paste0("TFActivity_",m,"_PTcells_FeaturePlot.pdf"), plot = p2, width = 4, height = 3)
}

DefaultAssay(object = PTcells) <- "RNA"

for (n in tfs){
  p3 <- (FeaturePlot(PTcells, features = n) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) +
  ggtitle(paste0(n, ' expression'))
  
  ggsave(paste0("TFExpression_",n,"_PTcells_FeaturePlot.pdf"), plot = p3, width = 4, height = 3)

}
# add their activity in Dotplot
Idents(PTcells)
# get the order of row name
GeneOrderByPheatmap <- rownames(t(top_acts_mat))[PTTFheatmap$tree_row$order]
#IMMTFheatmap$tree_row$order

?DotPlot
TFDotPlot <- DotPlot(PTcells, features = rev(GeneOrderByPheatmap)) +
  RotatedAxis() + coord_flip()+ 
  #ggtitle("Predicted TF expression")+
   theme(legend.position = "right")
ggsave("Immune_TFDotPlot.pdf", plot = TFDotPlot, width = 5, height = 6)

```


# Using PROGENy model from DecoupleR
```{r}
setwd(ProgenyDir)
net <- get_progeny(organism = 'mouse', top = 500)

# Activity inference with Multivariate Linear Model (MLM)
# Extract the normalized log-transformed counts
mat <- as.matrix(PTcells@assays$RNA@data)

# Run mlm
acts <- run_mlm(mat=mat, net=net, .source='source', .target='target',
                .mor='weight', minsize = 5)

acts
# Visualization:
# Extract mlm and store it in pathwaysmlm in data
PTcells[['pathwaysmlm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)


# Change assay
DefaultAssay(object = PTcells) <- "pathwaysmlm"

# Scale the data
PTcells <- ScaleData(PTcells)
PTcells@assays$pathwaysmlm@data <- PTcells@assays$pathwaysmlm@scale.data

# run the source of pathways
ProgenyFeatures <- levels(factor(acts$source))
ProgenyFeatures
for (i in ProgenyFeatures ){
  p2 <-FeaturePlot(PTcells, features = i) & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0(i," activity"))
  
  ggsave(paste0(i, "_PT_cluster_featurePlot.pdf"), plot = p2, width = 6, height = 5)
}

#(c("royalblue", "white","lightcoral"))
FeaturePlot(PTcells, features = "Hypoxia", min.cutoff = -4, max.cutoff = 6, split.by = "DataSet") & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0("Hypoxia"," activity"))

FeaturePlot(PTcells, features = "NFkB", min.cutoff = -4, max.cutoff = 6, split.by = "DataSet") & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0("NFkB"," activity"))


FeaturePlot(PTcells, features = "TNFa", min.cutoff = -4, max.cutoff = 6, split.by = "DataSet") & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0("TNFa"," activity"))

p1 <- DimPlot(PTcells, reduction = "umap", label = TRUE, pt.size = 0.5) + 
  NoLegend() + ggtitle('Cell types')
p2 <- (FeaturePlot(PTcells, features = c("Trail")) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) +
  ggtitle('Trail activity')



# Expore the mean activiey per group across pathways:
# Extract activities from object as a long dataframe
df <- t(as.matrix(PTcells@assays$pathwaysmlm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(PTcells)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Transform to wide matrix
top_acts_mat <- df %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-0.5, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 0.5, length.out=floor(palette_length/2)))

# Plot
PTcellsHeatmap<- pheatmap(top_acts_mat, border_color = NA, color=my_color, breaks = my_breaks) 
ggsave("PTcellsHeatmap_Progeny.pdf", plot = PTcellsHeatmap, width = 6, height = 6)

```

# Transcription factor activity inference from PT scRNA-seq
```{r}
# CollecTRI network
# set up the directory
setwd(TFDir)
netTF <-  readRDS("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Datasets/get_collectri_human_mouse20230716/get_collectri_mouse_process_table.rds")

mat <- as.matrix(PTcells@assays$RNA@data)
# Run wmean
acts <- run_ulm(mat=mat, net=netTF, .source='source', .target='target',
                .mor='mor', minsize = 5)

# Activity inference with Univariate Linear Model (ULM)
# Extract ulm and store it in tfsulm in pbmc
PTcells[['tfsulm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = PTcells) <- "tfsulm"

# Scale the data
PTcells <- ScaleData(PTcells)
PTcells@assays$tfsulm@data <- PTcells@assays$tfsulm@scale.data

p1 <- DimPlot(PTcells, reduction = "umap", label = TRUE, pt.size = 0.5) + 
  NoLegend() + ggtitle('Cell types')
p2 <- (FeaturePlot(PTcells, features = c("PAX5")) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) +
  ggtitle('PAX5 activity')
DefaultAssay(object = PTcells) <- "RNA"
p3 <- FeaturePlot(data, features = c("PAX5")) + ggtitle('PAX5 expression')
DefaultAssay(object = PTcells) <- "tfswmean"
p1 | p2 | p3

# Explore the top 20 more varaible TFs
n_tfs <- 25
# Extract activities from object as a long dataframe
df <- t(as.matrix(PTcells@assays$tfswmean@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(PTcells)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Get top tfs with more variable means across clusters
tfs <- df %>%
  group_by(source) %>%
  summarise(std = sd(mean)) %>%
  arrange(-abs(std)) %>%
  head(n_tfs) %>%
  pull(source)

# Subset long data frame to top tfs and transform to wide matrix
top_acts_mat <- df %>%
  filter(source %in% tfs) %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-3, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 3, length.out=floor(palette_length/2)))

# Plot
pheatmap(top_acts_mat, border_color = NA, color=my_color, breaks = my_breaks) 

```

# generate the subclusters of PT, DCT, IC, PC
```{r}
# dividing into cycling, transitioning, adap- tive (successful or maladaptive repair) and degenerative (damaged or stressed). 
# cycling: Mki67
# Early Injury: Cdh13, Cdh6, Havcr1
# Late Injury: Ccl2, Gm17268, Cald1
# conserved: Slc13a1, Slc5a2,  Slc7a13, Cyp7b1



# Panoptosis
# Il1b, MLKL, Gsdmd,Ripk3

# Fibrosis 
# Fn1, Col1a1, Col3a1, Tgfb1
# adaptive
# VCAM1, DCDC2 and HAVCR1, C3, ICAM1
# VCAM1 PROM1 ITGB8
# Vcam1, Havcr1, Dcdc2, Icam1, Prom1, ITgb8

#degenerative
# SPP1, CST3, CLU and IGFBP7, APOE ALDOB
# B2M S100A6
# Spp1, Cst3, GDF15,Clu, Igfbp7, Apoe, Aldob, B2m, S100a6
# agionegious PEC,
# EGFR, VEGF, Cdh5

# Egfr, Vegf, Cdh5
KidneyStatuesMarkers<- c("Slc13a1", "Slc5a2", "Slc7a13", "Cyp7b1","Mki67","Cdh13", "Cdh6", "Havcr1","Ccl2", "Gm17268", "Cald1")

KidneyStatuesTypeMarkers<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "Spp1", "Cst3","Gdf15", "Clu", "Igfbp7", "Apoe", "Aldob", "B2m", "S100a6")

DotPlot(scrna, features =  KidneyStatuesMarkers, col.min = -0.5) +RotatedAxis()
DotPlot(scrna, features =  KidneyStatuesTypeMarkers, col.min = -0.5) +RotatedAxis()


for (i in KidneyStatuesMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}

for (i in KidneyStatuesTypeMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesTypeMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}


# with the inflammation markers

# with the fibrosis markers

# with the apoptosis markers

# with the necrosis markers

# with the endothelial markers


```


