---
title: "SpatialDEGFigure_RegionalKidney"
author: "Xin Wang"
date: "2023-07-20"
output: html_document
---
---
title: "Differentially expressed genes detection for comparisions within timepoints to control in spatial transcriptomic datasets using all the niches"
output: html_document
Author: Xin Wang
Email: xin.wang@nationwidechildrens.org 
Copyright (c) 2023 Kidney and Urology Tract Center
Nationwide Children's Hospital
date: "2023-05-18"
Description: 
#   Detection of differentially expressed genes for different infection regions
#    - Detection of DEGs comparing the clusters
#    - Detection of DEGs under several different conditions
#    - Functional enrichments for the DEGs
#    - Several of associated figures to show the differences
#    - KEGG pathway enrichments
#    - Heatmap, spatialfeature plots, pathview of interesting KEGG pathways.

note: the genes including the list of
       cell type: Proximal tubule, connecting tubule, IC, PC, Loop of henle, Endothelium, NK, CD45, T cells, B Cells, Macrophage, Neutorphil

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = TRUE)
```


## loading the libraries

```{r libraries, cache=FALSE, warning=FALSE, error=FALSE, message=FALSE,}

# loading libraries required for seurat
library(Seurat)
#library(SeuratData)
library(ggplot2)
library(patchwork)
library(dplyr)
library(Matrix)
library(tidyverse)
library(SeuratObject)
#install.packages("harmony")
library(harmony)
library(SPOTlight)
# library(scater)
# library(scran)
library(SingleCellExperiment)

#install.packages("SeuratDisk")
#library(SeuratDisk)

# loading for the table bubble updates
library(L1pack)
library(reshape)
library(tidyverse)
library(stringr)
library(cowplot)

# loading the libraries for plot and colors
library(ggplot2)
library(reshape2)
library(Hmisc)
library(stats)
library(corrplot)
library(RColorBrewer)
## Deconvolve
#install.packages("L1pack")
#install.packages("reshape")

# loading libraries required for the DEG and functionally enrichment
#install.packages("msigdbr")
library(msigdbr)
library(fgsea)
library(data.table)
library(RColorBrewer)
library(ggrepel)
library(DESeq2)
library(pheatmap)
#BiocManager::install("MAST")
library(MAST)
library(edgeR)
library(clusterProfiler)
library(org.Mm.eg.db)
library(enrichplot)
#keytypes(org.Mm.eg.db) 

#BiocManager::install("ReactomePA")
library(ReactomePA)

# install thje KEGG packages
#MusKEGG<- search_kegg_organism('Mus musculus', by='scientific_name')
#dim(MusKEGG)
#head(MusKEGG)
# loading for KEGG analyes
library("pathview")


library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))

```


# The main analyses section:
# setting up the environments and randomization
```{r}
## set directory pathway:

rm(list=ls())
setwd("/Users/XXW004/Documents/Projects/RuizRosado/SpatialTranscriptomic/Results/NichesDEGs_V09062023/KidneyAllNiches/")
Indir = "/Users/XXW004/Documents/Projects/RuizRosado/SpatialTranscriptomic/Results/Deconvolution_Six_V09062023/"
# Here is the Kidney different regions
Outdir =c("/Users/XXW004/Documents/Projects/RuizRosado/SpatialTranscriptomic/Results/NichesDEGs_V09062023/KidneyAllNiches/")

# Generate the Different folders to store the results
DEGDir <- paste0(Outdir,"DEGs/")
# directory for Progeny
ProgenyDir <- paste0(Outdir,"Progeny/")
# directory for GO enrichments
GOenrichDir <- paste0(Outdir,"GOenrich/")
# directory for Transcript Factors
TFDir <-  paste0(Outdir,"TranscriptFactors/")
# directory for KEGG
KEGGDir <-  paste0(Outdir,"KEGG/")
# directory for CellCommunication
CellCommDir <-  paste0(Outdir,"CellCommunication/")
# directory for pyseudotime
PyseudotimeDir <-  paste0(Outdir,"Pyseudotime/")

# directory for biomarkers

BioMarkerDir <- paste0(Outdir,"Biomarkers/")

dir.create(DEGDir, showWarnings = FALSE)
dir.create(ProgenyDir, showWarnings = FALSE)
dir.create(GOenrichDir, showWarnings = FALSE)
dir.create(TFDir, showWarnings = FALSE)
dir.create(KEGGDir, showWarnings = FALSE)
dir.create(CellCommDir, showWarnings = FALSE)
dir.create(BioMarkerDir, showWarnings = FALSE)
dir.create(PyseudotimeDir, showWarnings = FALSE)
# setting up the seed 
set.seed(1000001)
```

# read the input object that have been well convoluted and annotated
```{r}
Pyelonephritis_Infection <- readRDS(file = paste0(Indir,"Pyelonephritis_Infection_Deconvoluted.RDS"))

rownames(Pyelonephritis_Infection)
colnames(Pyelonephritis_Infection)
# define the selcted slices
selectedslices<- c("slice1","slice3","slice4","slice7","slice8","slice9")

Pyelonephritis_Infection
# define infected niches
Infectedniches <- c("N12","N13","N14","N15","N16","N17")


```

# choose the subset of niches for the following analysis
```{r}
# Define the default assay as SCT 
DefaultAssay(Pyelonephritis_Infection) <- "SCT"

# select the subset of spatial transcriptomics
## we only choose the slides from PyeloD0_1, PyeloD1_2, PyeloD3_1, PyeloD5_1, PyeloD7_1, PyeloD28_1
#SlidesSelected_Condtion<-c("PyeloD0_1", "PyeloD1_1", "PyeloD3_1", "PyeloD5_2", "PyeloD7_1", "PyeloD28_1")

Idents(Pyelonephritis_Infection)
FinalAbcessClustersObjSubSlidesInfection_subcluster<- subset(Pyelonephritis_Infection, idents = c("N12","N13","N14","N15","N16","N17"))
# Reorder the identity levels

Idents(FinalAbcessClustersObjSubSlidesInfection_subcluster)

DimPlot(Pyelonephritis_Infection)
?SpatialDimPlot
#DefaultAssay(FinalAbcessClustersObjSubSlidesNonInfection_Subcluster) <- "SCT"
pdf("SpatialIntegrated_Pyeloneprhitis_Infected_area_nocrop.pdf", height =12, width = 6)
SpatialDimPlot(FinalAbcessClustersObjSubSlidesInfection_subcluster, images =selectedslices,crop = FALSE, label = FALSE,ncol = 2)
dev.off()
pdf("SpatialIntegrated_Pyeloneprhitis_Infected_area_dimplot.pdf", height =4, width = 5)
DimPlot(FinalAbcessClustersObjSubSlidesInfection_subcluster,label = T)
dev.off()
```

# Detection all the DEGs for all the regions and only show the top 10 for N13-N17
```{r}
setwd(BioMarkerDir)
DefaultAssay(Pyelonephritis_Infection) <- "SCT"
# check the specific marker by the MAST
#AllCategories_Clustermarkers <- FindAllMarkers(Pyelonephritis_Infection,
#            min.pct = 0.25,test.use= "MAST") # only markers that are in 25% of cells

# save all biomarkers as a dataframe
#write.csv(AllCategories_Clustermarkers, "AllCategories_Clustermarkers_MAST.csv")

# here we draw the specific biomarkers for the infected niches
AllCategories_Clustermarkers <- read.csv("AllCategories_Clustermarkers_MAST.csv", header = T)
AllCategories_Clustermarkers %>% head()
 
Top_InfectedrmarkersComparingAll<-AllCategories_Clustermarkers %>%
                group_by(cluster) %>% # group by cluster (cell type)
                slice_max(avg_log2FC, n = 10) %>% filter(cluster %in% Infectedniches)

Top25_InfectedrmarkersComparingAll<-AllCategories_Clustermarkers %>%
                group_by(cluster) %>% # group by cluster (cell type)
                slice_max(avg_log2FC, n = 25) %>% filter(cluster %in% Infectedniches)


Top_InfectedrmarkersComparingAllGene<- Top_InfectedrmarkersComparingAll %>% pull(gene) %>% unique()

Top25_InfectedrmarkersComparingAllGene<- Top25_InfectedrmarkersComparingAll %>% pull(gene) %>% unique()

  # generate the top 10 genes for each biomarker
# generate the heatmap
Top_InfectedrmarkersComparingAllGenesheatmap<- DoHeatmap(Pyelonephritis_Infection, features = Top_InfectedrmarkersComparingAllGene, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom") 
?DoHeatmap
ggsave("Top_InfectedrmarkersComparingAllGenesheatmap.pdf", plot =Top_InfectedrmarkersComparingAllGenesheatmap, height = 10, width = 9 )

Top_InfectedrmarkersComparingAllGeneDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = Top_InfectedrmarkersComparingAllGene) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))+RotatedAxis() 
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("Top_InfectedrmarkersComparingAllGeneDotPlot.pdf", plot =Top_InfectedrmarkersComparingAllGeneDotPlot, height = 9, width = 9 )


Top25_InfectedrmarkersComparingAllGeneDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = Top25_InfectedrmarkersComparingAllGene) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))+RotatedAxis() 
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 
ggsave("Top25_InfectedrmarkersComparingAllGeneDotPlot.pdf", plot =Top25_InfectedrmarkersComparingAllGeneDotPlot, height = 18, width = 9 )


```


```{r}
setwd(BioMarkerDir)

# generate the top 30 genes for each biomarker
top_Infectedrmarkers <- as.data.frame(AllCategories_Clustermarkers %>%
                group_by(cluster) %>% # group by cluster (cell type)
                slice_max(avg_log2FC, n = 30))
# generate the featureplot for each biomarker
top_Infectedrmarkers

top_InfectedrmarkersGenes<- top_Infectedrmarkers %>% pull(gene) %>% unique()

for (t in top_InfectedrmarkersGenes){
  
  # generate the figures for the top 10 biomarkers for each section
  RegionalSpecificSpatilaFeatureVariouCategoriesCut5 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
     # save into a file
    ggsave(filename = paste0(BioMarkerDir,t, "_SpatialFeatureplot_Cut5.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut5, height = 8, width = 15 )
    
    RegionalSpecificSpatilaFeatureVariouCategoriesCut8 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,8))
     # save into a file
    ggsave(filename = paste0(BioMarkerDir,t, "_SpatialFeatureplot_Cut8.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut8, height = 8, width = 15 )
    
     RegionalSpecificSpatilaFeatureVariouCategoriesCut10 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,10))
     # save into a file
    ggsave(filename = paste0(BioMarkerDir,t, "_SpatialFeatureplot_Cut10.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut10, height = 8, width = 15 )
    
      RegionalSpecificSpatilaFeatureVariouCategoriesCut20 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,20))
     # save into a file
    ggsave(filename = paste0(BioMarkerDir,t, "_SpatialFeatureplot_Cut20.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut20, height = 8, width = 15 )

 
}
```

# generate the dotplot and heatmap
```{r}
Idents(Pyelonephritis_Infection)
levels(Idents(Pyelonephritis_Infection))

# only for the top 5
# generate the top 5 genes for each biomarker
top5_Infectedrmarkers_genes <- as.data.frame(InfectedCategories_Clustermarkers %>%
                group_by(cluster) %>% # group by cluster (cell type)
                slice_max(avg_log2FC, n = 5)) %>% pull(gene) %>% unique()
# generate the heatmap
top_InfectedrmarkersGenesheatmap<- DoHeatmap(Pyelonephritis_Infection, features = top5_Infectedrmarkers_genes, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("top_InfectedrmarkersGenesheatmap.pdf", plot =top_InfectedrmarkersGenesheatmap, height = 10, width = 9 )

top_InfectedrmarkersGeneDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = top5_Infectedrmarkers_genes) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("top_InfectedrmarkersGeneDotPlot.pdf", plot =top_InfectedrmarkersGeneDotPlot, height = 8, width = 6 )

```

# Generate the dotplot for different CDK biomarkers
```{r}
# For Glomerular: 
GlomerularMarker <- paste0(BioMarkerDir,"/Glomerular/")
dir.create(GlomerularMarker)
setwd(GlomerularMarker)
GlomerularCKD_markers<- c("Alb", "Ddn", "Nphs1", "Trf", "Col4a1","Col4a2","Col4a3", "Col4a4","Col4a5", "Fn1", "Pdpn", "Synpo", "Cp", "Ptgds", "Ighm", "Des", "Smad1", "Actn1", "Adam10", "Chi3l1", "Podxl", "Vegfa", "Ptpro", "Myb", "Nphs2", "Eno3", "B2m")

# generate the heatmap
GlomerularCKD_markersheatmap<- DoHeatmap(Pyelonephritis_Infection, features = GlomerularCKD_markers, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("GlomerularCKD_markersGenesheatmap.pdf", plot =GlomerularCKD_markersheatmap, height = 10, width = 10 )

GlomerularCKD_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = GlomerularCKD_markers) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("GlomerularCKD_markersDotPlot.pdf", plot =GlomerularCKD_markersDotPlot, height = 8, width = 10 )

# make a directory for the spatial features
GlomerularMarkerSpFeature <- paste0(GlomerularMarker,"/SpatialFeature/")
dir.create(GlomerularMarkerSpFeature)
setwd(GlomerularMarkerSpFeature)

for (t in GlomerularCKD_markers){
  
  # generate the figures for the top 10 biomarkers for each section
  RegionalSpecificSpatilaFeatureVariouCategoriesCut5 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
     # save into a file
    ggsave(filename = paste0(GlomerularMarkerSpFeature,t, "_SpatialFeatureplot_Cut5.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut5, height = 8, width = 15 )
    
    RegionalSpecificSpatilaFeatureVariouCategoriesCut8 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,8))
     # save into a file
    ggsave(filename = paste0(GlomerularMarkerSpFeature,t, "_SpatialFeatureplot_Cut8.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut8, height = 8, width = 15 )
    
     RegionalSpecificSpatilaFeatureVariouCategoriesCut10 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,10))
     # save into a file
    ggsave(filename = paste0(GlomerularMarkerSpFeature,t, "_SpatialFeatureplot_Cut10.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut10, height = 8, width = 15 )
    
      RegionalSpecificSpatilaFeatureVariouCategoriesCut20 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,20))
     # save into a file
    ggsave(filename = paste0(GlomerularMarkerSpFeature,t, "_SpatialFeatureplot_Cut20.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut20, height = 8, width = 15 )

 
}

TubulointerstitialCKD_markers<- c("Cst3", "Havcr1", "Lcn2", "Oga", "Fabp3", "Ccn2", "Myb", "Umod", "Il18", "Lgals3", "Vnn1", "Nes", "Ambp", "Timp1", "Fabp1")
####
#### For Tubulointerstitial: 
TubulointerstitialMarker <- paste0(BioMarkerDir,"Tubulointerstitial/")
dir.create(TubulointerstitialMarker)
setwd(TubulointerstitialMarker)

# generate the heatmap
TubulointerstitialCKD_markersheatmap<- DoHeatmap(Pyelonephritis_Infection, features = TubulointerstitialCKD_markers, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("TubulointerstitialCKD_markersGenesheatmap.pdf", plot =TubulointerstitialCKD_markersheatmap, height = 10, width = 10 )

TubulointerstitialCKD_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = TubulointerstitialCKD_markers) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("TubulointerstitialCKD_markersDotPlot.pdf", plot =TubulointerstitialCKD_markersDotPlot, height = 8, width = 10 )

# make a directory for the spatial features
TubulointerstitialMarkerSpFeature <- paste0(TubulointerstitialMarker,"SpatialFeature/")
dir.create(TubulointerstitialMarkerSpFeature)
setwd(TubulointerstitialMarkerSpFeature)

for (t in TubulointerstitialCKD_markers){
  
  # generate the figures for the top 10 biomarkers for each section
  RegionalSpecificSpatilaFeatureVariouCategoriesCut5 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
     # save into a file
    ggsave(filename = paste0(TubulointerstitialMarkerSpFeature,t, "_SpatialFeatureplot_Cut5.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut5, height = 8, width = 15 )
    
    RegionalSpecificSpatilaFeatureVariouCategoriesCut8 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,8))
     # save into a file
    ggsave(filename = paste0(TubulointerstitialMarkerSpFeature,t, "_SpatialFeatureplot_Cut8.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut8, height = 8, width = 15 )
    
     RegionalSpecificSpatilaFeatureVariouCategoriesCut10 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,10))
     # save into a file
    ggsave(filename = paste0(TubulointerstitialMarkerSpFeature,t, "_SpatialFeatureplot_Cut10.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut10, height = 8, width = 15 )
    
      RegionalSpecificSpatilaFeatureVariouCategoriesCut20 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,20))
     # save into a file
    ggsave(filename = paste0(TubulointerstitialMarkerSpFeature,t, "_SpatialFeatureplot_Cut20.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut20, height = 8, width = 15 )
 
}


# Ahsg	Ccl2	Il18	Cd40lg	Chit1	
####
#### For Endothelial dysfunction

EndothelialDysCKD_markers <- c("Ahsg", "Ccl2", "Cxcl15", "Cd40lg", "Chit1")
EndothelialDysMarker <- paste0(BioMarkerDir,"EndothelialDys/")
dir.create(EndothelialDysMarker)
setwd(EndothelialDysMarker)

# generate the heatmap
EndothelialDysCKD_markersheatmap<- DoHeatmap(Pyelonephritis_Infection, features = EndothelialDysCKD_markers, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("EndothelialDysCKD_markersGenesheatmap.pdf", plot =EndothelialDysCKD_markersheatmap, height = 10, width = 10 )

EndothelialDysCKD_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = EndothelialDysCKD_markers) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("EndothelialDysCKD_markersDotPlot.pdf", plot =EndothelialDysCKD_markersDotPlot, height = 8, width = 10 )

# make a directory for the spatial features
EndothelialDysMarkerSpFeature <- paste0(EndothelialDysMarker,"SpatialFeature/")
dir.create(EndothelialDysMarkerSpFeature)
setwd(EndothelialDysMarkerSpFeature)

for (t in EndothelialDysCKD_markers){
  
  # generate the figures for the top 10 biomarkers for each section
  RegionalSpecificSpatilaFeatureVariouCategoriesCut5 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
     # save into a file
    ggsave(filename = paste0(EndothelialDysMarkerSpFeature,t, "_SpatialFeatureplot_Cut5.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut5, height = 8, width = 15 )
    
    RegionalSpecificSpatilaFeatureVariouCategoriesCut8 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,8))
     # save into a file
    ggsave(filename = paste0(EndothelialDysMarkerSpFeature,t, "_SpatialFeatureplot_Cut8.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut8, height = 8, width = 15 )
    
     RegionalSpecificSpatilaFeatureVariouCategoriesCut10 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,10))
     # save into a file
    ggsave(filename = paste0(EndothelialDysMarkerSpFeature,t, "_SpatialFeatureplot_Cut10.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut10, height = 8, width = 15 )
    
      RegionalSpecificSpatilaFeatureVariouCategoriesCut20 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,20))
     # save into a file
    ggsave(filename = paste0(EndothelialDysMarkerSpFeature,t, "_SpatialFeatureplot_Cut20.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut20, height = 8, width = 15 )
 
}


# Inflammation

#####
#### For Inflammation dysfunction

InflammationCKD_markers <- c("Crp", "Ptx3","Tnfrsf1a", "Tnfrsf1b", "Il18", "Il6", "Il10","Gdf15","Ccl6")

# Final used
FinalInflammationCKD_markers <- c("Tnfrsf1a", "Tnfrsf1b", "Il18", "Il6", "Ccl6")

InflammationMarker <- paste0(BioMarkerDir,"Inflammation/")
dir.create(InflammationMarker)
setwd(InflammationMarker)

# generate the heatmap
InflammationCKD_markersheatmap<- DoHeatmap(Pyelonephritis_Infection, features = InflammationCKD_markers, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("InflammationCKD_markersGenesheatmap.pdf", plot =InflammationCKD_markersheatmap, height = 10, width = 10 )

InflammationCKD_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = InflammationCKD_markers) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("InflammationCKD_markersDotPlot.pdf", plot =InflammationCKD_markersDotPlot, height = 8, width = 10 )

FinalInflammationCKD_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = FinalInflammationCKD_markers) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("FinalLimitedInflammationCKD_markersDotPlot.pdf", plot =FinalInflammationCKD_markersDotPlot, height = 4, width = 9 )


# make a directory for the spatial features
InflammationMarkerSpFeature <- paste0(InflammationMarker,"SpatialFeature/")
dir.create(InflammationMarkerSpFeature)
setwd(InflammationMarkerSpFeature)

for (t in InflammationCKD_markers){
  
  # generate the figures for the top 10 biomarkers for each section
  RegionalSpecificSpatilaFeatureVariouCategoriesCut5 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
     # save into a file
    ggsave(filename = paste0(InflammationMarkerSpFeature,t, "_SpatialFeatureplot_Cut5.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut5, height = 8, width = 15 )
    
    RegionalSpecificSpatilaFeatureVariouCategoriesCut8 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,8))
     # save into a file
    ggsave(filename = paste0(InflammationMarkerSpFeature,t, "_SpatialFeatureplot_Cut8.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut8, height = 8, width = 15 )
    
     RegionalSpecificSpatilaFeatureVariouCategoriesCut10 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,10))
     # save into a file
    ggsave(filename = paste0(InflammationMarkerSpFeature,t, "_SpatialFeatureplot_Cut10.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut10, height = 8, width = 15 )
    
      RegionalSpecificSpatilaFeatureVariouCategoriesCut20 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,20))
     # save into a file
    ggsave(filename = paste0(InflammationMarkerSpFeature,t, "_SpatialFeatureplot_Cut20.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut20, height = 8, width = 15 )
 
}




#### For Fibrosis dysfunction

FibrosisCKD_markers <- c("Tgfb1", "Cst3", "Tgfa", "Egf", "Il1rl1",  "Postn", "Pth", "Inhba", "Acp5", "Fgf23", "Kl", "Mmp9", "Mmp2", "Mmp7", "Dkk3")

FibrosisMarker <- paste0(BioMarkerDir,"Fibrosis/")
dir.create(FibrosisMarker)
setwd(FibrosisMarker)

# generate the heatmap
FibrosisCKD_markersheatmap<- DoHeatmap(Pyelonephritis_Infection, features = FibrosisCKD_markers, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("FibrosisCKD_markersGenesheatmap.pdf", plot =FibrosisCKD_markersheatmap, height = 10, width = 10 )

FibrosisCKD_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = FibrosisCKD_markers) +coord_flip() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("FibrosisCKD_markersDotPlot.pdf", plot =FibrosisCKD_markersDotPlot, height = 8, width = 10 )

# make a directory for the spatial features
FibrosisMarkerSpFeature <- paste0(FibrosisMarker,"SpatialFeature/")
dir.create(FibrosisMarkerSpFeature)
setwd(FibrosisMarkerSpFeature)

for (t in FibrosisCKD_markers){
  
  # generate the figures for the top 10 biomarkers for each section
  RegionalSpecificSpatilaFeatureVariouCategoriesCut5 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
     # save into a file
    ggsave(filename = paste0(FibrosisMarkerSpFeature,t, "_SpatialFeatureplot_Cut5.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut5, height = 8, width = 15 )
    
    RegionalSpecificSpatilaFeatureVariouCategoriesCut8 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,8))
     # save into a file
    ggsave(filename = paste0(FibrosisMarkerSpFeature,t, "_SpatialFeatureplot_Cut8.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut8, height = 8, width = 15 )
    
     RegionalSpecificSpatilaFeatureVariouCategoriesCut10 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,10))
     # save into a file
    ggsave(filename = paste0(FibrosisMarkerSpFeature,t, "_SpatialFeatureplot_Cut10.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut10, height = 8, width = 15 )
    
      RegionalSpecificSpatilaFeatureVariouCategoriesCut20 <- 
          SpatialFeaturePlot(Pyelonephritis_Infection,
                             features = t, 
                             images = selectedslices) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(0,20))
     # save into a file
    ggsave(filename = paste0(FibrosisMarkerSpFeature,t, "_SpatialFeatureplot_Cut20.pdf"), 
       plot =RegionalSpecificSpatilaFeatureVariouCategoriesCut20, height = 8, width = 15 )
 
}

# for AKI
AKI<- c("S100a8", "S100a9", "Lcn2", "Il18",  "Havcr1", "Igfbp7", "Fabp1", "Cst3","Umod") 
AKIMarker <- paste0(BioMarkerDir,"AKI/")
dir.create(AKIMarker)
setwd(AKIMarker)
# generate the heatmap
AKIMarker_markersheatmap<- DoHeatmap(Pyelonephritis_Infection, features = AKI, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("AKIMarker_markersGenesheatmap.pdf", plot =AKIMarker_markersheatmap, height = 10, width = 10 )

AKIMarker_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = rev(AKI)) +coord_flip() +RotatedAxis() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("AKIMarker_markersDotPlot.pdf", plot =AKIMarker_markersDotPlot, height = 5, width = 8 )

# for final figure, including the tubular function, tubular injury, Inflammation, Fibrosis

FinalAKICKDMarkers<- c("Fabp1", "Cst3","Umod","Igfbp7","Havcr1","Lcn2","Il18", "S100a8","S100a9", "Tnfrsf1a","Tnfrsf1b", "Il6", "Ccl6","Egf","Egfr","Kl","Tgfb1","Acp5","Mmp2","Mmp7","Mmp9")

# generate the heatmap
FinalAKICKD_markersheatmap<- DoHeatmap(Pyelonephritis_Infection, features = FinalAKICKDMarkers, size = 2.5)  + scale_fill_gradientn(colors = c("lightblue", "white", "FireBrick")) +  theme(legend.position="bottom")
?DoHeatmap
ggsave("FinalAKICKD_markersGenesheatmap.pdf", plot =FinalAKICKD_markersheatmap, height = 10, width = 10 )

FinalAKICKD_markersDotPlot<- 
   DotPlot(Pyelonephritis_Infection, features = rev(FinalAKICKDMarkers)) +coord_flip() +RotatedAxis() +
  theme(legend.text = element_text(size = 12), legend.title = element_text(size = 8))
  #DotPlot(Pyelonephritis_Infection, features = top_InfectedrmarkersGenes)+RotatedAxis() 

ggsave("FinalAKICKD_markersDotPlot.pdf", plot =FinalAKICKD_markersDotPlot, height = 6, width = 8 )

```


# generate the Dotplot for the functional enrichment of each classification:
```{r}
# change the folder back
setwd(file.path(DEGDir))
InfectedCategories_Clustermarkers %>% head()
SectionClusters <- levels(Idents(Pyelonephritis_Infection))
SectionClusters

# create a gene list for differnt regions
SectionDEGlistWithInfection <- list()
SectionClusters

# generate the DEGs for the different regions :"Cortex","Medulla","Pelvis","Vascular","Processing Infection"
for (m in SectionClusters){
  # put he DEG into the DEGlist for comparision
  # we only consider the Cortex, Medulla, Pelvis 
  if( m == "N12" | m == "N13" | m == "N14" | m== "N15" | m== "N16"| m== "N17"){
     SectionDEGlistWithInfection[[m]] <- InfectedCategories_Clustermarkers %>% filter(cluster %in% m & avg_log2FC >= 0.25) %>% pull(gene)
  }
 }

# Comparing the Cluster 
SectionDEGlistGOPlotWithInfection <- compareCluster(geneClusters = SectionDEGlistWithInfection, fun = "enrichGO", OrgDb = "org.Mm.eg.db",ont="BP",keyType = 'SYMBOL')

# simplify the GO enrichment
SectionDEGlistGOPlotWithInfectionSimplifed <- clusterProfiler::simplify(SectionDEGlistGOPlotWithInfection)

SectionDEGlistGOPlotWithInfectionSimplifed
SectionDEGlistGOPlotWithInfectionDotPlot<- dotplot(SectionDEGlistGOPlotWithInfection,font.size = 12,label_format = 80,showCategory = 6) +RotatedAxis()
?dotplot
#low = "LightBlue", midpoint=0, mid="white", high = "FireBrick",
ggsave(paste0("InfectedSectionDEGlistGOPlotDotPlot_withinfection.pdf"), plot = SectionDEGlistGOPlotWithInfectionDotPlot, height = 10, width = 10 )
write.csv(SectionDEGlistGOPlotWithInfectionSimplifed, "InfectedSectionDEGlistWithInfectionGOPlotSimplifed.csv")

# considering the GO enrichment took a long time, we read the saved files

SectionDEGlistGOPlotWithInfectionSimplifed <- read.csv("InfectedSectionDEGlistWithInfectionGOPlotSimplifed.csv", header = T)

```


# generate the KEGG pathway oever-representation analysis

```{r}

setwd(KEGGDir)
library(org.Mm.eg.db)
KEGGpathways <-list()
KEGGpathwaysForView<-list()
## put the list of UNIPROT into the list
# we generate this for the kegg pathview
SectionDEGTransferUnirotID <-list()
SectionClusters
#SectionDEGlistWithInfection
for (m in SectionClusters){
  # put he DEG into the DEGlist for comparision
  # we only consider the infected niches 
  if( m == "N12" | m == "N13" | m == "N14" | m== "N15" | m== "N16"| m== "N17"){
    
     #m= "Processing Infection"
     SectionDEGlistWithInfection[[m]] <- InfectedCategories_Clustermarkers %>% filter(cluster %in% m & avg_log2FC >= 0.25) %>% pull(gene)
     
     # transfer biolgical ID to KEGG ID
     TransferIDs<- bitr(SectionDEGlistWithInfection[[m]] , fromType = "SYMBOL", toType = c("UNIPROT","ENSEMBL"), OrgDb=org.Mm.eg.db)
     TransferIDs %>% head()
      
     SectionDEGTransferUnirotID[[m]] <- TransferIDs$UNIPROT
     
     egID = bitr(SectionDEGlistWithInfection[[m]] , fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Mm.eg.db")

     KEGGpathwaysForView[[m]] <-enrichKEGG(gene = egID$ENTREZID,
                                           organism = 'mmu',
                                           pvalueCutoff = 0.05)
     KEGGpathways[[m]] <- enrichKEGG(gene =  TransferIDs$UNIPROT, organism = 'mmu', pvalueCutoff = 1,keyType = 'uniprot')
  
     write.csv(KEGGpathways[[m]], file = paste0(KEGGDir, m, "_KeggEnrichpath.csv"))
  }
 }
```


## compare the KEGG pathways among Cortex, medulla, Pelvis and Processing Infection
```{r}

#SectionDEGTransferUnirotID
length(SectionDEGTransferUnirotID)
 # compare four up-regulated gene list with GO enrichment, we did not set up the p value 
SectionDEGTransferKEGG <- compareCluster(geneCluster = SectionDEGTransferUnirotID, fun = enrichKEGG, organism="mmu", keyType = 'uniprot',pvalueCutoff=1)

# setting into readable ID
SectionDEGTransferKEGGReadable <- setReadable(SectionDEGTransferKEGG, OrgDb = "org.Mm.eg.db", keyType="UNIPROT")

# remove the string - Mus musculus (house mouse)
SectionDEGTransferKEGGReadable@compareClusterResult$Description <-

gsub(pattern = " - Mus musculus (house mouse)", replacement = "", SectionDEGTransferKEGGReadable@compareClusterResult$Description, fixed = T) 


#%>% mutate(Description = str_remove(Description, "- Mus musculus (house mouse)")) %>% head()
# 
# SectionDEGTransferKEGGReadable$Description <- gsub(pattern = " - Mus musculus (house mouse)", replacement = "", SectionDEGTransferKEGGReadable$Description, fixed = T)

# Show all the top15 pathways
Top15Comparion<-  
  dotplot(SectionDEGTransferKEGGReadable,font.size = 12,label_format = 80,showCategory = 15) 

ggsave(paste0(KEGGDir,"Top15Comparion_Sections_KEGG_comp.pdf"), plot = Top15Comparion, width = 15, height = 15)
# save into a enriched pathway file
write.csv(SectionDEGTransferKEGGReadable, file = paste0(KEGGDir,"AllSections_KeggEnrichpath.csv") )


# save the top ID

# show the candidate pathway in the infected regions
selectedPathway<- c("Carbon metabolism","Glyoxylate and dicarboxylate metabolism","Galactose metabolism","Cell adhesion molecules","ECM-receptor interaction","Focal adhesion","Tight junction","Necroptosis","Complement and coagulation cascades","Apoptosis","Endocytosis","Lysosome","Hippo signaling pathway","Phagosome","Neutrophil extracellular trap formation","Chemokine signaling pathway","Hematopoietic cell lineage","Cytokine-cytokine receptor interaction","NOD-like receptor signaling pathway","Osteoclast differentiation","NF-kappa B signaling pathway","Th17 cell differentiation","Antigen processing and presentation","Inflammatory bowel disease","Toll-like receptor signaling pathway","Human cytomegalovirus infection","TNF signaling pathway","C-type lectin receptor signaling pathway","Th1 and Th2 cell differentiation","IL-17 signaling pathway","Fc gamma R-mediated phagocytosis","Cytosolic DNA-sensing pathway","Natural killer cell mediated cytotoxicity","B cell receptor signaling pathway","JAK-STAT signaling pathway","Intestinal immune network for IgA production","Leukocyte transendothelial migration","T cell receptor signaling pathway","Rap1 signaling pathway","Bacterial invasion of epithelial cells","Fc epsilon RI signaling pathway","p53 signaling pathway")

# select these interesting pathway
SelectedPathwayForfigure<-SectionDEGTransferKEGGReadable %>% as.data.frame() %>% filter(Description %in% selectedPathway) %>% arrange(p.adjust)

# setting up the minium p adjust value for size visulization 
maximum<- max(c(-log10(SelectedPathwayForfigure$p.adjust)))

## set the levels of Description and reorder them
SelectedPathwayForfigure$Description<- ordered(SelectedPathwayForfigure$Description, levels=selectedPathway)
#reorder(variable, value)
p<-ggplot(SelectedPathwayForfigure, aes(x=SelectedPathwayForfigure$Cluster, y=SelectedPathwayForfigure$Description))

InterestingComparionsDotplot<- p+ geom_point(aes(size=-log10(p.adjust)),
                                             colour= ifelse(SelectedPathwayForfigure$p.adjust <=0.05,"red",
                                                            ifelse(SelectedPathwayForfigure$p.adjust==1,"white","gray")), 
                                             fill= ifelse(SelectedPathwayForfigure$p.adjust<=0.05,"red",
                                                          ifelse(SelectedPathwayForfigure$p.adjust==1,"white","gray"))) +
  labs(title = "KEGG Pathway Enrichment", x = "Spatial region", y = "KEGG pathway enrichment") +
  theme_bw()+
  theme(plot.background = element_blank(),
        text = element_text(size=10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1)
  ) + scale_size_area(limits=c(0,maximum),breaks=c(0,4,6,8,12))


InterestingComparions <- dotplot(SectionDEGTransferKEGGReadable,font.size = 12,label_format = 80,showCategory = selectedPathway)  +RotatedAxis()

ggsave(paste0(KEGGDir, "InterestingPathwayComparion_Sections_KEGG_comp.pdf"), plot = InterestingComparions, width = 8, height = 10)
ggsave(paste0(KEGGDir, "InterestingPathwayComparion_Sections_KEGG_comp_format.pdf"), plot = InterestingComparionsDotplot, width = 8, height = 10)
```


# Using PROGENy model from DecoupleR
```{r}
library(decoupleR)
# set the identity as orginal
Idents(Pyelonephritis_Infection) <- Pyelonephritis_Infection @meta.data$Annotation
#devtools::install_github("saezlab/progeny")
library(decoupleR)
setwd(ProgenyDir)
net <- decoupleR::get_progeny(organism = 'mouse', top = 500)

Pyelonephritis_Infection@assays$SCT@data %>% head()
# Activity inference with Multivariate Linear Model (MLM)
# Extract the normalized log-transformed counts
mat <- as.matrix(Pyelonephritis_Infection@assays$SCT@data)

# Run mlm
acts <- run_mlm(mat=mat, net=net, .source='source', .target='target',
                .mor='weight', minsize = 5)

acts 
# Visualization:
# Extract mlm and store it in pathwaysmlm in data
Pyelonephritis_Infection[['pathwaysmlm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)


# Change assay
DefaultAssay(object = Pyelonephritis_Infection) <- "pathwaysmlm"

# Scale the data
Pyelonephritis_Infection <- ScaleData(Pyelonephritis_Infection)
Pyelonephritis_Infection@assays$pathwaysmlm@data <- Pyelonephritis_Infection@assays$pathwaysmlm@scale.data

# run the source of pathways
ProgenyFeatures <- levels(factor(acts$source))
ProgenyFeatures
for (i in ProgenyFeatures ){
  p2 <-SpatialFeaturePlot(Pyelonephritis_Infection, features = i) +
  ggtitle(paste0(i," activity")) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))
  
  ggsave(paste0(i, "_Pyelonephritis_Infection_cluster_featurePlot_DecoupleR.pdf"), plot = p2, width = 20, height = 4)
}

#(c("royalblue", "white","lightcoral"))
Pyelonephritis_Infection@meta.data$orig.ident
SpatialFeaturePlot(Pyelonephritis_Infection, features = "Hypoxia", min.cutoff = -4, max.cutoff = 6) & 
   ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))& 
  ggtitle(paste0("Hypoxia"," activity"))

# SpatialFeaturePlot(Pyelonephritis_Infection, features = "NFkB", min.cutoff = -4, max.cutoff = 6, split.by = "orig.ident") & 
#     ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5)) & 
#   ggtitle(paste0("NFkB"," activity"))
# 
# 
# SpatialFeaturePlot(Pyelonephritis_Infection, features = "TNFa", min.cutoff = -4, max.cutoff = 6) & 
#    ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))& 
#   ggtitle(paste0("TNFa"," activity"))
# 
# p1 <- DimPlot(Pyelonephritis_Infection, reduction = "umap", label = TRUE, pt.size = 0.5) + 
#   NoLegend() + ggtitle('Cell types')
# p2 <- (SpatialFeaturePlot(Pyelonephritis_Infection, features = c("Trail") & 
#      ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5))&
#   ggtitle('Trail activity')



# Expore the mean activiey per group across pathways:
# Extract activities from object as a long dataframe
df <- t(as.matrix(Pyelonephritis_Infection@assays$pathwaysmlm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(Pyelonephritis_Infection)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Transform to wide matrix
top_acts_mat <- df %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
#my_color = colorRampPalette(c("dodgerblue2", "white","firebrick1"))(palette_length)
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)


my_breaks <- c(seq(-2, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 2, length.out=floor(palette_length/2)))

# Plot
Pyelonephritis_InfectionHeatmap<- pheatmap(t(top_acts_mat), border_color = NA, cluster_cols = F, color=my_color, breaks = my_breaks, angle_col = 45) 
ggsave("Pyelonephritis_InfectioncellsHeatmap_Progeny_Decouple.pdf", plot = Pyelonephritis_InfectionHeatmap, width = 6, height = 4)

```

# Transcription factor activity inference from Immune scRNA-seq
```{r}
# CollecTRI network
# set up the directory
library(decoupleR)
setwd(TFDir)
netTF <-  readRDS("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Datasets/get_collectri_human_mouse20230716/get_collectri_mouse_process_table.rds")

mat <- as.matrix(Pyelonephritis_Infection@assays$SCT@data)
# Run ulm
#acts <- run_ulm(mat=mat, net=netTF, .source='source', .target='target',
 #               .mor='mor', minsize = 5)

#saveRDS(acts, file="Pyelonephritis_InfectionNiches_Transcriptfactor.RDS")
# this took a long time,
#acts <- readRDS(file = "Pyelonephritis_Infection_Transcriptfactor.RDS")

# Activity inference with Univariate Linear Model (ULM)
# Extract ulm and store it in tfsulm in pbmc
Pyelonephritis_Infection[['tfsulm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = Pyelonephritis_Infection) <- "tfsulm"

# Scale the data
Pyelonephritis_Infection <- ScaleData(Pyelonephritis_Infection)
Pyelonephritis_Infection@assays$tfsulm@data <- Pyelonephritis_Infection@assays$tfsulm@scale.data

# Explore the top 20 more varaible TFs
n_tfs <- 20

t(as.matrix(Pyelonephritis_Infection@assays$tfsulm@data))
Idents(Pyelonephritis_Infection)
Idents(Pyelonephritis_Infection)

#t(as.matrix(Pyelonephritis_Infection@assays$tfsulm@data)) 
# Extract activities from object as a long dataframe
df <- t(as.matrix(Pyelonephritis_Infection@assays$tfsulm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(Pyelonephritis_Infection)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))


# # get the top 10 biomarkers 
# TopTF<- df %>% arrange(mean,cluster)%>% group_by(cluster) %>% slice_max(mean,n=10)
# nrow(Top)
# Get top tfs with more variable means across clusters
tfs <- df %>%
  group_by(source) %>%
  summarise(std = sd(mean)) %>%
  arrange(-abs(std)) %>%
  head(n_tfs) %>%
  pull(source)



# Transform to wide matrix for topTF
top_acts_mat <- df %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

top_acts_mat
# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-2, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 2, length.out=floor(palette_length/2)))

# Plot
pheatmap(top_acts_mat, border_color = NA, color=my_color, breaks = my_breaks) 
# drop "Znf335" due to not exist in the expression
#tfs <- tfs[!tfs == "Znf335"]

#tfs[!tfs == "Znf335"]
# Subset long data frame to top tfs and transform to wide matrix
top_acts_mat <- df %>%
  filter(source %in% tfs) %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-3, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 3, length.out=floor(palette_length/2)))

# Plot
IMMTFheatmap<- pheatmap(t(top_acts_mat),border_color = NA, cluster_cols = F, 
                        color=my_color,fontsize=12, fontsize_row = 10, 
                        breaks = my_breaks, 
                        #main = "Predicted TF activities", 
                        angle_col = 45,
                        treeheight_col = 0, treeheight_row=0)


ggsave("Pyelonephritis_InfectedNichesHeatmap_Transcriptfactor_Progeny_Decouple.pdf", plot = IMMTFheatmap, width = 5, height = 5)

# draw the transcriptor factor activity and their expression
# generate the figure of transcript factor activity:
for (m in tfs){
  p2 <- (SpatialFeaturePlot(Pyelonephritis_Infection, features = m))  +
  ggtitle(paste0(m," activity")) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5)) 
  
  ggsave(paste0("TFActivity_",m,"_Pyelonephritis_Infection_FeaturePlot.pdf"), plot = p2, width = 16, height = 4)
}

p3 <- (SpatialFeaturePlot(Pyelonephritis_Infection, features = "Hbp1"))  +
  ggtitle(paste0("Hbp1"," activity")) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-2,5)) 
p3  
  ggsave(paste0("TFActivity_","Hbp1","_Pyelonephritis_Infection_FeaturePlot.pdf"), plot = p3, width = 16, height = 4)

  p4 <- (SpatialFeaturePlot(Pyelonephritis_Infection, features = "Nfkb1"))  +
  ggtitle(paste0("Nfkb1"," activity")) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-2,6)) 
p4  
  ggsave(paste0("TFActivity_","Nfkb1","_Pyelonephritis_Infection_FeaturePlot.pdf"), plot = p4, width = 16, height = 4)

Pyelonephritis_Infection
DefaultAssay(object = Pyelonephritis_Infection) <- "SCT"

for (n in tfs){
  p3 <- (SpatialFeaturePlot(Pyelonephritis_Infection, features = n) +
  ggtitle(paste0(m," activity")) &
          ggplot2::scale_fill_gradientn(colours = myPalette(4), limits=c(-1,5)))
  
  ggsave(paste0("TFExpression_",n,"_Pyelonephritis_Infection_FeaturePlot.pdf"), plot = p3, width = 16, height = 4)

}

# add their activity in Dotplot
Idents(Pyelonephritis_Infection)
# get the order of row name
GeneOrderByPheatmap <- rownames(t(top_acts_mat))[IMMTFheatmap$tree_row$order]
#IMMTFheatmap$tree_row$order

?DotPlot
TFDotPlot <- DotPlot(Pyelonephritis_Infection, features = rev(GeneOrderByPheatmap)) +
  RotatedAxis() + coord_flip()+ 
  #ggtitle("Predicted TF expression")+
   theme(legend.position = "right")
ggsave("Spatial_TFDotPlot.pdf", plot = TFDotPlot, width = 10, height = 6)

```



