---
title: "PipseqReclustering_V0803"
author: "Xin Wang"
date: "2023-08-10"
output: html_document
description: This script is to generate the clustering and annoation infection for Pipseq:
  1. UMAP of single cell landscape of mouse Pyelonephritis : including the quality control, integration of objects, harmony batch effect removal
  2. Annotation of Pyelonephritis : with known published biomarkers
  3. Subcluster of Proximal tubule cells : with candidate known PT status biomarkers
  4. Psudotime analysis of Proximal tubule cells
---

# Loading the packages
```{r}
# check the session information
sessionInfo()
library(sf)
library(dplyr)
library(Seurat)
library(patchwork)

#install.packages('devtools') #assuming it is not already installed

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
library(devtools)
#install_github('andreacirilloac/updateR')
# install.packages("sctransform")
# install.packages("ggthemes")
# loading the libarary
library(updateR)

library("sctransform")
library("ggthemes")
library(Seurat)

# BiocManager::install("limma")
# BiocManager::install("ComplexHeatmap")
# BiocManager::install("biomaRt")
# BiocManager::install("methrix")BiocManager::install("org.Hs.eg.db")
# devtools::install_github("saeyslab/nichenetr")
library(limma)
library(nichenetr)
library(tidyverse)
library(dplyr)

library(biomaRt)
library("data.table")


library(umap)
library(patchwork)
library(cowplot)
library(installr)

#install.packages("Matrix", repos = "http://cran.r-project.org")
### we used discre color palettes from ggsci

library("ggsci")
library("ggplot2")
library("gridExtra")


#options(buildtools.check = function(action) TRUE )
#devtools::install_url('https://cran.r-project.org/src/contrib/Archive/Matrix.utils/Matrix.utils_0.9.7.tar.gz')
#devtools::install_github('cole-trapnell-lab/monocle3')
#library(monocle3)

library(harmony)
library(Seurat)
#library(SeuratData)
library(tidyverse)

#library(BPCells)
library(ggplot2)
library("devtools")
library("AnnotationDbi")
library("org.Hs.eg.db")
#library(tibble)
#library(future)
library(here)
library(patchwork)
library(future)
library(monocle3)
#devtools::install_dev("remotes")
#remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
#remotes::install_github(repo ='chris-mcginnis-ucsf/DoubletFinder' )
#install.packages("DoubletFinder")
library(DoubletFinder)
library(future.callr)
library("ggsci")
library("ggplot2")
library("gridExtra")
```


## Step 2: Set up the workspace environment:
```{r}
setwd(
  "/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/"
)
Indir <-
  c(
    "/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/PreprocessedPipseeker/"
  )
Outdir <-
  c(
    "/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/"
  )
## check the files in the current directory
Datafiles <-
  list.files(
    path = paste0(Indir, "/"),
    recursive = F,
    full.names = F
  )

# setting up the seed 
set.seed(10000)
```

# Step 3. Prepare the input datasets and create objects
```{r}
# create an empty list to store the objects
scrna.list = list()

# create a vector of convenient sample names, such as "D0","D3","D7" and "D28"
for (i in Datafiles) {
  # initiate the seurate object
  scrna.list[[i]] = ReadMtx(
    mtx = paste0(Indir, i, "/matrix.mtx.gz"),
    features = paste0(Indir, i, "/features.tsv.gz"),
    cells = paste0(Indir, i, "/barcodes.tsv.gz")
  )
  # initiate the seurate object
  #scrna.list[[name2]]<- CreateSeuratObject(counts=rds, project= name2)
  scrna.list[[i]] = CreateSeuratObject(
    counts = scrna.list[[i]],
    Project = samples[i],
    min.cells = 3,
    min.features = 200
  )
  
  
  # rename the cell with name id
  scrna.list[[i]] <- RenameCells(scrna.list[[i]], add.cell.id = i)
  ## using the assign the name to the objects
  assign (i, scrna.list[[i]])
  ## create a group that show the name of samples
  scrna.list[[i]]$DataSet <-
    rep(i, length(scrna.list[[i]]$orig.ident))
  
  scrna.list[[i]]$percent.mt <-
    PercentageFeatureSet(scrna.list[[i]], pattern = "^mt-")
  scrna.list[[i]]$rDNA <-
    PercentageFeatureSet(scrna.list[[i]], pattern = "^Rp[sl][[:digit:]]")
  
  scrna.list[[i]] <-
    subset(
      scrna.list[[i]],
      subset = nFeature_RNA > 200  &
        nFeature_RNA < 8000 &
        nCount_RNA < 16000 & percent.mt < 20 & rDNA < 40
    )
  scrna.list[[i]] <- NormalizeData(scrna.list[[i]])
  scrna.list[[i]] <- ScaleData(scrna.list[[i]])
  scrna.list[[i]] <-
    FindVariableFeatures(scrna.list[[i]],
                         selection.method = "vst",
                         nfeatures = 2000)
  scrna.list[[i]] <- RunPCA(scrna.list[[i]])
  # ElbowPlot(seurat_obj)
  scrna.list[[i]] <- FindNeighbors(scrna.list[[i]], dims = 1:20)
  scrna.list[[i]] <- RunUMAP(scrna.list[[i]], dims = 1:20)
  
}

```

# merge the seurate objects into a single one
```{r}
## merge might not be a good way for the intergration, instead we used the integration by find the integration features
scrna <-
  merge(
    x = scrna.list[[1]],
    y = c(scrna.list[[2]], scrna.list[[3]], scrna.list[[4]]),
    project = "Pyelonephritis"
  )

## change the order of days 
levels(factor(scrna@meta.data$DataSet))

scrna@meta.data$DataSet <-
  ordered(factor(scrna@meta.data$DataSet),
          levels = c("0DPI", "3DPI", "7DPI", "28DPI"))

scrna@meta.data %>% tail()
# create a condition details with and without infection
scrna@meta.data <- scrna@meta.data %>%
  mutate(Condition = case_when(scrna@meta.data$DataSet == "0DPI" ~ "Noninfect",
                               TRUE ~ "Infect"))

# measure the mitochondrial percentage
scrna[["percent.mt"]] <-
  PercentageFeatureSet(scrna, pattern = "^mt-")
scrna[["rDNA"]] <-
  PercentageFeatureSet(scrna, pattern = "^Rp[sl][[:digit:]]")

#VlnPlot(scrna, features = c("percent.mt", "nCount_RNA", "nFeature_RNA","rDNA"))
 QCReport <-
   VlnPlot(
     scrna,
     features = c("rDNA", "percent.mt", "nCount_RNA", "nFeature_RNA") ,
     split.by = "DataSet",
     pt.size=0,
     ncol = 1
   )

ggsave(filename = "SampleQC_reports.pdf",plot = QCReport, height = 12, width = 12 )

pdf("PIPseq_Feature.pdf", height = 4, width = 6)
VlnPlot(scrna, features = c("nFeature_RNA"), group.by = "DataSet",pt.size=0 )
dev.off()
?VlnPlot
pdf("PIPseq_nCounts.pdf", height = 4, width = 6)
VlnPlot(scrna, features = c("nCount_RNA"), group.by = "DataSet",pt.size=0 )
dev.off()
?VlnPlot

pdf("PIPseq_mtPercent.pdf", height = 4, width = 6)
VlnPlot(scrna, features = c("percent.mt"), group.by = "DataSet",pt.size=0,y.max=80 )
dev.off()
?VlnPlot

### filtered the low quality cells that have less than 200 RNA and larger than 8000 and percentage mt high than 20%, rDNA higher than 40
scrna <-
  subset (scrna,
          subset = nFeature_RNA > 200  &
            nFeature_RNA < 8000 &
            nCount_RNA < 16000 & percent.mt < 20 & rDNA < 40)
```

# Calculate the Frequency of Cells in each Datasets
```{r}
SummarizedCells <- as.data.frame(table(scrna@meta.data$DataSet))
nrow(scrna@meta.data)
OrignalFrequencyCellTypes <-
  ggplot(data = SummarizedCells , aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(
    aes(label = Freq),
    position = position_dodge(width = 0.9),
    vjust = -0.25,
    size = 2
  ) +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
    ),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black")
  ) + xlab("Samples") + ylab("Number of cells ")

ggsave(
  "OrignalFrequencyCells.pdf",
  plot = OrignalFrequencyCellTypes,
  width = 8,
  height = 4
)
```

# Removing batch effects using Harmony
```{r}

# # enable parallel processing via future package
# plan("multiprocess", workers = 10) 
# options(future.globals.maxSize = 100000 * 1024^2) # for 100 Gb RAM
# plan()
#"mitoPercent","Rpl", "nCount_RNA", "nFeature_RNA"
# 
scrna <- scrna %>%
    NormalizeData() %>%
    FindVariableFeatures(selection.method = "vst") %>% 
    ScaleData() %>%
    SCTransform()

scrna <- RunPCA(scrna, verbose = TRUE, assay = "SCT", npcs = 30)
# ElbowPlot(rnaAggr, ndims = 50) # to determine number of dimensions for clustering
scrna <- RunHarmony(scrna, group.by.vars = c("DataSet"), assay.use="SCT",reduction.save = "harmony")
scrna <- RunUMAP(scrna, dims = 1:30, reduction = "harmony", assay = "SCT")
scrna <- FindNeighbors(scrna, dims = 1:30,  reduction = "harmony")
scrna <- FindClusters(scrna,  resolution = 0.1,algorithm = 1)

scrna
# scrna <- FindNeighbors(scrna, dims = 1:30,  reduction = "harmony", graph.name = "Test")
# scrna <- FindClusters(scrna,  resolution = 0.1,algorithm = 1, graph.name = "Test")

#rnaAggr@meta.data$doublet_viz <- ifelse(rnaAggr@meta.data$doublet_id == "Singlet",0,1)
#DefaultAssay(scrna) <-"SCT"
# draw the dimplot grouped by doublet_id
DimplotScrna <- DimPlot(scrna, reduction = "umap", label = T) 
DimplotScrna
DimplotScrnaSplit <- DimPlot(scrna, reduction = "umap",split.by = "DataSet", label = T) 
DimplotScrnaSplit

ggsave(filename = "DimplotScrnat.pdf", plot = DimplotScrna, width = 6, height = 5)
ggsave(filename = "DimplotScrnaSplit.pdf", plot = DimplotScrnaSplit, width = 20, height = 5)

saveRDS(scrna, file = "Pyelonephritis_singlecell_harmony_v0803.RDS")

scrna<- readRDS("Pyelonephritis_singlecell_harmony_v0803.RDS")

```


# We further reclustering the cluster 1 and 5 for a higher resolution
```{r}
# we further divide the meyeloblast into neutrophil, basophils and microphage
# the default cluster names were saved in the:
scrna@commands$FindClusters$graph.name

Idents(scrna) <- scrna@meta.data$seurat_clusters
# use the SCT cluster graph name do subcluster for 1
scrna<-FindSubCluster(scrna,cluster=c("1"),graph.name = "SCT_snn",subcluster.name = "Sub1", resolution = 0.05)

# set up the cluster details into Sub1
Idents(scrna) <- scrna@meta.data$Sub1

# scrna<-FindSubCluster(scrna,cluster=c("4"),graph.name = "SCT_snn",subcluster.name = "Sub4", resolution = 0.05)
# Idents(scrna) <- scrna@meta.data$Sub4

# use the SCT cluster graph name do subcluster for 5
scrna<-FindSubCluster(scrna,cluster=c("5"),graph.name = "SCT_snn",subcluster.name = "Sub5", resolution = 0.05)
Idents(scrna) <- scrna@meta.data$Sub5


# use the SCT cluster do subcluster for 

# generate the DimPlots
DimplotScrnaSub <- DimPlot(scrna, reduction = "umap", label = T) 
DimplotScrnaSplitSub <- DimPlot(scrna, reduction = "umap",split.by = "DataSet", label = T) 
DimplotScrnaSplitSub

ggsave(filename = "DimplotScrnatSub.pdf", plot = DimplotScrnaSub, width = 6, height = 5)
ggsave(filename = "DimplotScrnaSplitSub.pdf", plot = DimplotScrnaSplitSub, width = 16, height = 5)

```


# Generate the biomarker distribution
```{r}
DefaultAssay(scrna) <-"RNA"

# Cfh: Parietal epithelial cells
celltype.markers<- c("Lrp2","Slc34a1","Acsm1","Atp11a","Slc13a3","Notch2","Cubn","Vcam1","Havcr1","Fn1",
                     "Col1a1","Il6","Il1a","Il1b","Tnf","Mmp3","Mmp9","Slc5a1",
                     "Slc5a2", # proximal connvoluted tubule
                     "Cfh",
               "Ireb2","Alas2","C78334","Slc12a3","Pvalb","Wnk1","Atp6v1g3","Aqp6", "Atp6v0d2","Slc4a1",
                     "Slc26a7","Slc26a4","Aqp2","Hsd11b2","Scnn1g","Fst","Bst1","Slc14a2","Umod","Slc12a1",
                     "Cldn10","Krt5","Krt14","Upk1b","Nphs1","Nphs2","Pecam1","Cdh5","Igfbp3","Acta2",
                     "Ehd3","Nrp1","Col3a1","Lum","Fbn1","Prkg1","C1qa","C1qb","Aif1","Ptprc","Cd3e",
                     "Cd14","Lst1","Lyz2","Cebpb","Cxcr6","Cd247","Nkg7","Cd7","Ccl5","Cd79a","Cd79b",
                     "Ms4a1","Igkc","Ms4a7","Adap2","S100a8", "S100a9", "Ifitm1","Ly6g", "Lcn2","Cxcl1",
                     "Acod1","Slc22a8","Slc7a12", "Krt19","Gadd45b", "Ripk3", "Tnf", "Tgfb1","Cd9","Nrxn3","Nphs1","Nphs2","Plvap","Enox1","Postn","Mki67","Serpine2","Fhl2","Des",
                    
                     "Tagln","Gata3", # SMC
                     "Plat","Emcn","Kdr", # Glomerular endothelial cells
                    "Apoe","Cst3","Aldob", # damaged Promimal tubule
               "Clu","Tpm1","Spp1","Vim" # adaptive PT
               ) %>% unique()

Clustermakrers<- c("Lrp2","Slc34a1","Acsm1","Slc5a2","Slc5a1","Havcr1","Fn1","Col1a1","Notch2","Enox1","Klh13","Atp11a","Cfh","Ireb2","Alas2","C78334","Slc12a3","Pvalb","Wnk1","Atp6v1g3","Aqp6","Slc4a1","Rcan2","Kit","Slc26a4","Hmx2","Spink8","Slc26a7","Slc8a1", "Trpv5","Nphs1", "Calb1","Skap1", "Upk2a","Upk3a","Upk3b",
                   "Aqp2","Atp6v1b1","Hsd11b2","Scnn1g","Fst","Bst1","Slc14a2","Umod","Slc12a1","Cldn10","Krt5","Krt14","Upk1b", 
                   "Pecam1","Cdh5","Nrp1","Col3a1","Lum","Fbn1","C1qa","C1qb","Aif1","Fhl2","Lst1","Lyz2","Ptprc","Cebpb","Cxcr6","Cd247","Nkg7","Cd7","Ccl5","Cd79a","Cd79b","Ms4a1","Ly6g", "Lcn2","Slc22a8","Slc12a1", "Krt19","Mki67")

FinalUsedMarkers<- c("Lrp2","Slc34a1","Acsm1","Atp11a","Slc5a2","Atp6v1g3","Slc26a4","Hmx2","Spink8","Slc26a7","Aqp6","Slc4a1","Rcan2","Kit","Aqp2","Hsd11b2" ,"Scnn1g","Slc8a1","Slc12a3","Wnk1","Fst","Bst1","Slc14a2","Umod","Slc12a1","Nphs1","Nrp1","Col3a1","Fbn1", "Fhl2","Lst1","Lyz2","Ptprc","Atp6v1b1","S100a9", "Ifitm1")

ICAMarkers <- c("Slc26a7","Rcan2","Atp4a","Slc4a1","Aqp6","Kit")
ICBMarkers<- c("Slc26a4","Hmx2","Spink8")
PCMarkers <- c("Aqp2","Hsd11b2" ,"Scnn1g")

#FinalUssedMarkers2<- c("Lrp2","Slc34a1","Acsm1","Cubn","Havcr1","Slc12a3","Trpm6","Pvalb","Wnk1","Slc8a1","Calb1","Trpv5","Slc12a1","Cldn10","Atp6v1g3","Aqp6","Slc26a4","Aqp2","Hsd11b2" ,"Scnn1g","Fst","Bst1","Slc14a2","Krt5","Krt14","Upk1b", "Pecam1","Cdh5","Igfbp3","Col3a1","Lum","Fbn1","Ireb2","Alas2","C78334","Ptprc","Lyz2","Lst1","C1qa","C1qb","Cxcr6","Cd247", "Nkg7","Cd7","Ccl5","Cd79b","Ms4a1","Igkc","Il1b","Cd14")

FinalUssedMarkers2<- c("Lrp2","Slc34a1","Acsm1","Cubn","Havcr1","Slc12a3","Slc8a1","Calb1","Trpm6","Pvalb","Wnk1","Trpv5","Fst","Slc14a2","Slc12a1","Cldn10","Atp6v1g3","Aqp6","Slc26a4","Aqp2","Hsd11b2" ,"Scnn1g","Krt5","Krt14","Upk1b", "Col3a1","Lum","Fbn1","Pecam1","Igfbp3","Cdh5","Ptprc","Lyz2","Lst1","C1qa","C1qb","Il1b","Cd14","Cxcr6","Cd247", "Cd79b","Ms4a1","Igkc","Nkg7","Cd7","Ccl5","Ireb2","Alas2","C78334")

# macrophage 1 and macrophage 2:

# M1 can choose CD80, CD86, CD64, CD16 and CD32 as markers. 
Macrophage2TypeMarkers <- c("Ifngr1","Il17a","Cd80","Cd86","Fcgr1","Fcgr3","Fcgr2b","Il13","Cd163","Mrc1","Arg1", "Cd68")

DotPlot(scrna, features = Macrophage2TypeMarkers)
Macrophage1Markers <- c("Ifngr1","Il17a","Cd80","Cd86","Fcgr1","Fcgr3","Fcgr2b")


# M2 can choose CD163 and CD206 are major markers, Arg1 DECTIN
Macrophage2Markers <- c("Il13","Cd163","Mrc1","Arg1", "Cd68")

CFilePath<- paste0(Outdir,"Candidatemarkers")
dir.create(file.path(CFilePath))
setwd(file.path(CFilePath))

for (i in celltype.markers){

  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()

  ggsave(paste0(Outdir,"/CandidateMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 10)

}

for (i in Macrophage1Markers){

  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()

  ggsave(paste0(Outdir,"/CandidateMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 10)

}


for (i in Macrophage2Markers){

  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()

  ggsave(paste0(Outdir,"/CandidateMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 10)

}
#
# setwd(file.path(Outdir))


pdf("Pyelonephritis_DotPlot_Finalused2.markers.pdf", height =8, width = 20)
DotPlot(scrna, features =  FinalUssedMarkers2, col.min =-1, col.max = 2 ) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_DotPlot_candidatecelltype.markers.pdf", height =6, width = 25)
DotPlot(scrna, features =  celltype.markers) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_VlnPlot_FinalUsedMarkers.pdf", height =20, width = 25)
VlnPlot(scrna, features =  FinalUsedMarkers, split.by = "DataSet",pt.size = 0,ncol=3) 

dev.off()


pdf("Pyelonephritis_DotPlot_Clustermakrers.pdf", height =6, width = 25)
DotPlot(scrna, features =  Clustermakrers) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_DotPlot_FinalUsedMarkers.pdf", height =6, width = 15)
DotPlot(scrna, features = FinalUsedMarkers ) +RotatedAxis()
dev.off()

all.genes <- rownames(scrna)

# scrna <- ScaleData(scrna, features = all.genes)
# pdf("Pyelonephritis_Heatmap_candidatecelltype.markers.pdf", height =15, width = 25)
# DoHeatmap(scrna, features = celltype.markers) + NoLegend()
# dev.off()

```

# add the cell type
```{r}
# Cell Type
FinalCluster<- as.character(levels(factor(scrna$Sub5)))
FinalCluster
CellType <- data.frame(cluster=FinalCluster , cell= FinalCluster)
CellType
# add the cluster name
#CellType[CellType$cluster %in% c(0,3,9,12,18,19),2] <- "Proximal tubule"

#CellType[CellType$cluster %in% c(0,3,9,12,18,19),2] <- "Proximal tubule"

# CellType[CellType$cluster %in% c("0"),2] <- "Proximal tubule I"
# CellType[CellType$cluster %in% c("9"),2] <- "Proximal tubule II"
# CellType[CellType$cluster %in% c("12"),2] <- "Proximal tubule III"
# #CellType[CellType$cluster %in% c("18"),2] <- "Proximal tubule IV"
# CellType[CellType$cluster %in% c("19"),2] <- "Proximal tubule V"
# 
# 
# CellType[CellType$cluster %in% c("18"),2] <- "Glomerular Endothelial Cell"
# 

CellType[CellType$cluster %in% c("1_1","0", "9","12","18", "19","3"),2] <- "PT" # "Proximal tubule"
CellType[CellType$cluster %in% c("1_0"),2] <- "MAC" # macrophage
#CellType[CellType$cluster %in% c("1_1"),2] <- "Inflammated proximal tubule"
CellType[CellType$cluster %in% c("2"),2] <- "T" #"T cell"
CellType[CellType$cluster %in% c("4"),2] <- "DTL" # Descending thin limb cells
#CellType[CellType$cluster %in% c("4_1"),2] <- "Maladaptive limb of LOH"

CellType[CellType$cluster %in% c("5_0"),2] <- "TAL" # "Thick ascending limb cells"
CellType[CellType$cluster %in% c("5_1"),2] <- "DCT-TAL" #"Distal convoluted tubule and Thick ascending limb of LOH"

CellType[CellType$cluster %in% c("6"),2] <- "IC" #"Intercalated cells"
CellType[CellType$cluster %in% c("7"),2] <- "PC" # "Principal cells"
CellType[CellType$cluster %in% c("8"),2] <- "CNT" # "Connecting tubule"

CellType[CellType$cluster %in% c("10"),2] <- "NK" # "NK Cells"
CellType[CellType$cluster %in% c("11"),2] <- "END" # "Endothelium"
CellType[CellType$cluster %in% c("13"),2] <- "PL" # "Plasma cells"
CellType[CellType$cluster %in% c("14"),2] <- "FIB" # "Fibroblast"
CellType[CellType$cluster %in% c("15"),2] <- "URO" #"Urothelium"
CellType[CellType$cluster %in% c("16"),2] <-"B"  # "B cells"
CellType[CellType$cluster %in% c("17"),2] <- "NEU"  # "Neutrophil"

# add the cell types
scrna@meta.data$Raw_cell_type <- sapply(scrna@meta.data$Sub5, function(x){CellType[CellType$cluster %in% x,2]})
levels(factor(scrna@meta.data$Raw_cell_type))
scrna@meta.data %>% head()
levels(factor(scrna@meta.data$Sub5))
#scrna@meta.data$Sub5
# generate the dimplot for these raw_cell_type

DimClusterAnnotate <-DimPlot(scrna, reduction = "umap",group.by = "Raw_cell_type", label = T)
ggsave(filename = "DimClusterAnnotate.pdf", plot = DimClusterAnnotate, width = 6, height = 5)

Idents(scrna) <- scrna@meta.data$Raw_cell_type

Idents(scrna) <- ordered(factor(scrna@meta.data$Raw_cell_type), levels= c("PT","DCT-TAL","TAL","CNT","DTL","PC","IC","URO","END","FIB","PL","NEU","MAC","T","B","NK"))

DimClusterAnnotate <-DimPlot(scrna, reduction = "umap",group.by = "Raw_cell_type", label = T)
ggsave(filename = "DimClusterAnnotate.pdf", plot = DimClusterAnnotate, width = 6, height = 5)

scrna@meta.data %>% head()
DimClusterAnnotateSplit <-DimPlot(scrna, reduction = "umap",group.by = "Raw_cell_type", split.by = "DataSet",label = T)
ggsave(filename = "DimClusterAnnotateSplit.pdf", plot = DimClusterAnnotateSplit, width = 20, height = 5)

# save the scrna object

saveRDS(scrna, file = "Pyelonephritis_singlecell_harmony_withAnnotation_v0803.RDS")
```

# add the proportion within timepoints
```{r}
# Firstly, create a new column to combine Dataset and cluster

#scrna <-readRDS("Pyelonephritis_singlecell_harmony_withAnnotation_v0803.RDS")
scrna@meta.data %>% head()
table(scrna@meta.data$DataSet,scrna@meta.data$Raw_cell_type) %>% as.data.frame()

AllPropotionWithinTime <-table(scrna@meta.data$DataSet,scrna@meta.data$Raw_cell_type) %>% as.data.frame() %>%
  group_by(Var1) %>%
  mutate(pct= prop.table(Freq) )  %>% ggplot(aes(x=Var1, y=pct, fill=Var2)) +geom_bar(position="stack", stat="identity") + theme_classic() +scale_y_continuous(labels = scales::percent)

ggsave("CellTypePropotionWithinTime.pdf", plot = AllPropotionWithinTime, width = 4, height = 6)
```


# final used cell biomarkers
```{r}
# Cfh: Parietal epithelial cells
Finalcelltypemarkers<- c("Lrp2","Slc34a1","Slc13a3", # Proximal tubule cells (PT)
                      "Slc12a3","Pvalb","Wnk1",# Distal connecting tublue (DCT)

                     "Umod","Slc12a1","Cldn10", # Thick ascending limb cells (TAL)
                      "Klk1","Slc8a1","Calb1", # Connecting tubule (CNT)
                     "Slc14a2","Fst","Bst1", # descending thin  limb cells (DTL)
                     #"Akr1b1","Sh3gl3","Prox1", # ascending thin limb cells (ATL)
                     #"Enox1","Thsd4", #Macula densa
                      "Aqp2","Hsd11b2","Scnn1g", # principal cells (PC)
                    "Atp6v1g3","Aqp6","Slc26a7", # Intercalated cells
                    "Krt5","Upk1b","Krt14", # Urothelium
                    "Cdh5","Igfbp3","Pecam1", # Endothelium
                    "Col3a1","Fbn1","Lum","Col1a1", # Fibroblast
                    "Ireb2","Alas2","C78334", # Plasmacytoid dentritic cells
                    "S100a8","S100a9","Il1b", # Neutrophil
                    "C1qa","C1qb","Aif1", # Macrophage
                    "Cxcr6","Cd247","Cd3e", # T cells
                    "Igkc","Cd79a","Cd79b", # B cells
                    "Ccl5","Nkg7","Cd7" # NK/Cd8 cells
)

MarkerGenesDotPlots<- DotPlot(scrna, features =  Finalcelltypemarkers, col.min = -0.5) +RotatedAxis()
ggsave(filename = "MarkerGenesExpression_ClusterAnnotate.pdf", plot = MarkerGenesDotPlots, width = 15, height = 8)


for (i in Finalcelltypemarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/CandidateMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}


```








# The following is for PT, please ignore it. We did have the another script focusing on PT cluster
# generate the health cluster of PT cells
```{r}
# We only collect the PT cells
CellType[CellType$cluster %in% c("1_1","0", "9","12","18", "19","3"),2] <- "PT" # "Proximal tubule"

scrna @meta.data

PTcells <- subset(scrna, subset = Raw_cell_type == "PT")

PTcells@meta.data
# We then recluster the PT cells
# recluster the cells
PTcells <- SCTransform(PTcells, verbose = TRUE)
PTcells <- RunPCA(PTcells,npcs = 5, verbose = TRUE)

# Run harmony for the PT cells
PTcells <- RunHarmony(PTcells, 
				group.by.vars = c("DataSet"), 
				reduction = "pca", assay.use = "SCT", reduction.save = "harmony")
PTcells <- RunUMAP(PTcells, dims = 1:5, verbose = TRUE, reduction = "harmony")
PTcells <- FindNeighbors(PTcells, dims = 1:5,  verbose = TRUE, reduction = "harmony")

PTcells<- FindClusters(PTcells,  verbose = TRUE, resolution = 0.2)

PTcellsDim<- DimPlot(PTcells,label = T)
PTcellsDim
ggsave(filename = "PTcellsDim.pdf", plot = PTcellsDim, width = 8, height = 6)

PTcellsDimSplit<-DimPlot(PTcells, split.by = "DataSet",label = T)

ggsave(filename = "PTcellsDimSplit.pdf", plot = PTcellsDimSplit, width = 18, height = 6)

```

# check the markers
```{r}
# the healthy cells
HealthMarkers<- c("Slc22a30", "Atp11a", "Trim7", "Slc22a6", "Slc7a13", "Slc5a2", "Slc5a10", "Spp2",  "Gatm", "Dcxr")

DotPlot(PTcells, features =  HealthMarkers, col.min = -0.5) +RotatedAxis()

CyclingMarker <- "Mki67"
DotPlot(PTcells, features =  CyclingMarker, col.min = -0.5) +RotatedAxis()

EarlyInjuryMarkers <- c("Cdh13", "Cdh6", "Havcr1")
DotPlot(PTcells, features =  EarlyInjuryMarkers, col.min = -0.5) +RotatedAxis()
LateInjuryMarkers <-c( "Ccl2", "Gm17268", "Cald1")
DotPlot(PTcells, features =  LateInjuryMarkers, col.min = -0.5) +RotatedAxis()

ReparingMarkers <- "Top2a"

DotPlot(PTcells, features =  ReparingMarkers, col.min = -0.5) +RotatedAxis()

FailedFepairMarkers <- c("Ypel2","Kcnip4")

DotPlot(PTcells, features =  FailedFepairMarkers, col.min = -0.5) +RotatedAxis()
# 
MaladptiveMarkers <- c("Vcam1", "Havcr1",  "Icam1", "Prom1","Cxcl2","Ccl3","Myc")

DotPlot(PTcells, features =  MaladptiveMarkers, col.min = -0.5) +RotatedAxis()


# Combined 
#"Lrp2","Slc34a1","Slc13a1"
KidneyStatuesMarkers<- c( "Slc5a2","Slc34a1","Inmt","Apom", "Slc7a13","Slc22a6", "Prom1","Top2a","Cyp7b1","Mki67","Cdh13", "Cdh6", "Havcr1","Ccl2", "Gm17268", "Cald1","Vcam1", "Ypel2","Kcnip4","Hsp90aa1","Col27a1","Plin2")

 DotPlot(PTcells, features =  KidneyStatuesMarkers,col.min = -1) +RotatedAxis()

KidneyStatuesTypeMarkers<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "Spp1", "Cst3","Gdf15", "Clu", "Igfbp7", "Apoe", "Aldob", "B2m", "S100a6","S100a8","S100a9","Pdgfb","Bak1","Bad","Casp8","Ptgs2","Casp1","Hmox1")
 DotPlot(PTcells, features =  KidneyStatuesTypeMarkers) +RotatedAxis()
 
 KidneyStatuesTypeMarkers2<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "S100a6","S100a8","S100a9","Pdgfb","Bak1","Bad","Casp8","Ptgs2","Casp1","Hmox1","Nfkbia")
  DotPlot(PTcells, features =  KidneyStatuesTypeMarkers2) +RotatedAxis()

 #, cluster 5: Fn1, Havcr1, Cst3, Igfbp7, S100a8, Il1b, Slc7a13 : Profibrotic PT
 # cluster 1: Spp1, Prom1, Kcnip4, Col27a1,Top2a,Nfkbia: Cycling PT
 # cluster 8: Cdh13, Hsp90aa1,Inmt, Mki67 : Proliferating PT
 
 # cluster 1: ASpp1, Prom1, Kcnipp4: Progenitor PT
 
 # cluster 0: Aldob
```

# the biomarkers for the 
```{r}
CandidateMarkers <- c("Aldob","Slc34a1","Spp1","Spp2", "Slc22a8","Slc5a2","Inmt","Prom1","Defb1","Top2a","Nfkbia", "Kcnip4", "Slc22a6", "Apoe","Havcr1", "Fn1","C3", "Il1b","S100a8", "Igfbp7", "Cst3", "Col27a1", "Hsp90aa1","Cdh13")

CandidateMarkersDotPlot<-  DotPlot(PTcells, features =  CandidateMarkers, col.min = 0) +RotatedAxis()


ggsave("PT_CandidateMarkersDotPlot.pdf", plot = CandidateMarkersDotPlot, height = 6, width = 12)
```

# #####################################
## generate the Trajectories of PT cells
# #####################################
# 2.2 Prepared cds the object
```{r}

# build cds object

PTcells_expr_matrix = GetAssayData(PTcells, slot = "counts")

# Cell meta annotation
p_data <- PTcells@meta.data 
head(p_data)

# Gene meta annotation
f_data<- data.frame(gene_short_name = row.names(PTcells), row.names = row.names(PTcells))


pre_cds <- new_cell_data_set(PTcells_expr_matrix,
                         cell_metadata = p_data,
                         gene_metadata = f_data)

pData(pre_cds) %>% colnames()

cds <- preprocess_cds(pre_cds) #标准化+PCA降维

```

# 2.2 Umap by Monocle3
```{r}

cds <- reduce_dimension(cds, preprocess_method = "PCA")
head(cds@int_colData$reducedDims$UMAP)

plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="SCT_snn_res.0.2")
dev.off()
# UMAP by Seurat
seurat_umap <- Embeddings(PTcells, reduction = "umap")[colnames(cds),]
cds@int_colData$reducedDims$UMAP <- seurat_umap
head(cds@int_colData$reducedDims$UMAP)
plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="SCT_snn_res.0.2")

```

# Draw the monocle state
```{r}
#分群(类似monocle的State)

cds <- cluster_cells(cds)

#预测轨迹
cds <- learn_graph(cds)
#交互式确定root节点，可以选择多个。我这里选择了一个
cds <- order_cells(cds) 

PT_pseudotime <-plot_cells(cds, color_cells_by = "pseudotime", 
           label_cell_groups = FALSE, 
           label_leaves = TRUE,  
           label_branch_points = FALSE)

PT_pseudotime
ggsave(filename = "PT_pseudotime_MonocaCluster.pdf", plot = PT_pseudotime, width = 4, height = 3)


```

```{r}
# Determine the psedudotime related genes
Track_genes <- graph_test(cds, neighbor_graph="principal_graph")

Track_genes %>% head()
library(monocle3)
# select the top
```


```{r}
Track_genes <- Track_genes[,c(5,2,3,4,1,6)] %>% 
  dplyr::arrange(desc(morans_I),q_value)
KapTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[1],] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_KapTrack.pdf", plot = KapTrack, width = 5, height = 3)
NapsaTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[2],] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_NapsaTrack.pdf", plot = NapsaTrack, width = 5, height = 3)

TtrTrack <- plot_genes_in_pseudotime(cds["Ttr",] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_TtrTrack.pdf", plot = TtrTrack, width = 5, height = 3)

Igfbp7Track <- plot_genes_in_pseudotime(cds["Igfbp7",] , 
                         min_expr=0.5)
ggsave(filename = "PT_pseudotime_Igfbp7Track.pdf", plot = Igfbp7Track, width = 5, height = 3)

Track_genes$gene_short_name[1:5]


Track_genes$gene_short_name[2]
Track_genes$gene_short_name[1]
Havcr1Track <- plot_genes_in_pseudotime(cds["Havcr1",] , 
                         min_expr=0.5)

ggsave(filename = "PT_pseudotime_Havcr1Track.pdf", plot = Havcr1Track, width = 5, height = 3)

# For the 
Prom1Track <- plot_genes_in_pseudotime(cds["Prom1",],min_expr=0.5 )
ggsave(filename = "PT_pseudotime_Prom1Track.pdf", plot = Prom1Track, width = 5, height = 3)

FeaturePlot(PTcells, features = "Havcr1")

```

# choose the top signifcant for visualization
```{r}
diff_pseudo_gene <- Track_genes %>% 
  dplyr::arrange(q_value) %>%
  # remove the genes with mitochondrion
  filter(!grepl('^mt-|Hbb-', gene_short_name)) %>%
  rownames() %>% head(60)

# we ignore the mitochondria genes



diff_pseudo_gene

# generate the heatmap for those genes

diff_pseudo_gene.matrix <- exprs(cds)[match(diff_pseudo_gene,rownames(rowData(cds))),order(pseudotime(cds))]

diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){smooth.spline(x,df=3)$y}))
diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){(x-mean(x))/sd(x)}))
rownames(diff_pseudo_gene.matrix) <- diff_pseudo_gene;

#pheatmap(diff_pseudo_gene.matrix)

library(ComplexHeatmap)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(circlize)
library(monocle3)
#install.packages("magick")
library(magick)
#K means with 6 groups


# because the plot_pseudotime_heatmap no longer have, we used the complexheatmap to draw the figure
# Reference: https://github.com/cole-trapnell-lab/monocle-release/issues/295
pdf("PT_pseudotime_psudotimeGenes.pdf", height = 6, width = 8)
htkm <- Heatmap(
  diff_pseudo_gene.matrix,
  name    = "z-score",
  col     = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names  = TRUE,
  show_column_names = FALSE,
  row_names_gp  = gpar(fontsize = 6),
  km = 8,
  row_title_rot = 0,
  use_raster=TRUE,
  cluster_rows  = TRUE,
  cluster_row_slices = FALSE,
  cluster_columns   = FALSE)

print (htkm)
dev.off()
#ggsave(filename = "PT_pseudotime_psudotimeGenes.pdf", plot = htkm, width = 8, height = 6)


#Ward.D2 Hierarchical Clustering
pdf("PT_pseudotime_psudotimeGenes_Hierarchicalcluster.pdf", height = 7, width = 6)
hthc <- Heatmap(
  diff_pseudo_gene.matrix,
  name                         = "z-score",
  col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names               = TRUE,
  show_column_names            = FALSE,
  row_names_gp                 = gpar(fontsize = 6),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  use_raster= TRUE,
  cluster_rows                 = TRUE,
  cluster_row_slices           = FALSE,
  cluster_columns              = FALSE)

print(hthc)
dev.off()

cds["Havcr1",] %>% head()
# Generate the Dotplot for these genes
plot_genes_in_pseudotime(cds[c("Havcr1"),] , 
                         min_expr=0.5)

DotPlot(PTcells, features =diff_pseudo_gene ) +RotatedAxis()

DotPlot(PTcells, features =c("Havcr1","Itgb1",
                             "Prom1","Pdgfb","Defb1","Lcn2"), col.min = -0.5 ) +RotatedAxis()

# the following no longer have 

Track_genes
p = plot_pseudotime_heatmap(cds[diff_pseudo_gene,], 
                        num_clusters = 3,  # default 6
                        return_heatmap=T)



```





# generate the subclusters of PT, DCT, IC, PC
```{r}
# dividing into cycling, transitioning, adap- tive (successful or maladaptive repair) and degenerative (damaged or stressed). 
# cycling: Mki67
# Early Injury: Cdh13, Cdh6, Havcr1
# Late Injury: Ccl2, Gm17268, Cald1
# conserved: Slc13a1, Slc5a2,  Slc7a13, Cyp7b1



# Panoptosis
# Il1b, MLKL, Gsdmd,Ripk3

# Fibrosis 
# Fn1, Col1a1, Col3a1, Tgfb1
# adaptive
# VCAM1, DCDC2 and HAVCR1, C3, ICAM1
# VCAM1 PROM1 ITGB8
# Vcam1, Havcr1, Dcdc2, Icam1, Prom1, ITgb8

#degenerative
# SPP1, CST3, CLU and IGFBP7, APOE ALDOB
# B2M S100A6
# Spp1, Cst3, GDF15,Clu, Igfbp7, Apoe, Aldob, B2m, S100a6
# agionegious PEC,
# EGFR, VEGF, Cdh5

# Egfr, Vegf, Cdh5
KidneyStatuesMarkers<- c("Slc13a1", "Slc5a2", "Slc7a13", "Cyp7b1","Mki67","Cdh13", "Cdh6", "Havcr1","Ccl2", "Gm17268", "Cald1")

KidneyStatuesTypeMarkers<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "Spp1", "Cst3","Gdf15", "Clu", "Igfbp7", "Apoe", "Aldob", "B2m", "S100a6")

DotPlot(scrna, features =  KidneyStatuesMarkers, col.min = -0.5) +RotatedAxis()
DotPlot(scrna, features =  KidneyStatuesTypeMarkers, col.min = -0.5) +RotatedAxis()


for (i in KidneyStatuesMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}

for (i in KidneyStatuesTypeMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesTypeMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}





# with the inflammation markers

# with the fibrosis markers

# with the apoptosis markers

# with the necrosis markers

# with the endothelial markers


```



# seting the identity

```{r}

Idents(rnaAggrSingle) <- rnaAggrSingle@meta.data$seurat_clusters
rnaAggrSingle<-RenameIdents(rnaAggrSingle, `0`= "Proximal tubule I", `1`= "Myeloblast", `2`= "T cells", 
                    `3`= "Adaptive proximal tubule", `4`= "Thin limb of LOH", `5`= "Thick ascending limb of LOH", 
                    `6`= "Intercalated cells", `7`= "Principle cells", `8`= "Distal convoluted", `9`= "Parietal epithelial cells",
                    `10`= "NK cells", `11`= "Endothelium", `12`= "Proximal tubule II", 
                    `13`= "Plasmacytoid dendritic cell", `14`= "Fibroblast", `15`= "Urothelium", `16`= "B cells",`17`= "Granulocyte",`18`= "Proximal tubule", "")



# reorder the levels
levels(x=rnaAggrSingle) <- c("Proximal tubule I", "Proximal tubule II", "Proximal tubule III" ,"Proximal tubule IV","Proximal tubule V","Distal convoluted/connecting tubule","Thin limb of LOH",  "Thick ascending limb of LOH", "Intercalated cells","ICA","Principle cells","Urothelium", "Endothelium/Fibroblast","Glomerular endothelial cell","Myeloblast","Granulocyte",
                     "T cells","B cells",  "NK cells" ,"Plasmacytoid dendritic cell")



```
##  visualization of the dimplot using the harmony with annotated cells
```{r}

pdf("Pyelonephritis_DimPlot_samples_dataset_ann.pdf", height =5, width = 20)
DimHarmonySamples<-DimPlot(rnaAggrSingle, reduction = 'umap', split.by = 'DataSet', label = TRUE)
DimHarmonySamples
dev.off()

pdf("Pyelonephritis_DimPlot_samples_condition_ann.pdf", height =5, width = 10)
DimHarmonySamplesTec<-DimPlot(rnaAggrSingle, reduction = 'umap', split.by = 'Condition', label = TRUE)
DimHarmonySamplesTec
dev.off()

pdf("Pyelonephritis_DimPlot_all_ann.pdf", height =10, width = 12)
DimHarmonySamplesSam<-DimPlot(rnaAggrSingle, reduction = 'umap', label = TRUE)
DimHarmonySamplesSam
dev.off()
```


# step 7 generate the biomarker distribution
```{r}

DefaultAssay(rnaAggrSingle) <-"RNA"

# Cfh: Parietal epithelial cells
celltype.markers<- c("Lrp2","Slc34a1","Acsm1","Cubn","Havcr1","Fn1","Col1a1","Il6","Il1a","Il1b","Tnf","Mmp3","Mmp9","Slc5a1","Slc5a2","Cfh","Vcam1","Ireb2","Alas2","C78334","Slc12a3","Pvalb","Wnk1","Atp6v1g3","Aqp6", "Atp6v0d2","Slc4a1","Slc26a7","Slc26a4",
                   "Aqp2","Hsd11b2","Scnn1g","Fst","Bst1","Slc14a2","Umod","Slc12a1","Cldn10","Krt5","Krt14","Upk1b","Nphs1","Nphs2",
                   "Pecam1","Cdh5","Igfbp3","Acta2","Ehd3","Nrp1","Col3a1","Lum","Fbn1","Prkg1","C1qa","C1qb","Aif1","Ptprc","Cd3e","Cd14","Lst1","Lyz2","Cebpb","Cxcr6","Cd247","Nkg7","Cd7","Ccl5","Cd79a","Cd79b","Ms4a1","Igkc","Ms4a7","Adap2")

Clustermakrers<- c("Lrp2","Slc34a1","Acsm1","Slc5a2","Slc5a1","Havcr1","Fn1","Col1a1","Notch2","Enox1","Klh13","Atp11a","Cfh","Ireb2","Alas2","C78334","Slc12a3","Pvalb","Wnk1","Atp6v1g3","Aqp6","Slc4a1","Rcan2","Kit","Slc26a4","Hmx2","Spink8","Slc26a7","Slc8a1", "Trpv5","Nphs1", "Calb1","Skap1", "Upk2a","Upk3a","Upk3b",
                   "Aqp2","Atp6v1b1","Hsd11b2","Scnn1g","Fst","Bst1","Slc14a2","Umod","Slc12a1","Cldn10","Krt5","Krt14","Upk1b", 
                   "Pecam1","Cdh5","Nrp1","Col3a1","Lum","Fbn1","C1qa","C1qb","Aif1","Fhl2","Lst1","Lyz2","Ptprc","Cebpb","Cxcr6","Cd247","Nkg7","Cd7","Ccl5","Cd79a","Cd79b","Ms4a1")

FinalUsedMarkers<- c("Lrp2","Slc34a1","Acsm1","Atp11a","Slc5a2","Atp6v1g3","Slc26a4","Hmx2","Spink8","Slc26a7","Aqp6","Slc4a1","Rcan2","Kit","Aqp2","Hsd11b2" ,"Scnn1g","Slc8a1","Slc12a3","Wnk1","Fst","Bst1","Slc14a2","Umod","Slc12a1","Nphs1","Nrp1","Col3a1","Fbn1", "Fhl2","Lst1","Lyz2","Ptprc","Atp6v1b1")

ICAMarkers <- c("Slc26a7","Rcan2","Atp4a","Slc4a1","Aqp6","Kit")
ICBMarkers<- c("Slc26a4","Hmx2","Spink8")
PCMarkers <- c("Aqp2","Hsd11b2" ,"Scnn1g")

#FinalUssedMarkers2<- c("Lrp2","Slc34a1","Acsm1","Cubn","Havcr1","Slc12a3","Trpm6","Pvalb","Wnk1","Slc8a1","Calb1","Trpv5","Slc12a1","Cldn10","Atp6v1g3","Aqp6","Slc26a4","Aqp2","Hsd11b2" ,"Scnn1g","Fst","Bst1","Slc14a2","Krt5","Krt14","Upk1b", "Pecam1","Cdh5","Igfbp3","Col3a1","Lum","Fbn1","Ireb2","Alas2","C78334","Ptprc","Lyz2","Lst1","C1qa","C1qb","Cxcr6","Cd247", "Nkg7","Cd7","Ccl5","Cd79b","Ms4a1","Igkc","Il1b","Cd14")

FinalUssedMarkers2<- c("Lrp2","Slc34a1","Acsm1","Cubn","Havcr1","Slc12a3","Slc8a1","Calb1","Trpm6","Pvalb","Wnk1","Trpv5","Fst","Slc14a2","Slc12a1","Cldn10","Atp6v1g3","Aqp6","Slc26a4","Aqp2","Hsd11b2" ,"Scnn1g","Krt5","Krt14","Upk1b", "Col3a1","Lum","Fbn1","Pecam1","Igfbp3","Cdh5","Ptprc","Lyz2","Lst1","C1qa","C1qb","Il1b","Cd14","Cxcr6","Cd247", "Cd79b","Ms4a1","Igkc","Nkg7","Cd7","Ccl5","Ireb2","Alas2","C78334")


# CFilePath<- paste0(Outdir,"Candidatemarkers")
# dir.create(file.path(CFilePath))
# setwd(file.path(CFilePath))
# 
# for (i in celltype.markers){
#   
#   FeaturePlots<-FeaturePlot(rnaAggrSingle, features =  i, split.by = "DataSet" ) +RotatedAxis()
# 
#   ggsave(paste0("CandidateMarkers", i, "Diabiet_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 10)
#   
# }
# 
# 
# setwd(file.path(Outdir))

?DotPlot

pdf("Pyelonephritis_DotPlot_Finalused2.markers.pdf", height =8, width = 20)
DotPlot(rnaAggrSingle, features =  FinalUssedMarkers2, col.min =-1, col.max = 2 ) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_DotPlot_candidatecelltype.markers.pdf", height =6, width = 25)
DotPlot(rnaAggrSingle, features =  celltype.markers) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_VlnPlot_FinalUsedMarkers.pdf", height =20, width = 25)
VlnPlot(rnaAggrSingle, features =  FinalUsedMarkers, split.by = "DataSet",pt.size = 0,ncol=3) 

dev.off()


pdf("Pyelonephritis_DotPlot_Clustermakrers.pdf", height =6, width = 25)
DotPlot(rnaAggrSingle, features =  Clustermakrers) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_DotPlot_FinalUsedMarkers.pdf", height =6, width = 15)
DotPlot(rnaAggrSingle, features = FinalUsedMarkers ) +RotatedAxis()
dev.off()

all.genes <- rownames(rnaAggrSingle)

rnaAggrSingle <- ScaleData(rnaAggrSingle, features = all.genes)
pdf("Pyelonephritis_Heatmap_candidatecelltype.markers.pdf", height =15, width = 25)
DoHeatmap(rnaAggrSingle, features = celltype.markers) + NoLegend()
dev.off()

```


# add the dotplot for the innative immune genes and adaptive immune genes
```{r}
AdaptiveImmune<- read.csv(file = "/Users/XXW004/Documents/Projects/RuizRosado/SpatialTranscriptomic/Results/Deconvolution_V04142023/Seurat/SelectedGenesFromMicroArray/HeN Adaptive Immunity Comprehensive.csv", header = T)
InnativeImmune<-read.csv(file ="/Users/XXW004/Documents/Projects/RuizRosado/SpatialTranscriptomic/Results/Deconvolution_V04142023/Seurat/SelectedGenesFromMicroArray/HeN Innate Immune Response Comprehensive.csv",  header = T)
AllInnativeImmuneGenes<- InnativeImmune %>% pull(geneID) %>% strsplit("/") %>% unlist() %>% unique()

AllAdaptiveImmuneGenes<- AdaptiveImmune %>% pull(geneID) %>% strsplit("/") %>% unlist() %>% unique()

# check the trend within timepoints


AllInnativeImmuneGene2 <-  AllInnativeImmuneGenes[AllInnativeImmuneGenes !="Pde4d"]

pdf("Pyelonephritis_DotPlot_innativeimmunegenes.pdf", height =6, width = 25)
DotPlot(rnaAggrSingle, features =  AllInnativeImmuneGenes) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_DotPlot_adaptiveimmunegenes.pdf", height =6, width = 15)
DotPlot(rnaAggrSingle, features = AllAdaptiveImmuneGenes ) +RotatedAxis()
dev.off()

AllInnativeImmuneGenes["Pde4d"]

Idents(rnaAggrSingle)<-rnaAggrSingle@meta.data$DataSet
pdf("Pyelonephritis_DotPlot_innativeimmunegenes_timepoints.pdf", height =6, width = 25)
DotPlot(rnaAggrSingle, features =  AllInnativeImmuneGenes) +RotatedAxis()
dev.off()

pdf("Pyelonephritis_DotPlot_adaptiveimmunegenes_timepoints.pdf", height =6, width = 15)
DotPlot(rnaAggrSingle, features = AllAdaptiveImmuneGenes ) +RotatedAxis()
dev.off()

Fibrosis <- c("Fn1","Col1a1","Col3a1","Tgfb1")
pdf("Pyelonephritis_DotPlot_Fibrosisgenes_timepoints.pdf", height =6, width =6)
DotPlot(rnaAggrSingle, features = Fibrosis ) +RotatedAxis()
dev.off()


```



