---
title: "PipseqImmuneSubcluster_V0822"
author: "Xin Wang"
date: "2023-08-22"
description: Immune subclusters analyses:
  - Recluster of the immune cells
  - Annotation of subclusters using the candidate markers
  - Psudotime of Microphage
  - Basic statistics of Microphage
output: html_document
---
# Loading the packages
```{r}
# check the session information
sessionInfo()
library(dplyr)
library(Seurat)
library(patchwork)

#install.packages('devtools') #assuming it is not already installed

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
library(devtools)
#install_github('andreacirilloac/updateR')
# install.packages("sctransform")
# install.packages("ggthemes")
# loading the libarary
library(updateR)

library("sctransform")
library("ggthemes")
library(Seurat)

# BiocManager::install("limma")
# BiocManager::install("ComplexHeatmap")
# BiocManager::install("biomaRt")
# BiocManager::install("methrix")BiocManager::install("org.Hs.eg.db")
# devtools::install_github("saeyslab/nichenetr")
library(limma)
library(nichenetr)
library(tidyverse)
library(dplyr)

library(biomaRt)
library("data.table")


library(umap)
library(patchwork)
library(cowplot)
library(installr)

#install.packages("Matrix", repos = "http://cran.r-project.org")
### we used discre color palettes from ggsci

library("ggsci")
library("ggplot2")
library("gridExtra")

#options(buildtools.check = function(action) TRUE )
#devtools::install_url('https://cran.r-project.org/src/contrib/Archive/Matrix.utils/Matrix.utils_0.9.7.tar.gz')
#devtools::install_github('cole-trapnell-lab/monocle3')
#library(monocle3)

library(harmony)
library(Seurat)
#library(SeuratData)
library(tidyverse)

#library(BPCells)
library(ggplot2)
library("devtools")
library("AnnotationDbi")
library("org.Hs.eg.db")
#library(tibble)
#library(future)
library(here)
library(patchwork)
library(future)
library(monocle3)
#devtools::install_dev("remotes")
#remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
#remotes::install_github(repo ='chris-mcginnis-ucsf/DoubletFinder' )
#install.packages("DoubletFinder")
library(DoubletFinder)
library(future.callr)
library(ComplexHeatmap)

library(RColorBrewer)
library(circlize)
library(monocle3)
#install.packages("magick")
library(magick)
#K means with 6 groups

```


## Step 2: Set up the workspace environment:
```{r}
rm(list=ls())
Indir =c("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/")
Outdir =c("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Results/Evaluation_V022723/CellEvaluation_Subclustering_V0803/Immune_Subclusters/")

DEGDir <- paste0(Outdir,"DEGs/")
# directory for Progeny
ProgenyDir <- paste0(Outdir,"Progeny/")
# directory for GO enrichments
GOenrichDir <- paste0(Outdir,"GOenrich/")
# directory for Transcript Factors
TFDir <-  paste0(Outdir,"TranscriptFactors/")
# directory for KEGG
KEGGDir <-  paste0(Outdir,"KEGG/")
# directory for CellCommunication
CellCommDir <-  paste0(Outdir,"CellCommunication/")
# directory for pyseudotime
PyseudotimeDir <-  paste0(Outdir,"Pyseudotime/")

# directory for biomarkers

BioMarkerDir <- paste0(Outdir,"Biomarkers/")

dir.create(DEGDir, showWarnings = FALSE)
dir.create(ProgenyDir, showWarnings = FALSE)
dir.create(GOenrichDir, showWarnings = FALSE)
dir.create(TFDir, showWarnings = FALSE)
dir.create(KEGGDir, showWarnings = FALSE)
dir.create(CellCommDir, showWarnings = FALSE)
dir.create(BioMarkerDir, showWarnings = FALSE)
dir.create(PyseudotimeDir, showWarnings = FALSE)
# setting up the seed 
set.seed(10000)
```


# generate the health cluster of PT cells

```{r}
setwd(Outdir)
# Read the preclustered object

scrna <- readRDS(file = paste0(Indir,"Pyelonephritis_singlecell_harmony_withAnnotation_v0803.RDS"))

# We only collect the PT cells
#CellType[CellType$cluster %in% c("1_1","0", "9","12","18", "19","3"),2] <- "PT" # "Proximal tubule"

scrna @meta.data

IMMcells <- subset(scrna, subset = Raw_cell_type == "MAC"| Raw_cell_type == "NEU"| Raw_cell_type == "B"| Raw_cell_type == "NK"| Raw_cell_type == "T")
IMMcells

IMMcells@meta.data
# We then recluster the Immune cells
# recluster the cells
IMMcells <- SCTransform(IMMcells, verbose = TRUE)
IMMcells <- RunPCA(IMMcells,npcs = 10, verbose = TRUE)

# Run harmony for the Immune cells
IMMcells <- RunHarmony(IMMcells, 
				group.by.vars = c("DataSet"), 
				reduction = "pca", assay.use = "SCT", reduction.save = "harmony")
IMMcells <- RunUMAP(IMMcells, dims = 1:10, verbose = TRUE, reduction = "harmony")
IMMcells <- FindNeighbors(IMMcells, dims = 1:10,  verbose = TRUE, reduction = "harmony")

IMMcells<- FindClusters(IMMcells,  verbose = TRUE, resolution = 0.3)

IMMcellsDim<- DimPlot(IMMcells,label = T)

ggsave(filename = "IMMcellsDim.pdf", plot = IMMcellsDim, width = 8, height = 6)

IMMcellsDimSplit<-DimPlot(IMMcells, split.by = "DataSet",label = T)
IMMcellsDimSplit 
ggsave(filename = "IMMcellsDimSplit.pdf", plot = IMMcellsDimSplit, width = 24, height = 6)

```


```{r}

# change the levels of the immune cells

levels(IMMcells) <- c("4","5","0","6","1","2","3")

# FinalMarkers <- c("Il1b","S100a8","S100a9", "Cd79a","Cd79b","C1qa","C1qb","Itgam","Cd81","Ctsh","Mgl2","Cx3cr1", "Ccr2", "Nos2","Tnf","Ifngr1","Arg1","Mrc1","Il13","Cd163","Cd3g", "Cxcr6","Cd247","Cd3e","Capg","Lef1","Tcf7","Gzmk","Gimap7","Ms4a4b","Ccl5","Nkg7","Cd7","Lrp2","Umod","Aqp2","Igfbp7","Havcr1","Lcn2","Ly6c1")
CFilePath<- paste0(Outdir,"Candidatemarkers")

dir.create(file.path(CFilePath))
setwd(file.path(CFilePath))


FinalMarkers <- c("Ccl5","Nkg7","Cd7","Cd79a","Cd79b","Ms4a1","Cd3g", "Cxcr6","Cd247","Cd3e","Il1b","S100a8","S100a9", "C1qa","C1qb","Itgam","Ctsh","Cd81","Mgl2","H2-Ab1","Cx3cr1", "Ccr2", "Nos2","Tnf","Arg1","Mrc1","Il13","Lrp2","Umod","Aqp2","Igfbp7","Havcr1","Lcn2","Ly6c1")
MYEGenesDotPlots<- DotPlot(IMMcells, features =  FinalMarkers, col.min = -1) +RotatedAxis()
MYEGenesDotPlots
ggsave(filename = "MYEGenesDotPlots.pdf", plot = MYEGenesDotPlots, width = 10, height = 5)


# 
# FinalMarkers <- c("S100a8","S100a9","Il1b","Itgam","Ly6g","Ly6c1", "Cd79a","Cd79b","C1qa","C1qb","Ctsh","Selenop","Cd81","Mgl2","H2-Ab1","Cx3cr1", "Ccr2", "Nos2","Tnf","Ifngr1","Arg1","Lrp2","Umod","Aqp2","Havcr1","Lcn2","Igfbp7")
# 
# 
# # Cfh: Parietal epithelial cells
# Finalcelltypemarkers<- c("Lrp2","Slc34a1","Slc13a3", # Proximal tubule cells (PT)
#                       "Slc12a3","Pvalb","Wnk1",# Distal connecting tublue (DCT)
# 
#                      "Umod","Slc12a1","Cldn10", # Thick ascending limb cells (TAL)
#                       "Klk1","Slc8a1","Calb1", # Connecting tubule (CNT)
#                      "Slc14a2","Fst","Bst1", # descending thin  limb cells (DTL)
#                      #"Akr1b1","Sh3gl3","Prox1", # ascending thin limb cells (ATL)
#                      #"Enox1","Thsd4", #Macula densa
#                       "Aqp2","Hsd11b2","Scnn1g", # principal cells (PC)
#                     "Atp6v1g3","Aqp6","Slc26a7", # Intercalated cells
#                     "Krt5","Upk1b","Krt14", # Urothelium
#                     "Cdh5","Igfbp3","Pecam1", # Endothelium
#                     "Col3a1","Fbn1","Lum","Col1a1", # Fibroblast
#                     "Ireb2","Alas2","C78334", # Plasmacytoid dentritic cells
#                     "S100a8","S100a9","Il1b","Lcn2", # Neutrophil
#                     "C1qa","C1qb","Aif1", # Macrophage
#                     "Cxcr6","Cd247","Cd3e", # T cells
#                     "Igkc","Cd79a","Cd79b", # B cells
#                     "Ccl5","Nkg7","Cd7", # NK/Cd8 cells
#                     "Hdc","Ifitm1","Ncr",
#                     "Mmp9","Ly6c1",
#                     "Havcr1")
# 
# MarkerGenesDotPlots<- DotPlot(IMMcells, features =  Finalcelltypemarkers, col.min = -0.5) +RotatedAxis()
# MarkerGenesDotPlots
# ggsave(filename = "MarkerGenesExpression_ClusterAnnotate.pdf", plot = MarkerGenesDotPlots, width = 15, height = 8)
# 

for (i in FinalMarkers){
# 
  FeaturePlots<-FeaturePlot(IMMcells, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/CandidateMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =4, width = 14)
}

# heatmap for all the cells

DoHeatmap(IMMcells, features = FinalMarkers)

```

# add the Immune cell names
```{r}
# add the annotation using case_when
setwd(Outdir)

IMMcells@meta.data %>% head() 
  
levels(IMMcells@meta.data$SCT_snn_res.0.3)
IMMcells@meta.data <- IMMcells@meta.data %>%   mutate(ImmuneIntegrateAnnot = 
           case_when(SCT_snn_res.0.3 %in% c("0") ~ "T", # T cells
                     SCT_snn_res.0.3 %in% c("1") ~ "RecMAC", # Recruited MAC
                     SCT_snn_res.0.3 %in% c("2") ~ "M1_M2", # M1/M2 Macrophage
                     SCT_snn_res.0.3 %in% c("3") ~ "ResIMM", # Resident immune
                     SCT_snn_res.0.3 %in% c("4") ~ "NK", # NK cell
                     SCT_snn_res.0.3 %in% c("5") ~ "B", # B 
                     SCT_snn_res.0.3 %in% c("6") ~ "NEU", # Neutrophil
                 )) 

IMMcells@meta.data %>% head()

Idents(IMMcells) <- IMMcells@meta.data$ImmuneIntegrateAnnot
Idents(IMMcells)

# Reorder the levels:
Idents(IMMcells)<-ordered(factor(Idents(IMMcells)),levels=c("NK","B","T","NEU","RecMAC","M1_M2","ResIMM"))

FinalMarkersWithAnn <- c("Ccl5","Nkg7","Cd7","Cd79a","Cd79b","Ms4a1","Cd3g", "Cxcr6","Cd247","Cd3e","Il1b","S100a8","S100a9", "C1qa","C1qb","Itgam","Ctsh","Cd81","Mgl2","H2-Ab1","Cx3cr1", "Ccr2", "Nos2","Tnf","Arg1","Mrc1","Il13","Lrp2","Umod","Aqp2","Igfbp7","Havcr1","Lcn2","Ly6c1")

MYEGenesDotPlotsWithAnn<- DotPlot(IMMcells, features =  FinalMarkersWithAnn, col.min = -1) +RotatedAxis()
MYEGenesDotPlotsWithAnn
ggsave(filename = "MYEGenesDotPlots_WithAnn.pdf", plot = MYEGenesDotPlotsWithAnn, width = 12, height = 5)

IMMcellsDimWithAnno<- DimPlot(IMMcells,label = T)
IMMcellsDimWithAnno
ggsave(filename = "IMMcellsDimWithAnno.pdf", plot = IMMcellsDimWithAnno, width = 6, height =4)

```

# add the Immune cells types into the orginal datasets
```{r}
IMMcells@meta.data %>% head()

ImmuneSubAnnotation <- IMMcells@meta.data %>% dplyr::select(ImmuneIntegrateAnnot)
ImmuneSubAnnotation %>% head()
## add this annotation into the meta dataset
scrna@meta.data<- scrna@meta.data %>% rownames_to_column('CellID') %>% 
  left_join( ImmuneSubAnnotation %>% rownames_to_column('CellID')) %>% 
# then add a new column that change the NA into Raw_cell_type
  # mutate(NewAnnoation = coalesce(PTIntegrateAnnot,Raw_cell_type))
  mutate(ImmuneSub = ifelse(is.na(ImmuneIntegrateAnnot), Raw_cell_type,ImmuneIntegrateAnnot))
### join the meta data to meta annoation

Idents(scrna) <- scrna@meta.data$ImmuneSub

scrna@meta.data %>% head()

row.names(scrna@meta.data) <- scrna@meta.data$CellID
 DimPlot(scrna, reduction = "umap", label = TRUE, pt.size = 0.5)

 saveRDS(scrna, file = "PIPSeq_ImmuneSub_Annotated_v0823.RDS")

```

# add the frequency
```{r}
table(IMMcells@meta.data$DataSet)

IMMcellsPropotionWithinTime <- table(IMMcells@meta.data$DataSet,Idents(IMMcells)) %>% as.data.frame() %>%
  group_by(Var1) %>%
  mutate(pct= prop.table(Freq) )  %>% ggplot(aes(x=Var1, y=pct, fill=Var2)) +geom_bar(position="stack", stat="identity") + theme_classic() +scale_y_continuous(labels = scales::percent)
ggsave("IMMcellsPropotionWithinTime.pdf", plot = IMMcellsPropotionWithinTime, width = 4, height = 3)

```

# check the markers
```{r}
# macrophage 1 and macrophage 2:
CFilePath<- paste0(Outdir,"Candidatemarkers")

dir.create(file.path(CFilePath))
setwd(file.path(CFilePath))

# M1 can choose CD80, CD86, CD64, CD16 and CD32 as markers. 
Macrophage2TypeMarkers <- c("Nos2","Cd79a","Cxcr6","Lef1","Ccl5","Capg","Ccl3","Ccl4","Ly6d","Lyz2","C1qa","Ifitm1","Ifngr1","Il17a","Cd80","Cd86","Fcgr1","Fcgr3","Fcgr2b","Cd163","Mrc1","Arg1", "Cd68","Havcr1","Lcn2")

IMMcells2BiomarkerDotPlot <- DotPlot(IMMcells, features = Macrophage2TypeMarkers) +RotatedAxis()
ggsave("IMMcells2BiomarkerDotPlot.pdf", plot = IMMcells2BiomarkerDotPlot, height = 5, width = 8)

Macrophage1Markers <- c("Nos2","Ifngr1","Il17a","Cd80","Cd86","Fcgr1","Fcgr3","Fcgr2b")

IMMcellsT1BiomarkerDotPlot<- DotPlot(IMMcells, features = Macrophage1Markers) +RotatedAxis()
ggsave("IMMcellsT1BiomarkerDotPlot.pdf", plot = IMMcellsT1BiomarkerDotPlot, height = 4, width = 6)

# M2 can choose CD163 and CD206 are major markers, Arg1 DECTIN
Macrophage2Markers <- c("Il13","Cd163","Mrc1","Arg1", "Cd68")
IMMcellsT2BiomarkerDotPlot<- DotPlot(IMMcells, features = Macrophage2Markers) +RotatedAxis()
ggsave("IMMcellsT2BiomarkerDotPlot.pdf", plot = IMMcellsT2BiomarkerDotPlot, height = 4, width = 6)


for (i in Macrophage1Markers){

  FeaturePlots<-FeaturePlot(IMMcells, features =  i, split.by = "DataSet" ) +RotatedAxis()

  ggsave(paste0(Outdir,"/CandidateMarkers/", i, "_M1_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)

}

for (i in Macrophage2Markers){

  FeaturePlots<-FeaturePlot(IMMcells, features =  i, split.by = "DataSet" ) +RotatedAxis()

  ggsave(paste0(Outdir,"/CandidateMarkers/", i, "_M2_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)

}
#
# setwd(file.path(Outdir))


```

```{r}
# contribution of monocyte-derived macrophage to inflammation

InflammationMarkers<- c("Cxcl1","Ccl4", "Ccl2", "Cxcl2")
GrowthFactors<- c("Csf3","Csf2")
InjuredMarkers <- c("Havcr1","Lcn2")
CytokinesMarkers<- c("Il23a","Ifngr1","Ifngr2","Il12","Il1a","Tnf","Il6","Il1f6")
Signaling <- c("Myd88","Tlr2","Tlr4","Nlrp3","Nfkb1","Nos2","Cybb")
Profibrotic <- c("Mmp2","Mmp9","Tgfb1","Col4a1","Col1a1","Fn1","Col3a1")

InflammationMarkersDotPlot<- DotPlot(IMMcells, features = InflammationMarkers) +RotatedAxis()
GrowthFactorsDotPlot<- DotPlot(IMMcells, features = GrowthFactors) +RotatedAxis()
InjuredMarkersDotPlot<- DotPlot(IMMcells, features = InjuredMarkers) +RotatedAxis()
CytokinesMarkersDotPlot<- DotPlot(IMMcells, features = CytokinesMarkers) +RotatedAxis()
SignalingDotPlot<- DotPlot(IMMcells, features = Signaling) +RotatedAxis()
ProfibroticDotPlot<- DotPlot(IMMcells, features = Profibrotic) +RotatedAxis()
ggsave("InflammationMarkersDotPlot.pdf", plot = InflammationMarkersDotPlot, height = 4, width = 6)
ggsave("GrowthFactorsDotPlot.pdf", plot = GrowthFactorsDotPlot, height = 4, width = 6)
ggsave("InjuredMarkersDotPlot.pdf", plot = InjuredMarkersDotPlot, height = 4, width = 6)
ggsave("CytokinesMarkersDotPlot.pdf", plot = CytokinesMarkersDotPlot, height = 4, width = 6)
ggsave("SignalingDotPlot.pdf", plot = SignalingDotPlot, height = 4, width = 6)
ggsave("ProfibroticDotPlot.pdf", plot = ProfibroticDotPlot, height = 4, width = 6)

```


```{r}
# the healthy cells
HealthMarkers<- c("Slc22a30", "Atp11a", "Trim7", "Slc22a6", "Slc7a13", "Slc5a2", "Slc5a10", "Spp2",  "Gatm", "Dcxr")

DotPlot(IMMcells, features =  HealthMarkers, col.min = -0.5) +RotatedAxis()

CyclingMarker <- "Mki67"
DotPlot(IMMcells, features =  CyclingMarker, col.min = -0.5) +RotatedAxis()

EarlyInjuryMarkers <- c("Cdh13", "Cdh6", "Havcr1")
DotPlot(IMMcells, features =  EarlyInjuryMarkers, col.min = -0.5) +RotatedAxis()
LateInjuryMarkers <-c( "Ccl2", "Gm17268", "Cald1")
DotPlot(IMMcells, features =  LateInjuryMarkers, col.min = -0.5) +RotatedAxis()

ReparingMarkers <- "Top2a"

DotPlot(IMMcells, features =  ReparingMarkers, col.min = -0.5) +RotatedAxis()

FailedFepairMarkers <- c("Ypel2","Kcnip4")

DotPlot(IMMcells, features =  FailedFepairMarkers, col.min = -0.5) +RotatedAxis()
# 
MaladptiveMarkers <- c("Vcam1", "Havcr1",  "Icam1", "Prom1","Cxcl2","Ccl3","Myc")

DotPlot(IMMcells, features =  MaladptiveMarkers, col.min = -0.5) +RotatedAxis()


# Combined 
#"Lrp2","Slc34a1","Slc13a1"
KidneyStatuesMarkers<- c( "Slc5a2","Slc34a1","Inmt","Apom", "Slc7a13","Slc22a6", "Prom1","Top2a","Cyp7b1","Mki67","Cdh13", "Cdh6", "Havcr1","Ccl2", "Gm17268", "Cald1","Vcam1", "Ypel2","Kcnip4","Hsp90aa1","Col27a1","Plin2")

 DotPlot(IMMcells, features =  KidneyStatuesMarkers,col.min = -1) +RotatedAxis()

KidneyStatuesTypeMarkers<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "Spp1", "Cst3","Gdf15", "Clu", "Igfbp7", "Apoe", "Aldob", "B2m", "S100a6","S100a8","S100a9","Pdgfb","Bak1","Bad","Casp8","Ptgs2","Casp1","Hmox1")
 DotPlot(IMMcells, features =  KidneyStatuesTypeMarkers) +RotatedAxis()
 
 KidneyStatuesTypeMarkers2<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "S100a6","S100a8","S100a9","Pdgfb","Bak1","Bad","Casp8","Ptgs2","Casp1","Hmox1","Nfkbia")
  DotPlot(IMMcells, features =  KidneyStatuesTypeMarkers2) +RotatedAxis()

 #, cluster 5: Fn1, Havcr1, Cst3, Igfbp7, S100a8, Il1b, Slc7a13 : Profibrotic PT
 # cluster 1: Spp1, Prom1, Kcnip4, Col27a1,Top2a,Nfkbia: Cycling PT
 # cluster 8: Cdh13, Hsp90aa1,Inmt, Mki67 : Proliferating PT
 
 # cluster 1: ASpp1, Prom1, Kcnipp4: Progenitor PT
 
 # cluster 0: Aldob
```

# the biomarkers for the 
```{r}
CandidateMarkers <- c("Aldob","Slc34a1","Spp1","Spp2","Lcn2", "Slc22a8","Slc5a2","Inmt","Prom1","Defb1","Top2a","Nfkbia", "Kcnip4", "Slc22a6", "Apoe","Havcr1", "Fn1","C3", "Il1b","S100a8", "Igfbp7", "Cst3", "Col27a1", "Hsp90aa1","Cdh13")

CandidateMarkersDotPlot<-  DotPlot(IMMcells, features =  CandidateMarkers, col.min = 0) +RotatedAxis()
CandidateMarkersDotPlot
ggsave("IMM_CandidateMarkersDotPlot.pdf", plot = CandidateMarkersDotPlot, height = 6, width = 12)
```

# #############################################
## generate the Trajectories of Macrophage cells
# #############################################
# 2.2 Prepared cds the object for Macrophage cells
```{r}
# build cds object for MYOLEOD
setwd(PyseudotimeDir)
Idents(IMMcells)
IMMcells@meta.data %>% head()

MACCells <- 
subset(IMMcells, subset =  ImmuneIntegrateAnnot == "M1_M2" | ImmuneIntegrateAnnot == "RecMAC")



MACCells_expr_matrix = GetAssayData(MACCells, slot = "counts")

# Cell meta annotation
p_data <- MACCells@meta.data 
head(p_data)

# Gene meta annotation
f_data<- data.frame(gene_short_name = row.names(MACCells), row.names = row.names(MACCells))


pre_cds <- new_cell_data_set(MACCells_expr_matrix,
                         cell_metadata = p_data,
                         gene_metadata = f_data)

pData(pre_cds) %>% colnames()

cds <- preprocess_cds(pre_cds) #标准化+PCA降维

```

# 2.2 Umap by Monocle3
```{r}
cds <- reduce_dimension(cds, preprocess_method = "PCA")
head(cds@int_colData$reducedDims$UMAP)

plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="DataSet")
dev.off()
# UMAP by Seurat
seurat_umap <- Embeddings(MACCells, reduction = "umap")[colnames(cds),]
cds@int_colData$reducedDims$UMAP <- seurat_umap
head(cds@int_colData$reducedDims$UMAP)
plot_cells(cds, reduction_method="UMAP",
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE,
           color_cells_by="DataSet")

```

# Draw the monocle state
```{r}
#分群(类似monocle的State)

cds <- cluster_cells(cds)

#预测轨迹
cds <- learn_graph(cds)
#交互式确定root节点，可以选择多个。我这里选择了一个
#cds <- order_cells(cds) 
# a helper function to identify the root principal points:
get_earliest_principal_node <- function(cds, time_bin="0DPI"){
  cell_ids <- which(colData(cds)[, "DataSet"] == time_bin)
  
  closest_vertex <-
  cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
  igraph::V(principal_graph(cds)[["UMAP"]])$name[as.numeric(names
  (which.max(table(closest_vertex[cell_ids,]))))]
  
  root_pr_nodes
}
cds <- order_cells(cds, root_pr_nodes=get_earliest_principal_node(cds))


MACCells_pseudotime <-plot_cells(cds, color_cells_by = "pseudotime", 
           label_cell_groups = FALSE, 
           label_leaves = TRUE,  
           label_branch_points = FALSE,graph_label_size=1.5)
MACCells_pseudotime
ggsave(filename = "MACCells_pseudotime_MonocaCluster.pdf", plot = MACCells_pseudotime, width = 4, height = 3)

MACCells_pseudotime_Celltype <- plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="ImmuneIntegrateAnnot")
ggsave(filename = "MACCells_pseudotime_Celltype.pdf", plot = MACCells_pseudotime_Celltype, width = 5, height = 3)
MACCells_pseudotime_Celltype
```

```{r}
# Determine the psedudotime related genes
Track_genes <- graph_test(cds, neighbor_graph="principal_graph")

Track_genes %>% head()

```

# plot the expression of interesting genes during the PT pseudotime
```{r}

# re-order the tracked genes by the q value and morans I
Track_genes <- Track_genes[,c(5,2,3,4,1,6)] %>% 
  dplyr::arrange(desc(morans_I),q_value)

# Plot some interesting genes for the PT pseudotime : Top 2 genes, Napsa, Kap, Ttr, Igfbp7, Havcr1 and Prom1
KapTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[1],] , 
                         min_expr=0.5)
ggsave(filename = "MAC_pseudotime_KapTrack.pdf", plot = KapTrack, width = 5, height = 3)
NapsaTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[2],] , 
                         min_expr=0.5)
ggsave(filename = "MAC_pseudotime_NapsaTrack.pdf", plot = NapsaTrack, width = 5, height = 3)

TtrTrack <- plot_genes_in_pseudotime(cds["Ttr",] , 
                         min_expr=0.5)
ggsave(filename = "MAC_pseudotime_TtrTrack.pdf", plot = TtrTrack, width = 5, height = 3)

ApoeTrack <- plot_genes_in_pseudotime(cds["Apoe",] , 
                         min_expr=0.5)
ggsave(filename = "MAC_pseudotime_ApoeTrack.pdf", plot = ApoeTrack, width = 5, height = 3)

Fn1Track <- plot_genes_in_pseudotime(cds["Fn1",] , 
                         min_expr=0.5)

ggsave(filename = "MAC_pseudotime_Fn1Track.pdf", plot = Fn1Track, width = 5, height = 3)

# For the 
Cst3Track <- plot_genes_in_pseudotime(cds["Cst3",],min_expr=0.5 )
ggsave(filename = "MAC_pseudotime_Cst3Track.pdf", plot = Cst3Track, width = 5, height = 3)

H2Ab1Track <- plot_genes_in_pseudotime(cds["H2-Ab1",] , 
                         min_expr=0.5)
ggsave(filename = "MAC_pseudotime_H2Ab1Track.pdf", plot = H2Ab1Track, width = 5, height = 3)

```

# choose the top signifcant for visualization
```{r}
diff_pseudo_gene <- Track_genes %>% 
  dplyr::arrange(q_value) %>%
  # remove the genes with mitochondrion
  filter(!grepl('^mt-|Hbb-', gene_short_name)) %>%
  rownames() %>% head(40)

# we ignore the mitochondria genes

diff_pseudo_gene

# generate the heatmap for those genes

diff_pseudo_gene.matrix <- exprs(cds)[match(diff_pseudo_gene,rownames(rowData(cds))),order(pseudotime(cds))]

diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){smooth.spline(x,df=3)$y}))
diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){(x-mean(x))/sd(x)}))
rownames(diff_pseudo_gene.matrix) <- diff_pseudo_gene;

#pheatmap(diff_pseudo_gene.matrix)

# because the plot_pseudotime_heatmap no longer have, we used the complexheatmap to draw the figure
# Reference: https://github.com/cole-trapnell-lab/monocle-release/issues/295
pdf("MAC_pseudotime_psudotimeGenes.pdf", height = 6, width = 8)
htkm <- Heatmap(
  diff_pseudo_gene.matrix,
  name    = "z-score",
  col     = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names  = TRUE,
  show_column_names = FALSE,
  row_names_gp  = gpar(fontsize = 6),
  km = 8,
  row_title_rot = 0,
  use_raster=TRUE,
  cluster_rows  = TRUE,
  cluster_row_slices = FALSE,
  cluster_columns   = FALSE)

print (htkm)
dev.off()
#ggsave(filename = "PT_pseudotime_psudotimeGenes.pdf", plot = htkm, width = 8, height = 6)


#Ward.D2 Hierarchical Clustering
pdf("MAC_pseudotime_psudotimeGenes_Hierarchicalcluster.pdf", height = 6, width = 8)
hthc <- Heatmap(
  diff_pseudo_gene.matrix,
  name                         = "z-score",
  col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names               = TRUE,
  show_column_names            = FALSE,
  row_names_gp                 = gpar(fontsize = 6),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  use_raster= TRUE,
  cluster_rows                 = TRUE,
  cluster_row_slices           = FALSE,
  cluster_columns              = FALSE)

print(hthc)
dev.off()

cds["Havcr1",] %>% head()
# Generate the Dotplot for these genes
plot_genes_in_pseudotime(cds[c("Havcr1"),] , 
                         min_expr=0.5)

DotPlot(IMMcells, features =diff_pseudo_gene ) +RotatedAxis()

DotPlot(IMMcells, features =c("Spp1","Havcr1","Itgb1","Top2a",
                             "Prom1","Pdgfb","Defb1","Fn1","Col6a1"), col.min = -0.5 ) +RotatedAxis()

FeaturePlot(IMMcells, features = c("Lcn2","Defb1","Egf","Havcr1"))
# the following no longer have 

Track_genes
# p = plot_pseudotime_heatmap(cds[diff_pseudo_gene,], 
#                         num_clusters = 3,  # default 6
#                         return_heatmap=T)
# 

```


# #############################################
## generate the Trajectories of T cells
# #############################################
# 2.3 Prepared cds the object for T cells
```{r}
# build cds object for T
setwd(PyseudotimeDir)
Idents(IMMcells)
IMMcells@meta.data %>% head()

TCells <- 
subset(IMMcells, subset =   ImmuneIntegrateAnnot == "T")



TCells_expr_matrix = GetAssayData(TCells, slot = "counts")

# Cell meta annotation
p_data <- TCells@meta.data 
head(p_data)

# Gene meta annotation
f_data<- data.frame(gene_short_name = row.names(TCells), row.names = row.names(TCells))


pre_cds <- new_cell_data_set(TCells_expr_matrix,
                         cell_metadata = p_data,
                         gene_metadata = f_data)

pData(pre_cds) %>% colnames()

cds <- preprocess_cds(pre_cds) #标准化+PCA降维

```

# 2.2 Umap by Monocle3
```{r}
cds <- reduce_dimension(cds, preprocess_method = "PCA")
head(cds@int_colData$reducedDims$UMAP)

plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="DataSet")
dev.off()
# UMAP by Seurat
seurat_umap <- Embeddings(TCells, reduction = "umap")[colnames(cds),]
cds@int_colData$reducedDims$UMAP <- seurat_umap
head(cds@int_colData$reducedDims$UMAP)
plot_cells(cds, reduction_method="UMAP",
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE,
           color_cells_by="DataSet")

```

# Draw the monocle state
```{r}
#分群(类似monocle的State)

cds <- cluster_cells(cds)

#预测轨迹
cds <- learn_graph(cds)
#交互式确定root节点，可以选择多个。我这里选择了一个
#cds <- order_cells(cds) 
# a helper function to identify the root principal points:
get_earliest_principal_node <- function(cds, time_bin="0DPI"){
  cell_ids <- which(colData(cds)[, "DataSet"] == time_bin)
  
  closest_vertex <-
  cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
  igraph::V(principal_graph(cds)[["UMAP"]])$name[as.numeric(names
  (which.max(table(closest_vertex[cell_ids,]))))]
  
  root_pr_nodes
}
cds <- order_cells(cds, root_pr_nodes=get_earliest_principal_node(cds))


TCells_pseudotime <-plot_cells(cds, color_cells_by = "pseudotime", 
           label_cell_groups = FALSE, 
           label_leaves = TRUE,  
           label_branch_points = FALSE,graph_label_size=1.5)
TCells_pseudotime
ggsave(filename = "TCells_pseudotime_MonocaCluster.pdf", plot = TCells_pseudotime, width = 4, height = 3)

TCells_pseudotime_Celltype <- plot_cells(cds, reduction_method="UMAP", 
           show_trajectory_graph = FALSE,
           label_cell_groups = FALSE, 
           color_cells_by="ImmuneIntegrateAnnot")
ggsave(filename = "TCells_pseudotime_Celltype.pdf", plot = TCells_pseudotime_Celltype, width = 5, height = 3)
TCells_pseudotime_Celltype
```

```{r}
# Determine the psedudotime related genes
Track_genes <- graph_test(cds, neighbor_graph="principal_graph")

Track_genes %>% head()
# re-order the tracked genes by the q value and morans I
Track_genes <- Track_genes[,c(5,2,3,4,1,6)] %>% 
  dplyr::arrange(desc(morans_I),q_value)

```

# choose the top signifcant for visualization
```{r}
diff_pseudo_gene <- Track_genes %>% 
  dplyr::arrange(q_value) %>%
  # remove the genes with mitochondrion
  filter(!grepl('^mt-|Hbb-', gene_short_name)) %>%
  rownames() %>% head(40)

# we ignore the mitochondria genes

diff_pseudo_gene

# generate the heatmap for those genes

diff_pseudo_gene.matrix <- exprs(cds)[match(diff_pseudo_gene,rownames(rowData(cds))),order(pseudotime(cds))]

diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){smooth.spline(x,df=3)$y}))
diff_pseudo_gene.matrix <- t(apply(diff_pseudo_gene.matrix,1,function(x){(x-mean(x))/sd(x)}))
rownames(diff_pseudo_gene.matrix) <- diff_pseudo_gene;

#pheatmap(diff_pseudo_gene.matrix)

# because the plot_pseudotime_heatmap no longer have, we used the complexheatmap to draw the figure
# Reference: https://github.com/cole-trapnell-lab/monocle-release/issues/295
pdf("Tcells_pseudotime_psudotimeGenes.pdf", height = 6, width = 8)
htkm <- Heatmap(
  diff_pseudo_gene.matrix,
  name    = "z-score",
  col     = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names  = TRUE,
  show_column_names = FALSE,
  row_names_gp  = gpar(fontsize = 6),
  km = 8,
  row_title_rot = 0,
  use_raster=TRUE,
  cluster_rows  = TRUE,
  cluster_row_slices = FALSE,
  cluster_columns   = FALSE)

print (htkm)
dev.off()
#ggsave(filename = "PT_pseudotime_psudotimeGenes.pdf", plot = htkm, width = 8, height = 6)


#Ward.D2 Hierarchical Clustering
pdf("Tcells_pseudotime_psudotimeGenes_Hierarchicalcluster.pdf", height = 6, width = 8)
hthc <- Heatmap(
  diff_pseudo_gene.matrix,
  name                         = "z-score",
  col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names               = TRUE,
  show_column_names            = FALSE,
  row_names_gp                 = gpar(fontsize = 6),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  use_raster= TRUE,
  cluster_rows                 = TRUE,
  cluster_row_slices           = FALSE,
  cluster_columns              = FALSE)

print(hthc)
dev.off()

cds["Havcr1",] %>% head()
# Generate the Dotplot for these genes
plot_genes_in_pseudotime(cds[c("Havcr1"),] , 
                         min_expr=0.5)

DotPlot(IMMcells, features =diff_pseudo_gene ) +RotatedAxis()

DotPlot(IMMcells, features =c("Spp1","Havcr1","Itgb1","Top2a",
                             "Prom1","Pdgfb","Defb1","Fn1","Col6a1"), col.min = -0.5 ) +RotatedAxis()

FeaturePlot(IMMcells, features = c("Lcn2","Defb1","Egf","Havcr1"))
# the following no longer have 

Track_genes
# p = plot_pseudotime_heatmap(cds[diff_pseudo_gene,], 
#                         num_clusters = 3,  # default 6
#                         return_heatmap=T)
# 

```


# plot the expression of interesting genes during the T cell pseudotime
```{r}

# Plot some interesting genes for the PT pseudotime : Top 2 genes, Napsa, Kap, Ttr, Igfbp7, Havcr1 and Prom1
KapTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[1],] , 
                         min_expr=0.5)
ggsave(filename = "Tcells_pseudotime_KapTrack.pdf", plot = KapTrack, width = 5, height = 3)
NapsaTrack <- plot_genes_in_pseudotime(cds[Track_genes$gene_short_name[2],] , 
                         min_expr=0.5)
ggsave(filename = "Tcells_pseudotime_NapsaTrack.pdf", plot = NapsaTrack, width = 5, height = 3)

TtrTrack <- plot_genes_in_pseudotime(cds["Ttr",] , 
                         min_expr=0.5)
ggsave(filename = "Tcells_pseudotime_TtrTrack.pdf", plot = TtrTrack, width = 5, height = 3)

Igfbp7Track <- plot_genes_in_pseudotime(cds["Igfbp7",] , 
                         min_expr=0.5)
ggsave(filename = "Tcells_pseudotime_Igfbp7Track.pdf", plot = Igfbp7Track, width = 5, height = 3)

Track_genes$gene_short_name[1:5]


Track_genes$gene_short_name[2]
Track_genes$gene_short_name[1]
Defb1Track <- plot_genes_in_pseudotime(cds["Defb1",] , 
                         min_expr=0.5)
Defb1Track
ggsave(filename = "Tcells_pseudotime_Defb1Track.pdf", plot = Defb1Track, width = 5, height = 3)

# For the 
Ly6c2Track <- plot_genes_in_pseudotime(cds["Ly6c2",],min_expr=1 )
ggsave(filename = "Tcells_pseudotime_Ly6c2Track.pdf", plot = Ly6c2Track, width = 5, height = 3)

Il17aTrack <- plot_genes_in_pseudotime(cds["Il17a",],min_expr=0.5 )
ggsave(filename = "Tcells_pseudotime_Il17aTrack.pdf", plot = Il17aTrack, width = 5, height = 3)


Ccl5Track <- plot_genes_in_pseudotime(cds["Ccl5",],min_expr=0.5 )
ggsave(filename = "Tcells_pseudotime_Ccl5Track.pdf", plot = Ccl5Track, width = 5, height = 3)
Cdh16Track
#FeaturePlot(IMMcells, features = "Havcr1")

```


# ###########################################################
## generate siganaling pathways for all the immune cells
# ###########################################################

```{r}
# build cds object for MYOLEOD
setwd(ProgenyDir)
Idents(IMMcells)
IMMcells@meta.data %>% head()



```


# #####################################
## generate the Progeny signaling pathways
# #####################################

```{r}
# set up the directory
setwd(ProgenyDir)
Idents(IMMcells)
IMMcells@meta.data %>% head()

options(future.globals.maxSize = 891289600000) 
## We compute the Progeny activity scores and add them to our Seurat object
## as a new assay called Progeny. 

# create cluster data frame with the cells, Celltypes
IMMCluster<- data.frame(Cell = names(Idents(IMMcells)),   
                              CellType = as.character(Idents(IMMcells)),
                              stringsAsFactors = FALSE)
IMMCluster %>% head()


IMMcells <- progeny(IMMcells, scale=FALSE, organism="Mouse", top=500, perm=1, return_assay = TRUE)
IMMcells
# setup the default assay as progeny
DefaultAssay(object = IMMcells) <- "progeny"

IMMcells <- Seurat::ScaleData(IMMcells, assay = "progeny") 
## We can now directly apply Seurat functions in our Progeny scores. 
## For instance, we scale the pathway activity scores. 
progeny_scores_df <- 
  as.data.frame(t(GetAssayData(IMMcells, slot = "scale.data", 
                               assay = "progeny"))) %>%
  rownames_to_column("Cell") %>%
  gather(Pathway, Activity, -Cell) 

progeny_scores_df %>% head()

#progeny_scores_df %>% filter(Cell =="0DPI_AAAAAAAACCACCAGA")
## We match Progeny scores with the cell clusters.
progeny_scores_df <- inner_join(progeny_scores_df, IMMCluster)

progeny_scores_df %>% head()
## We summarize the Progeny scores by cellpopulation
summarized_progeny_scores <- progeny_scores_df %>% 
  group_by(Pathway, CellType) %>%
  summarise(avg = mean(Activity), std = sd(Activity))
summarized_progeny_scores


## We prepare the data for the plot
summarized_progeny_scores_df <- summarized_progeny_scores %>%
  dplyr::select(-std) %>%   
  spread(Pathway, avg) %>%
  data.frame(row.names = 1, check.names = FALSE, stringsAsFactors = FALSE) 

paletteLength = 100
myColor = colorRampPalette(c("royalblue", "white","lightcoral"))(paletteLength)

progenyBreaks = c(seq(min(summarized_progeny_scores_df), 0, 
                      length.out=ceiling(paletteLength/2) + 1),
                  seq(max(summarized_progeny_scores_df)/paletteLength, 
                      max(summarized_progeny_scores_df), 
                      length.out=floor(paletteLength/2)))

summarized_progeny_scores_df
ImmuneCellForHeatmap<-as.data.frame(t(summarized_progeny_scores_df[,-1]))
ImmuneCellForHeatmap
# generate the heatmap for ImmuneCells
pdf(file="ImmuneCells_Progeny_ForHeatmap.pdf",height = 4, width = 5)
library(pheatmap)
progeny_hmap = pheatmap(ImmuneCellForHeatmap,fontsize=14, 
                        fontsize_row = 10, 
                        color=myColor, breaks = progenyBreaks, 
                        main = "PROGENy (500)", angle_col = 45,
                        treeheight_col = 0,  border_color = NA)
print(progeny_hmap)
dev.off()
test<- FeaturePlot(PTcells, features = "NFkB")
ggsave("test_NFKB.pdf", plot = test, width = 5, height = 5)
# generate the featureplot 
ProgenyFeatures <- summarized_progeny_scores$Pathway %>% unique()
ProgenyFeatures
for (i in ProgenyFeatures ){
  p2 <-(FeaturePlot(IMMcells, features = i,split.by = "DataSet") & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red'))& 
  ggtitle(paste0(i," activity"))
  
  ggsave(paste0(i, "_Immune_cluster_featurePlot.pdf"), plot = p2, width = 20, height = 5)
}


```

# Using PROGENy model from DecoupleR
```{r}
library(decoupleR)
setwd(ProgenyDir)
net <- get_progeny(organism = 'mouse', top = 500)

# Activity inference with Multivariate Linear Model (MLM)
# Extract the normalized log-transformed counts
mat <- as.matrix(IMMcells@assays$RNA@data)

# Run mlm
acts <- run_mlm(mat=mat, net=net, .source='source', .target='target',
                .mor='weight', minsize = 5)

acts 
# Visualization:
# Extract mlm and store it in pathwaysmlm in data
IMMcells[['pathwaysmlm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)


# Change assay
DefaultAssay(object = IMMcells) <- "pathwaysmlm"

# Scale the data
IMMcells <- ScaleData(IMMcells)
IMMcells@assays$pathwaysmlm@data <- IMMcells@assays$pathwaysmlm@scale.data

# run the source of pathways
ProgenyFeatures <- levels(factor(acts$source))
ProgenyFeatures
for (i in ProgenyFeatures ){
  p2 <-FeaturePlot(IMMcells, features = i) & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0(i," activity"))
  
  ggsave(paste0(i, "_IMMcells_cluster_featurePlot_DecoupleR.pdf"), plot = p2, width = 6, height = 5)
}

#(c("royalblue", "white","lightcoral"))
FeaturePlot(IMMcells, features = "Hypoxia", min.cutoff = -4, max.cutoff = 6, split.by = "DataSet") & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0("Hypoxia"," activity"))

FeaturePlot(IMMcells, features = "NFkB", min.cutoff = -4, max.cutoff = 6, split.by = "DataSet") & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0("NFkB"," activity"))


FeaturePlot(IMMcells, features = "TNFa", min.cutoff = -4, max.cutoff = 6, split.by = "DataSet") & 
 scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')& 
  ggtitle(paste0("TNFa"," activity"))

p1 <- DimPlot(IMMcells, reduction = "umap", label = TRUE, pt.size = 0.5) + 
  NoLegend() + ggtitle('Cell types')
p2 <- (FeaturePlot(IMMcells, features = c("Trail")) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) +
  ggtitle('Trail activity')



# Expore the mean activiey per group across pathways:
# Extract activities from object as a long dataframe
df <- t(as.matrix(IMMcells@assays$pathwaysmlm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(IMMcells)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Transform to wide matrix
top_acts_mat <- df %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-0.5, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 0.5, length.out=floor(palette_length/2)))

# Plot
IMMcellsHeatmap<- pheatmap(t(top_acts_mat), border_color = NA, color=my_color, breaks = my_breaks) 
ggsave("IMMcellscellsHeatmap_Progeny_Decouple.pdf", plot = IMMcellsHeatmap, width = 6, height = 6)

```

# Transcription factor activity inference from PT scRNA-seq
```{r}
# CollecTRI network
# set up the directory
setwd(TFDir)
netTF <-  readRDS("/Users/XXW004/Documents/Projects/RuizRosado/PIPseq/Datasets/get_collectri_human_mouse20230716/get_collectri_mouse_process_table.rds")

mat <- as.matrix(IMMcells@assays$RNA@data)
# Run ulm
acts <- run_ulm(mat=mat, net=netTF, .source='source', .target='target',
                .mor='mor', minsize = 5)

saveRDS(acts, file="Immune_Transcriptfactor.RDS")
# Activity inference with Univariate Linear Model (ULM)
# Extract ulm and store it in tfsulm in pbmc
IMMcells[['tfsulm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = IMMcells) <- "tfsulm"

# Scale the data
IMMcells <- ScaleData(IMMcells)
IMMcells@assays$tfsulm@data <- IMMcells@assays$tfsulm@scale.data

p1 <- DimPlot(IMMcells, reduction = "umap", label = TRUE, pt.size = 0.5) + 
  NoLegend() + ggtitle('Cell types')
p2 <- (FeaturePlot(IMMcells, features = c("PAX5")) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) +
  ggtitle('PAX5 activity')
DefaultAssay(object = IMMcells) <- "RNA"
p3 <- FeaturePlot(data, features = c("PAX5")) + ggtitle('PAX5 expression')
DefaultAssay(object = IMMcells) <- "tfswmean"
p1 | p2 | p3

# Explore the top 20 more varaible TFs
n_tfs <- 25

t(as.matrix(IMMcells@assays$tfsulm@data))
# Extract activities from object as a long dataframe
df <- t(as.matrix(IMMcells@assays$tfsulm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(IMMcells)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Get top tfs with more variable means across clusters
tfs <- df %>%
  group_by(source) %>%
  summarise(std = sd(mean)) %>%
  arrange(-abs(std)) %>%
  head(n_tfs) %>%
  pull(source)

# Subset long data frame to top tfs and transform to wide matrix
top_acts_mat <- df %>%
  filter(source %in% tfs) %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-3, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 3, length.out=floor(palette_length/2)))

# Plot
IMMTFheatmap<- pheatmap(top_acts_mat, border_color = NA, color=my_color, breaks = my_breaks) 
ggsave("IMMcellsHeatmap_Transcriptfactor_Progeny_Decouple.pdf", plot = IMMTFheatmap, width = 8, height = 5)
```




```{r}
# dividing into cycling, transitioning, adap- tive (successful or maladaptive repair) and degenerative (damaged or stressed). 
# cycling: Mki67
# Early Injury: Cdh13, Cdh6, Havcr1
# Late Injury: Ccl2, Gm17268, Cald1
# conserved: Slc13a1, Slc5a2,  Slc7a13, Cyp7b1



# Panoptosis
# Il1b, MLKL, Gsdmd,Ripk3

# Fibrosis 
# Fn1, Col1a1, Col3a1, Tgfb1
# adaptive
# VCAM1, DCDC2 and HAVCR1, C3, ICAM1
# VCAM1 PROM1 ITGB8
# Vcam1, Havcr1, Dcdc2, Icam1, Prom1, ITgb8

#degenerative
# SPP1, CST3, CLU and IGFBP7, APOE ALDOB
# B2M S100A6
# Spp1, Cst3, GDF15,Clu, Igfbp7, Apoe, Aldob, B2m, S100a6
# agionegious PEC,
# EGFR, VEGF, Cdh5

# Egfr, Vegf, Cdh5
KidneyStatuesMarkers<- c("Slc13a1", "Slc5a2", "Slc7a13", "Cyp7b1","Mki67","Cdh13", "Cdh6", "Havcr1","Ccl2", "Gm17268", "Cald1")

KidneyStatuesTypeMarkers<- c("Il1b", "Mlkl", "Gsdmd","Ripk3","Fn1", "Col1a1", "Col3a1", "Tgfb1","Vcam1", "Havcr1", "Icam1", "Prom1", "Spp1", "Cst3","Gdf15", "Clu", "Igfbp7", "Apoe", "Aldob", "B2m", "S100a6")

DotPlot(scrna, features =  KidneyStatuesMarkers, col.min = -0.5) +RotatedAxis()
DotPlot(scrna, features =  KidneyStatuesTypeMarkers, col.min = -0.5) +RotatedAxis()


for (i in KidneyStatuesMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}

for (i in KidneyStatuesTypeMarkers){
# 
  FeaturePlots<-FeaturePlot(scrna, features =  i, split.by = "DataSet" ) +RotatedAxis()
  ggsave(paste0(Outdir,"/KidneyStatuesTypeMarkers/", i, "_Feature_GeneMarkers_Annoatation.pdf"), plot =FeaturePlots, height =6, width = 14)
}


# with the inflammation markers

# with the fibrosis markers

# with the apoptosis markers

# with the necrosis markers

# with the endothelial markers


```


